### 2017-04-07 draft/ms_v1.tex
Draft version 1

% Version 0; Only rough outline by SH
% Version 1; First draft by SH

%% ------------------------------------------------------------------------------------ %% 
\documentclass[useamsfonts]{pasj01}
\twocolumn

%% ------------------------------------------------------------------------------------ %% 
% Packages
\usepackage{url}    
\usepackage{natbib}
\usepackage{latexsym}
\usepackage{morefloats}
\usepackage{xspace}
\usepackage[usenames, dvipsnames]{xcolor}

% Package Settings
%\usepackage{hyperref}
%\hypersetup{colorlinks=true,
%            citecolor=cyan,
%            linkcolor=cyan,
%            filecolor=magenta,      
%            urlcolor=cyan}
%\urlstyle{same}
%% ------------------------------------------------------------------------------------ %%

% Figure extension
\DeclareGraphicsExtensions{.pdf,.png,.eps}

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\amin{$^\prime$}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over \sim \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over \sim \;$}}
\def\micron{{$\mu$m}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$}
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}

\def\logms{{$\log (M_{\star}/M_{\odot})$}~}
\def\mstar{{$M_{\star}$}~}
\def\logmh{{$\log (M_{\mathrm{halo}}/M_{\odot})$}~}
\def\m2l{{$M_{\star}/L_{\star}$}~}

\def\redm{\texttt{redMaPPer}}
\def\hscpipe{\texttt{hscPipe}}
\def\synpipe{\texttt{SynPipe}}
\def\cmodel{\texttt{cModel}}
\def\forced{\texttt{forced}}
\def\unforced{\texttt{unforced}}
\def\tract{\texttt{Tract}}
\def\visit{\texttt{Visit}}
\def\tracts{\texttt{Tracts}}
\def\visits{\texttt{Visits}}
\def\galsim{\texttt{G}{\scriptsize \texttt{AL}}\texttt{S}{\scriptsize \texttt{IM}}}
\def\hst{{\textit{HST}}}

%% ------------------------------------------------------------------------------------ %%

% Commenting:
%% ------------------------------------------------------------------------------------ %% 
\newcommand{\plan}[1]{\textcolor{blue} {\textbf{#1}}}
\newcommand{\todo}[1]{\textcolor{red} {\textbf{#1}}}
\newcommand{\song}[1]{\textcolor{cyan} {\textbf{#1}}}
\newcommand{\addref}{{\textcolor{red}{\textbf{REF}}}}
\newcommand{\term}[1]{\textbf{\texttt{#1}}}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %%

\begin{document}

\SetRunningHead{Huang et al.}{HSC-SSP SynPipe}

\Received{$\langle$April 2017$\rangle$}
\Accepted{$\langle$2017$\rangle$}
\Published{$\langle$2017$\rangle$}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\title{HSC Synthetic Pipeline I: Basic Design and Characterization of the 
       Photometric Performance of HSC}

\author{Song Huang \altaffilmark{1,2}
        Alexie Leauthaud \altaffilmark{1,2},
        Ryoma Murata \altaffilmark{1, 4}, 
        Jim Bosch \altaffilmark{3},
        Paul Price \altaffilmark{3},
        Robert Lupton \altaffilmark{3},
        Rachel Mandelbaum \altaffilmark{5},
        Claire Lackner \altaffilmark{1},
        Steven Bickerton \altaffilmark{1},
        Satoshi Miyazaki \altaffilmark{6} 
        HSC Collaboration \altaffilmark{7}}                                          
%% ------------------------------------------------------------------------------------ %%

\altaffiltext{1}{Kavli Institute for the Physics and Mathematics of the 
    Universe, The University of Tokyo Institutes for Advanced Study, 
    the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan}
\altaffiltext{2}{Department of Astronomy and Astrophysics, University of California, 
    Santa Cruz, 1156 High Street, Santa Cruz, CA 95064 USA}
\altaffiltext{3}{Department of Astrophysical Sciences, Princeton University, 
    4 Ivy Lane, Princeton, NJ 08544}
\altaffiltext{4}{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
\altaffiltext{5}{McWilliams Center for Cosmology, Department of Physics, 
    Carnegie Mellon University, Pittsburgh, PA 15213, USA}
\altaffiltext{6}{SOKENDAI (The Graduate University for Advanced Studies), Mitaka,
    Tokyo, 181--8588, Japan}
\altaffiltext{7}{Other affiliations}

\email{song.huang@ipmu.jp}
%% ------------------------------------------------------------------------------------ %%

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\KeyWords{Surveys, 
          Methods: observational, 
          Techniques: photometric}

\maketitle

%% ------------------------------------------------------------------------------------ %% 
\begin{abstract}

    The Subaru Strategy Survey Program (SSP) is an ambitious wide-area, multi-band 
    survey using the Hyper Suprime-Cam (HSC) on Subaru telescope.
    At the imaging depth of this survey, it becomes challenging to achieve accurate,
    unbiased, and consistent photometry across all five bands. 
    In this work, we present \synpipe{}, a \texttt{Python}-based, flexible framework 
    that interfaces with the HSC data reduction pipeline (\hscpipe{}) and test its 
    performance via injecting synthetic objects to the single-\texttt{visit} HSC images. 
    We use the coadd images in two \tracts{} from the Wide layer with representative 
    seeing conditions to demonstrate the usage of \synpipe{} and examine the general 
    photometric performance of \texttt{hscpipe}. 
    We find that the typical uncertainties of HSC \forced{} PSF photometry for stars 
    range from ${\sim}0.01$ mag at $i{\sim}18.0$ mag to ${\sim}0.06$ at $i{\sim}25.5$ 
    mag (1\%-5\% accuracy in the $i$-band). 
    For synthetic single-{S\'{e}rsic\ } model galaxies, the typical uncertainties of 
    HSC forced \cmodel{} photometry range from 0.14 mag at $i{\sim}20.0$ mag to 
    0.20 mag at $i{\sim} 25.2$ mag. 
    Over the range of colors and magnitudes in these tests, we find that both \forced{} 
    PSF and \cmodel{} photometry provide unbiased estimates of colors that are robust 
    against different seeing conditions. 
    Finally, we also noticed that 1) the degree that a object is blended with others 
    impact the photometric accuracy, especially for stars.  
    Highly blended object tends to have larger photometric uncertainty, systematically 
    underestimated flux, and slightly biased color. 
    2) a relative high fraction of faint stars (${\sim}20$\% at $i>22.5$ mag) could be 
    misclassified as extended objects. 
    3) the shape and size estimates for galaxies using \cmodel{} suffer from strong 
    biases.

\end{abstract}
%% ------------------------------------------------------------------------------------ %%

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\section{Introduction}
    \label{sec:intro}

    %% Background 
    Wide-field, multi-band imaging surveys have stepped onto the central stage of 
    modern astrophysics and cosmology in the past decade. 
    They also have a bright future as many ambitious projects are being carried out 
    or planned, like the Large Synoptic Survey Telescope 
    (LSST)\footnote{\url{https://www.lsst.org/}}, the Wide-Field Infrared Survey 
    Telescope 
    (WFIRST; \citealt{Dressler2012})\footnote{\url{https://wfirst.gsfc.nasa.gov/}}, 
    and the \textit{Euclid} project 
    (\citealt{Laureijs2012})\footnote{\url{http://sci.esa.int/euclid/}}.
    Among many ongoing efforts, the Subaru Strategic Program 
    (SSP; \citealt{HSCDR1})\footnote{\url{http://hsc.mtk.nao.ac.jp/ssp/}}
    using the Hyper Suprime-Cam (HSC; \citealt{Miyazaki2012}) on the prime focus of 
    the Subaru telescope is the most efficient one in term of Etendue. 
    These surveys together will greatly help us test cosmological models, trace the 
    evolution of galaxies, study the stellar structure of our Milky Way, and discover 
    large number of interesting transient objects.

    %% Challenges
    Before we can tackle any outstanding scientific question, we first need to learn 
    how to handle the huge amount of data generated by these projects, and achieve 
    accurate photometric measurements of different types of objects.  
    This target becomes increasingly challenging in the age of modern imaging 
    surveys. 
    The cameras that carry these surveys are often made up of multiple 
    charge-coupled devices (CCDs) with slightly different characteristics and have 
    large field-of-view (FoV) over not perfectly flat focus plane.
    During the observation, the seeing and background conditions always show spatial 
    and temporal variations across the FoV.
    The full-depletion, thick CCD adopted by these cameras enables long exposure time 
    and excellent sensitivity in the red-end, but they often have the so-called 
    ``brighter-fatter'' effect (\citealt{Antilogus2014}) which means the size of bright
    stars are enlarged.
    All these issues make accurate astrometric and photometric calibration, 
    background subtraction, and point spread function (PSF) modeling intrinsically
    difficult (e.g. \citealt{Schlafly2012}).
    In addition, the multi-epoch, multi-band nature of these surveys also complicate 
    the data reduction as one needs to pay special attentions to the processes of 
    image stacking, multi-band object detection and measurements. 
    Furthermore, as ongoing surveys reach to much deeper surface brightness level, 
    the number of objects in unit area, especially faint object, becomes much higher.
    Not only that, it is much more common to have objects that overlap each other at
    different levels (which is often referred to as ``blended''). 
    Both situations ask for improved methods for photometric measurements than the ones
    used in previous, shallower surveys.

%% ------------------------------------------------------------------------------------ %%     
    %% Reason
    These challenges mentioned above are not just technical details, they are crucial 
    systematics to key scientific goals of these surveys. 
    For instance, weak gravitational lensing (WL; \citealt{Kaiser1993, Bartelmann2001}) 
    is a powerful tool to measure the large-scale distribution of dark matter and 
    follow its evolution. 
    The WL measurements critically depend on the ability to model PSF and measure the
    shape of background galaxies in high precision (e.g. \citealt{Mandelbaum2015}). 
    Also, photometric redshift measurement (e.g. \citealt{Benitez2000, Bolzonella2000,
    Ilbert2009}) is fundamentally important for studying the evolution of galaxies; 
    and the ``drop-out'' method is very useful in selecting high-redshift galaxies 
    (e.g. \citealt{Steidel1996}). 
    Both methods are based on accurate multi-band photometric measurements. 
    The list goes on and on, but it becomes clear that we need to develop frameworks 
    to oversee the data reduction process of modern imaging surveys. 
    Several tools with similar goals have been developed (e.g. \citealt{Chang2015, 
    Suchyta2016}) for the Dark Energy Survey (DES; \citealt{DES2005}).

    %% Organization    
    In this work, we present \synpipe{}, a \texttt{Python}-based 
    software to test the data reduction quality of the HSC--SSP (will be referred to 
    as the HSC survey) using synthetic objects.   
    \synpipe{} is designed to be a flexible framework that has many applications 
    and can help future developments of the HSC data reduction pipeline.
    Here, after briefly introducing the HSC survey and the current status of 
    data reduction (Section \ref{sec:ssp}), we explain the design and implementation 
    of \synpipe{} (Section \ref{sec:synpipe}), then we demonstrate its usage via 
    straightforward tests (Section \ref{sec:test}) that focus on the general 
    photometric quality for both synthetic stars and galaxies 
    (Section \ref{sec:result}). 
    Finally, we summary the work and discuss future developments in the 
    Section \ref{sec:summary}.

    \synpipe{} code, along with basic documents and a few examples are available 
    on GitHub at: \url{https://github.com/dr_guangtou/synpipe}.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Flowchart
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_flowchart_new}
    \end{center}
    \caption{
        Illustration of the basic procedures of \synpipe{}.  
        Grey color boxes indicate the required inputs at different stages. 
        Blue box is for the \texttt{addFakes.py} step when \synpipe{} injects simulated
        images of synthetic objects onto single frame image. 
        We show an example CCD image with and without synthetic objects on it 
        (highlighted by green circles).
        Red boxes are for the following image coadding and multiband measurement steps
        using \texttt{stack.py} and \texttt{multiBand.py} from \hscpipe{}. 
        We show a small piece of coadd image with synthetic objects on it.    
        The figure on the right side of the middle row shows the spatial relation 
        between \tract{} (outlined by the black dashed lines; each small region 
        is a \texttt{Patch}) and all the \visits{} that contribute to it 
        (each \visit{} has its own unique color; each small region has the size 
        of a CCD).
        }
    \label{fig:flowchart}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Hyper-Suprime Cam Subaru Strategic Program}
    \label{sec:ssp}

    %% Basic
    Taking advantage of the new prime focus camera on the 8.2m Subaru telescope, 
    the ambitious HSC survey consists of three layers. 
    The Wide layer will map a total of ${\sim}1400$ deg$^2$ of sky in five broadband 
    ($grizy$; Kawanomoto\etal in prep.). 
    The Deep (4 separated fields; ${\sim}27$ deg$^2$) and UltraDeep (2 separated 
    fields; ${\sim}3.5$ deg$^2$) layers will use a few additional narrow-band filters 
    and employ slightly different survey strategy. 
    Please refers to the survey description paper for more detailed information
    (Strauss\etal in prep.).

    The HSC camera (\citealt{Miyazaki2012}) is made up of 112 full-depletion thick CCDs 
    for science and another 12 for guiding and focusing. 
    It provides a circular FoV with 1.5 deg diameter. 
    Each CCD contains 2048$\times$4096 pixels distributed in 4 amplifiers. 
    The pixel size equals to 0.168\asec{} on the sky.

    The HSC collaboration just released the first 1.7 years of data to public (DR1; 
    \citealt{HSCDR1})\footnote{\url{https://hsc_release.mtk.nao.ac.jp}}.
    For this work, we will focus on the Wide layer whose major scientific goals include
    WL cosmology, galaxy evolution since $z{\sim}1$, and studies of galaxy clusters. 
    The current Wide layer includes ${\sim}108$ deg$^2$ of sky that are isolated in 
    6 different regions (XMM--LSS, GAMA09H, WIDE12H, GAMA15H, HECTOMAP, and VVDS)
    \footnote{A small AEGIS field is also observed for calibration purpose.}.
    Most of these fields are close to the equator with the exception of the HECTOMAP 
    field. 
    In the Wide layer, the $g$ and $r$ bands have 10 mins exposure broken into 4 
    dithers\footnote{Large, ${\sim}1/3$ of FoV dither pattern is used to ensure 
    uniform coverage.}; for the rest 3 bands, 20 mins exposure time and 6 dithers are 
    used\footnote{Several 30s exposures are taken to improve the dynamic range.}. 
    The survey prioritizes the observations so that the $i$-band can have excellent 
    seeing condition to help accurate shape measurements for WL. 
    The current Wide layer has impressive median seeing with FWHM${\sim}0.6$\asec{}, 
    and has a $i{\sim} 26.4$ mag $5\sigma$ point source detection limit.
    For more details of the data status, please see the data release paper 
    (\citealt{HSCDR1}).  
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%     
\section{The HSC Data Reduction Pipeline}
    \label{ssec:hscpipe}

    The HSC survey data is currently reduced by \hscpipe{} \texttt{v4.0.5}, a 
    version of the LSST reduction pipeline (\citealt{Ivezic2008, Axelrod2010, 
    Juric2015}), which is described in detail in Bosch\etal (in prep.). 
    Since \synpipe{} is intrinsically connected to the complex data reduction 
    processes in \hscpipe{}, here we briefly introduce the main steps along with 
    a few key HSC/LSST terminologies.

    \begin{enumerate}

        \item \textbf{Single-\texttt{Visit~}Processing}: 
            Each individual exposure is called a \term{Visit} and is assigned with 
            an even integer. 
            After initial data screening (in background, seeing, and transparency; 
            Furusawa\etal in prep.), \hscpipe{} detrends these data (subtraction 
            of overscan, bias, and dark; flatfielding), generates variance and mask 
            images, subtracts background, and provides initial astrometric and 
            photometric calibrations.  
            \textbf{The spatially varying PSF models, photometric zeropoints, and the 
            World Coordinate System (WCS; \citealt{WCS1, WCS2}) corrected for optical 
            distortion will be used by \synpipe{}}.

        \item \textbf{Multi-\texttt{Visit~}Processing}:
            Then, \hscpipe{} then warps and mosaics these reduced CCD images into 
            much deeper \texttt{coadd} images while improving the astrometric and 
            photometric calibrations via processes similar to the \"{u}ber-calibration 
            in SDSS \citep{Padmanabhan2008}. 
            \textbf{\synpipe{} will use these improved calibrations.}
            \hscpipe{} organizes these \texttt{coadd} images into equi-area 
            rectangular regions that are pre-defined as an iso-latitude tessellation 
            and are referred to as \tract{}. 
            One \tract{} covers approximately $1.7\times 1.7$ deg$^2$ of 
            sky, and adjacent \tracts{} overlap each other by ${\sim}1$\amin{}.
            Each \tract{} is further divided into $9\times9$ \texttt{Patches}. 
            A \texttt{Patch} is a $4200\times4200$ pixels rectangular region, and
            adjacent \texttt{Patches} overlap each other by 100 pixels.
            \textbf{The \tract{} and \texttt{Patch} are basic units to deal 
            with HSC \texttt{coadd} images and \synpipe{} tests.}

        \item \textbf{Multi-band Measurements}: 
            To achieve consistent photometry across all filters, \hscpipe{} first 
            detects and deblends objects on \texttt{coadd} images in each band 
            independently (the \unforced{} measurements). 
            The above-threshold pixels for each object is referred to as 
            \texttt{footprint}.
            Then, \hscpipe{} merges the \texttt{footprints} and flux peaks in different
            bands, and selects a reference band for each object based on its $S/N$ in 
            each band (usually it is the $i$-band). 
            With the centroids, shape, and other non-amplitude parameters fixed at the
            reference values, \hscpipe{} perform \forced{} photometry which is 
            optimized to generate accurate colors.

        \item \textbf{HSC Photometry}: 
            The multiband catalogs generated by \hscpipe{} contains various photometric 
            information (see \citealt{HSCDR1} for details). 
            \textbf{Here, we focus on the \forced{} \texttt{psf} and \cmodel{} 
            photometry.}
            \texttt{psf} photometry should provide the most appropriate magnitudes and
            colors for point sources.
            Improved based on the SDSS version (\citealt{Lupton2001, Abazajian2004}), 
            the HSC \cmodel{}{} algorithm (described in Bosch\etal in prep.) fits the 
            flux distribution of each object using a combination of a de~Vaucouleur 
            and an exponential component taking the PSF convolution into accounts.
            \cmodel{} is designed to yield accurate colors for galaxies. 
            In addition, the star--galaxy separation task is performed based on the 
            difference between the \texttt{psf} and \cmodel{}{} magnitudes.

    \end{enumerate}

    The quality of the DR1 data have been vetted in \citet{HSCDR1} via cross-checking 
    with the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 1 
    imaging survey (\citealt{Schlafly2012, Tonry2012, Magnier2013}), examining the 
    behaviors of the stellar sequence and the shape of the PSF models. 
    Generally speaking, the PSF photometry is accurate to 1--2\% level, and the 
    astrometry is accurate to ${\sim}10$ and 40 max internally and externally.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{SynPipe Overview}
    \label{sec:synpipe}

%% ------------------------------------------------------------------------------------ %% 
\subsection{Basic Design of \synpipe{}}
    \label{ssec:design}

    As summarized above and described in greater detail in Bosch\etal (in prep.),
    \hscpipe{} is a complex data reduction pipeline that still faces many challenges 
    and is under active development.
    Unlike most photometric pipeline on the market, it involves high-level reduction 
    processes that produce not only deep \texttt{coadd} images but also a series of 
    science-ready catalogs. 
    To faithfully test \hscpipe{} and these data products, we make sure the 
    \synpipe{} satisfies a few basic standards:

    \begin{itemize}

        \item \textbf{Realistic Images}: 
            Instead of using the fully generative approach to simulate full HSC-like 
            images from the ground up (e.g. \citealt{Chang2015}), \synpipe{} injects
            synthetic objects to the observed image so that all the realistic 
            features of the real data (e.g. blended objects; approximity to bright 
            object, bleeding trails, other optical artifacts \etal) can be included 
            in the test. 
            The impact of these features can be important to any topic that cares 
            about the selection function and completeness of samples 
            (e.g. WL analysis; stellar mass function of galaxies). 
            In \synpipe{}, point source is simulated using the HSC PSF model, 
            and realistic model of galaxy is modelled by the modular galaxy image 
            simulation toolkit, \galsim{} (\citealt{Rowe2015})
            \footnote{\url{https://github.com/GalSim_developers/GalSim}}.

        \item \textbf{Authentic Data Processing}:
            To follow the data reduction process as close as possible, we choose to 
            start from the single-\visit{} images instead of directly putting
            synthetic objects on the final \texttt{coadd} images 
            (e.g. \citealt{Suchyta2016}).
            Through this approach, subtle effects like the seeing differences among 
            \visits{} and small errors in astrometric calibration can be taken 
            into accounts. 
            All the synthetic objects will experience all the steps for detection, 
            deblending, stacking, and measurement just like any real object. 
            This can be important for challenging tasks like the WL measurements
            or the detection of high-$z$ galaxies. 
            The only downside of this choice is that it will certainly slow down the 
            test for \texttt{coadd} images as it needs to inject synthetic object 
            to all the \visits{} that contribute to a \tract{}.

        \item \textbf{Flexible Capabilities}:
            To make \synpipe{} useful for HSC users with various scientific goals, 
            we make sure it is flexible enough. 
            Although we provide default catalogs to generate samples of synthetic stars 
            and galaxies with realistic magnitude and color distributions along with 
            tools to help work on data release of HSC survey, the user is allowed to 
            be creative in term of providing input catalogs and design the tests. 
            As we are speaking, \synpipe{} is being used in a wide range of 
            scientific applications (see Section \ref{sec:summary} for details).

    \end{itemize}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Implementation of \synpipe{}} 
    \label{ssec:flowchart}

    In this section, we will go through the implementation of \synpipe{} tests in 
    details (see Fig \ref{fig:flowchart}).

%% ------------------------------------------------------------------------------------ %%     
\subsubsection{Preparation}
    \label{sssec:prep}

    To start a \synpipe{} test, the user needs to provide:

    \begin{itemize}

        \item \textbf{Information about the data}: User needs to provide the 
            location of the HSC images and a list of \visits{} for the
            test.  
            For test of the \texttt{coadd} images, \synpipe{} can help user identify 
            all the \visits{} that contribute to one \tract{} in certain
            band. 
            It also provides tools to create optional bad-pixel mask (e.g. bright 
            object, bleeding trail, \etal) for specific \tract{} to avoid 
            putting synthetic objects on problematic regions.

        \item \textbf{Input catalog} in \texttt{FITS} format that contain necessary 
            information like the coordinate, magnitude, and other model parameters. 
            For certain sample of synthetic objects, \synpipe{} can help randomly 
            distribute them in a selected rectangular region with desired number 
            density. 
            If necessary, overlaps of synthetic objects can be greatly reduced 
            via setting a minimum distance between any two of them. 
            \synpipe{} can also help put synthetic objects on an evenly separated 
            coordinate grid, which is very useful for certain test 
            (e.g. Murata\etal in prep.).

    \end{itemize}

    For the synthetic object, \synpipe{} provides the following choices:

    \begin{itemize}

        \item \textbf{Point source}: 
            \synpipe{} simply uses the PSF model from \hscpipe{} as model for 
            point source like star or QSO. 
            \hscpipe{} uses a special version of \texttt{PSFEx} 
            (\citealt{Bertin2011, Bertin2013}) to characterize the PSF as a function 
            of the position so that a PSF model can be reconstructed for any location 
            on the image. 
            User only needs to provide the integrated magnitude to simulate 
            point sources.

        \item \textbf{Single- or Double-\ser{} model for galaxy}: 
            \texttt{SynPipe} uses \galsim{} \texttt{v1.4} (\citealt{Rowe2015}) to 
            simulate galaxy. 
            \galsim{} is an image simulation tool that has been well tested during 
            the GRavitational lEnsing Accuracy Testing 3 (GREAT3; 
            \citealt{Mandelbaum2014}). 
            Right now, \synpipe{} allows synthetic galaxy to be modelled by single 
            or double components that are described by the \citet{Sersic1963} 
            profile.  
            The \ser{} profile is flexible enough to describe the overall flux 
            distributions of galaxies near and far, early-type or late-type. 
            Double-\ser{} model can further simulate galaxy with more realistic 
            structural details (e.g. bulge$+$disk).
            For each \ser{} component, user needs to provide the magnitude, 
            effective radius ($R_{\mathrm{e}}$; in unit of arcsec), \ser{} index 
            ($n_{\mathrm{Ser}}$), axis ratio ($\mathrm{b}/\mathrm{a}$), and position 
            angle (PA).
            External shear can also be applied to synthetic galaxy ($g_1$ and $g_2$; 
            see \citealt{Rowe2015} for more details).

        \item \textbf{Models or real images of galaxy from the GREAT3 challenge}: 
            In stead of providing detailed model information for each galaxy, \synpipe{}
            also allows users to choose from the parametric models or real 
            high-resolution images of galaxies used in the GREAT3 WL challenge. 
            The images are from the \hst{}/ACS F814W ($I$-band) images of the galaxies 
            in the COSMOS field (e.g. \citealt{Scoville2007}). 
            These ACS exposures are 50\% completed down to $I=26.0$ mag in aperture with 
            0.5\asec{} diameter.  
            The depth and excellent resolution (0.09\asec FWHM) make it adequate to test 
            galaxies at Wide depth in HSC survey.
            In addition to the real cutout image, the COSMOS galaxies with 
            $I_{\mathrm{Auto}}\leq25.2$ mag are modelled via method developed by
            \citet{Lackner2012}, and the best parametric model for each of them is 
            selected between the single-\ser{} and the \ser{} bulge$+$Exponential disk 
            ones. 
            Information for these models are stored in the 
            \texttt{GalSim.COSMOSCatalog()} catalogs. 
            To use them, user can provide the ID number of the COSMOS galaxies in the 
            input catalog, or simply let \synpipe{} randomly select them for you.

    \end{itemize}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%     
\subsubsection{Injection of Synthetic Objects to Single-\visit{} Image}
    \label{sssec:addFakes}

    With the input catalog and information about the test data, \synpipe{} first 
    injects synthetic objects to single exposure images (step \texttt{addFakes.py}; 
    see the middle panel of Fig \ref{fig:flowchart}). 
    For each \visit{}, \synpipe{} goes through all the individual CCD images 
    that belong to it, and decides which synthetic objects from the input catalog need
    to be injected. 
    \synpipe{} makes use of the \textit{initial astrometric calibration} of each single 
    \visit{} to convert the input coordinates into locations on the CCD image. 
    Using the available models, \synpipe{} will reconstruct the PSF at the position of 
    each injected object.
    To convert the input magnitudes of synthetic objects into fluxes on the image, 
    \synpipe{} also uses the photometric zeropoint from single-\visit{} 
    calibration.

    For point sources, \synpipe{} generates a rectangular cutout image of the PSF model 
    with appropriate size and correct total flux. 
    For galaxy described by single- or double-\ser{} models, \synpipe{} passes the input 
    \ser{} index and effective radius to the \galsim{}\texttt{.Sersic} function to 
    create a \ser{} component.
    After stretching and rotating it to expected axis ratio and position angle 
    via the \texttt{shear} and \texttt{rotate} methods, \synpipe{} assigns correct flux 
    to this component. 
    For double-\ser{} model, \galsim{}\texttt{.Add} method is used to combine the two 
    \ser{} components together. 
    For parametric models from the GREAT3 catalog, \synpipe{} utilizes the 
    \texttt{COSMOSCatalog.makeGalaxy} method to generate corresponding model, 
    and the \galsim{}\texttt{.RealGalaxy} method is used for the case of real 
    high-resolution image. 
    If necessary, additional shear ($g_1$ and $g_2$) can be applied to the model at this 
    point. 
    Then, \synpipe{} passes the reconstructed PSF image into a 
    \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy model  
    using \galsim{}\texttt{.Convolve}. 
    After this, \synpipe{} uses \galsim{}\texttt{.drawImage} method to create the image 
    of the simulated galaxy. 
    Component with high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires large image 
    size to cover most of its total flux, so \synpipe{} allows user to truncate model 
    at certain radius (N$\times R_{\mathrm{e}}$). 
    It will also give warning information when component with $n_{\mathrm{Ser}} > 6$ is 
    encountered. 
    Such model takes much longer time to achieve accurate PSF convolution, and is usually 
    not realistic galaxy model.

    \synpipe{} then shifts the image of synthetic object according to is location on the 
    CCD on a pixel-by-pixel basis; it also crops the images when necessary. 
    Afterwards, \synpipe{} adds appropriate noise to the image based on the pixels it 
    covers and the calibration information of the detector, and create a variance image 
    that reflects the influence of the synthetic object. 
    Since each HSC CCD consists of four amplifiers with different characteristics, 
    \synpipe{} carefully creates corresponding gain map to provide accurate noise level. 
    Finally, \synpipe{} adds the noise-added image and the variance map to the original 
    CCD data while also adding a new \texttt{FAKE} mask bit to the mask plane. 
    On the middle panel of Fig \ref{fig:flowchart}, we demonstrate the result of the 
    \texttt{addFakes.py} step by showing an example CCD image before and after the 
    synthetic objects are injected. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%     
\subsubsection{Stacking and Multi-band Measurements}
    \label{sssec:multiband}

    These newly generated sing-\visits{} images have the same format and data 
    structure of the real HSC data, therefore \synpipe{} can directly feed them 
    to the rest steps of \hscpipe{}. 
    The \texttt{stack.py} step takes the improved astrometric and photometric 
    calibrations for each \visit{} from the original reduction and create 
    \texttt{coadd} images with synthetic objects on them. 
    Then the \texttt{multiBand.py} step will process these \texttt{coadd} images and 
    provide standard mulit-band measurements in FITS catalogs that are grouped by 
    \texttt{Patch} ID.

    Using the infrastructures provided by \hscpipe{}, the above processes can be 
    easily parallelized, and the overall efficiency of \synpipe{} is similar to the
    real data reduction process. 
    They also generate the same amount of images and catalog data with the original 
    one.

    As the last step, \synpipe{} matches the input catalog to the results within a 
    radius that is specified by the user, and merge all the \unforced{} and 
    \forced{} measurements from all the \texttt{Patches} together. 
    Unmatched object from the input catalog is also passed to the result catalog with 
    special label. 
    In case of multiple matches, user can choose between only returning the nearest 
    one or returning all objects within the matching radius (the nearest one is 
    labeled).  
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Star Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_star_sample}
    \end{center}
    \caption{
        Magnitude and color distributions of the synthetic stars. 
        \textbf{Upper panel} shows the 5-band magnitude distributions of the COSMOS 
        stars (filled histograms) and the resampled stars (dashed lines). 
        \textbf{Lower panel} shows their color-color distributions 
        (left: $g-r$ v.s. $r-i$; right: $i-z$ v.s. $z-y$). 
        We show the number density distribution of the COSMOS stars in hexagonal bins, 
        and use contours for the resampled stars.
        }
    \label{fig:star_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Galaxy Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sample}
    \end{center}
    \caption{
        Distributions of 5-band magnitudes (\textbf{upper panel}) and other properties 
        (\texttt{lower panel}; from left to right are the \ser{} index, axis ratio, 
        effective radius in arcsec, the $g-r$ and $r-z$ colors) of the synthetic 
        galaxies.
        Objects used for the \textbf{Tract} with good seeing (filled histograms) and 
        the one with bad seeing (dashed line) have almost identical distributions.
        }
    \label{fig:galaxy_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Generation of Synthetic Dataset}
    \label{sec:test}

    Here we demonstrate the application of \synpipe{} via testing the basic performance 
    of HSC photometry for both stars and galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Test Data}

    Given the design of \synpipe{} and the data volume of HSC DR1\footnote{${\sim}70$ 
    TB of images}, it is not realistic to test the whole dataset. 
    For this work, we select two \tracts{} in HSC Wide regions 
    that fit the following criteria:

    \begin{enumerate}

        \item Not at the edge of any Wide region, so the full \tract{} is 
            useful for test.

        \item Not bothered by extremely bright ($i<12$ mag), saturated star.

        \item Can represent good and bad seeing condition in the current $i$-band
            data of HSC survey.  
            For the same star or galaxy, different seeing conditions affect its 
            central S/N on the \texttt{coadd} images, hence it also impacts the 
            detection and deblending processes in \hscpipe{}. 
            Therefore, it is useful to show the photometric uncertainties under 
            different seeing conditions.

    \end{enumerate}

    \noindent These regions are: 
    \tract{}$=9699$ in the VVDS field (median FWHM$=0.449$\asec{} in $i$-band; 
    will be referred as \texttt{goodSeeing}) and
    \tract{}$=8764$ in the XMM-LSS field (median FWHM$=0.700$\asec{}; 
    \texttt{badSeeing}). 
    \tract{}$=9699$ also has better seeing in both $r$ and $z$-band; 
    their seeing conditions in $g$ and $y$ bands are very similar.    
    For both \tracts{}, the \texttt{coadd} image in each band includes data from 
    20--40 \visits{}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Input Models for Stars and Galaxies}
    \label{ssec:inputs}

    In this work, we use objects from the high-resolution \hst{}/ACS observations of 
    the COSMOS field to build input catalogs of realistic synthetic stars and galaxies 
    with appropriate completeness for the Wide survey and reasonable five-band colors.

    For stars, we choose to use the sample from \citet{Leauthaud2007}. 
    The authors select stars with clean photometry from the COSMOS images using not 
    only the \texttt{SExtractor} stellar classification index (\texttt{CLASS\_STAR}), 
    but also the peak surface brightness above background (\texttt{MU\_MAX}) and 
    half-light radius (\texttt{RHL}).
    The classification is reliable down to $I_{\mathrm{F814W}}{\sim}25.2$ mag.  
    We match this sample of stars with the HSC UltraDeep--COSMOS data that currently 
    reaches to about ${\sim}27.2$ mag in $i$-band. 
    After filtering the matched sample through basic quality cuts (see \ref{app:qc}),
    we gather 5-band HSC photometry for 14472 stars.
    To get a larger sample for the test, we resample the their 5-band magnitudes
    distributions using the \texttt{astroML} (\citealt{astroml}) alternative 
    implementation of the \texttt{extreme-deconvolution} algorithm developed by 
    \citet{Bovy2011}
    \footnote{\texttt{https://github.com/jobovy/extreme-deconvolution/}} to 
    generate 100000 synthetic stars. 
    In Fig \ref{fig:star_sample}, we show the magnitude and color distributions of 
    the real COSMOS stars are nicely recovered by the resampled stars in all 
    five-band magnitude and two color-color space.

    As for synthetic galaxies, we choose to use the single-\ser{} model of galaxies
    in the COSMOS field that have $I_{F814W} \leq 25.2$ mag. 
    These models are created via method developed by \citet{Lackner2012} on the 
    \textit{HST}/ACS images, and are included in \galsim{}. 
    Each model is described by a total of 5 parameters: magnitude, effective radius, 
    \ser{} index, axis ratio, and position angle. 
    Although the single-\ser{} model is not always the best choice, its flexibility 
    enables it to reasonably describe the flux distribution of most galaxies.
    Also its simplicity makes it easy to diagnose potential problems of the
    \cmodel{} photometry. 
    We exclude a tiny fraction of ill-behaved models that have very high \ser{} 
    index ($n_{\mathrm{Ser}} > 6.0$) or very low central surface brightness 
    ($\mu_{i} < 24.0$\sb). 
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data. 
    after removing the objects with problematic photometry, we obtain the a sample of 
    58210 synthetic galaxies with realistic 5-band HSC photometry. 
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are 
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec) galaxies. 
    This sample is appropriate to test the general behaviours of photometry in 
    \hscpipe{}, but is lack of relative bright galaxies ($i<20.5$ mag).

    We then create input catalogs via randomly distribute these synthetic stars and 
    galaxies into the regions covered by the good and bad seeing \tracts{}. 
    For stars (galaxies), we choose to use number density of 1000 (500)
    per CCD.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Running \synpipe{} Test}
    \label{ssec:running}

    We conduct these tests on the cluster (\texttt{Master}) at Kavli-IPMU using 
    the HSC DR1 data stored there. 
    The \texttt{addFakes.py} step takes ${\sim}1.5$ hours per \tract{} in each 
    band; same process takes longer for the test of galaxies (${\sim}3.0$ hours per
    \tract{} in each band) as the \galsim{} simulation is more time consuming. 
    The \texttt{stack.py} and \texttt{multiBand.py} together take ${\sim}3.5$ hours
    per \tract{}.

    We match the results with the input catalogs using a 2 pixels matching radius 
    to generate output catalogs for \forced{} and \unforced{} photometry 
    in each band. 
    We also match the input catalogs with the original HSC catalogs to identify the 
    ``ambiguously blended'' objects that are right on top of the flux peak of a real 
    object.  
    These objects are useful in the WL-related statistical tests (see Murata et al. 
    in prep. for more details), but will not help test photometry, hence they will 
    be excluded later

    The detailed log of the entire tests can be found here:
    \url{goo.gl/VINOVP}.
    User should be able to reproduce the results presented in this work using this 
    information. 
    It can also serve as a brief manual for running \synpipe{} test.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Results on Photometric Performance}
    \label{sec:result}

    Here we show the general performance of \hscpipe{} for stars (galaxies) 
    mainly through the \forced{} PSF (\cmodel{}) photometry. 
    Although \hscpipe{} provides more options (e.g. aperture and Kron photometry), 
    these two are the most appropriate and relevant ones for magnitude and colors of 
    star and galaxy as they consider the effects of PSF in different bands, 
    and are consistent across all bands in terms of position and shape.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relations between input magnitudes (\textbf{left}: $i$-band; \textbf{right}: 
        $g$-band) of synthetic stars and the $log(S/N)$ measured by \hscpipe{} PSF 
        photometry.  
        Hexagonal binned density plots and contours show the distributions for 
        stars in the \tract{} with good and bad seeing. 
        The highly blended stars are highlighted using scatter plots. 
        A grey dashed line marks $S/N = 5$.
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Accuracy of Forced PSF Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_psf_mag}
    \end{center}
    \caption{
        Accuracies of the \texttt{hscpipe} PSF photometry for synthetic stars measured
        by the difference between input and output \forced{} PSF magnitudes. 
        Plot [a, b, c, d, e] shows the results for [$g$, $r$, $i$, $z$, $y$]-band. 
        In each plot, left panel shows the relation between input magnitude and 
        magnitude difference. 
        Hexagonal binned density plot and contour are for the synthetic galaxies from 
        good and bad seeing \tracts{}. 
        Highly blended objects are highlighted using scattered plot. 
        The long-dashed line marks zero magnitude difference, while the pair of 
        dashed lines outline the running-median of PSF magnitude errors 
        (including the uncertainties in aperture correction). 
        Right panel shows the distributions of the magnitude differences for objects 
        in good (filled) and bad (solid line) seeing \tracts{}. 
        The dashed line is for the distribution of highly blended objects. 
        }
    \label{fig:psf_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        PSF photometry for synthetic stars in $g$- (\textbf{left} and $y$-band 
        (\textbf{right}). 
        Other formats are identical to the plots in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Accuracies of the PSF color
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_color}
    \end{center}
    \caption{
        Accuracies of the color measurements for synthetic stars via the differences
        between input and \forced{} PSF colors. 
        The \textbf{Upper panels} and \textbf{lower panels} are for $g-i$ and $i-y$ 
        colors separately.
        The \textbf{left} column shows the relation between input magnitude and 
        the color difference, and the \textbf{right} one uses the input colors as
        x--axis instead. 
        Other formats are identical to the plot in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_color}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Color-Color Distributions of the Input and Measurements for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_cdist}
    \end{center}
    \caption{
        Evaluations of the accuracy of color measurements for synthetic stars using 
        the color-color distributions. 
        \textbf{Left} uses $(g-r)$ v.s. $(r-i)$ colors, and the \textbf{right} one 
        uses $(i-z)$ and $(z-y)$ colors. 
        The filled contours and shaded histograms reflect the distributions for input 
        colors. 
        The empty contours and solid-line histograms show the distributions recovered 
        by \hscpipe{}. 
        }
    \label{fig:psf_cdist}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%     
\subsection{PSF photometry of Stars}
    \label{ssec:psf}

    In \hscpipe{}, the PSF magnitude is derived using matched-filter method that 
    depends on the best-fit PSF model and uses the Lanczos interpolation to shift 
    PSF model. 
    The error of PSF magnitude from \hscpipe{} only considers the per-pixel noise, 
    not the uncertainty of the centroid. 
    \hscpipe{} also estimates aperture correction for PSF photometry. 
    For more details please see Section 4.9.5 of Bosch\etal (in prep.).

    In our tests, we randomly inject ${\sim}100000$ stars to each \tract{} in all 
    five-bands. 
    Among these stars, ${\sim}3$--$4$\% of them locate within 2 pixels to the centroids
    of real objects. 
    We remove them from the sample to avoid confusion in the comparisons. 
    After this step, there are still ${\sim}1$--$3$\% of the remaining synthetic stars 
    have no match in the \synpipe{} outputs. 
    They could be either below the detection threshold or ambiguously blended with real
    objects. 
    The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
    mag fainter than the matched ones. 
    The fraction of unmatched stars is slightly higher in the bad seeing \tract{},
    which is consistent with the expectation of seeing-dependent limit magnitude,  
    although the difference in not very significant.     
    For the matched stars, we select the primary detections 
    (\texttt{detect.is-primary=True}) that have good photometric quality 
    (see Appendix \ref{app:qc} for details). 
    This gives us ${\sim}83000$ stars in each \tract{}.

    Next, we isolate the stars that are misidentified by \hscpipe{} as extended 
    objects using the \texttt{classification.extendedness} parameter in each band. 
    The fraction of misclassification is ${\sim}8$--$15$\%, and clearly depends on 
    seeing conditions. 
    For the same reason, $g$-band has the highest fraction of misclassification, 
    while $i$-band is recommended for selecting point sources.  
    The misclassified stars are on average ${\sim}1$ mag fainter than the others.
    We exclude them from the photometric comparison below, and will further discuss 
    the star--galaxy separation issue in Section \ref{ssec:sg}.

    In the following comparisons, we will always show the results from the 
    \textbf{good seeing} and \textbf{bad seeing} \tracts{} separately, because it 
    is important to understand the impact of seeing on photometric accuracies, 
    and to test whether \hscpipe{} can deliver unbiased photometry under different 
    seeing conditions.

    Besides seeing, ``blendedness'' ($b$; parameter that describes on what level 
    the flux distribution of one object is affected by others; please see 
    Appendix \ref{app:defineb} for precise definition, and see Murata\etal in prep. 
    for details) is another factor that influences the photometric accuracy. 
    We therefore separate the samples into relative isolated and highly blended 
    groups based on the $b$ values. 
    Assuming zero pixel noise, $b$ of object $A$ reflects the fraction of flux 
    within the \texttt{Footprint} of $A$ that comes from other objects.  
    Ideally, $b=0$ means perfect isolation, and $b{\sim}1$ means completely 
    confused with other object. 
    Here we use $b>0.05$ to define blended stars (${\sim}4$--$9$\%). 
    Fraction of highly blended stars slightly increases in the bad seeing \tract{}. 
    We highlight these stars separately as well, and will discuss more about the 
    impact of blendedness in Section \ref{ssec:blendedness}.

\subsubsection{Input magnitude and the $\mathrm{S}/\mathrm{N}$ of PSF photometry}

    \begin{itemize}

        \item In Fig \ref{fig:star_sn}, we show the relations between the input
            magnitude and the $\mathrm{S}/\mathrm{N}$ recovered by \hscpipe{}
            photometry in $g$- and $i$-band.

        \item \plan{Define $\mathrm{S}/\mathrm{N}$, quickly explain the purpose.}

        \item \plan{At fixed input magnitude, PSF $\mathrm{S}/\mathrm{N}$ depends 
            on seeing conditions. 
            For $i$-bands, the two \tracts{} have quite different seeing. 
            And bad seeing leads to clearly lower $\mathrm{S}/\mathrm{N}$.}

        \item \plan{Briefly discuss the detection limit of point sources. 
            Basically consistent with the reported values in \citet{HSCDR1}.}

        \item \plan{Blendedness can reduce the $\mathrm{S}/\mathrm{N}$ a little.}

    \end{itemize}

\subsubsection{Accuracy of PSF magnitude}

    \begin{itemize}

        \item In Fig \ref{fig:psf_mag}, we evaluate the accuracies of \hscpipe{} PSF 
            magnitudes in all five bands.

        \item \plan{The overall performance of PSF photometry is good. 
            In $i$-band, the median magnitude difference is around ${\sim}0.015$ mag
            (${\sim}1.5$\% accuracy) at $i_{\mathrm{Input}}{\sim}19.0$ mag. 
            At $i_{\mathrm{Input}}{\sim}24.0$ mag, the average accuracy of PSF 
            magnitude is around $3$\%. 
            Down to $i_{\mathrm{Input}}{\sim}25.2$ mag, the average accuracy is 
            around 7\%.
            This is consistent with the external check in \citet{HSCDR1}.}

        \item \plan{At fixed input magnitude, the accuracy does not strongly depends
            on filters. 
            The only exception is the $y$-band, which has slightly worse accuracy. 
            At $y_{\mathrm{Input}}{\sim}23.5$, the average accuracy is already close 
            to $10$\%.}
            \todo{Possible causes? background noise? Worse PSF model?}

       \item \plan{The \forced{} PSF magnitude provides unbiased estimates for
           isolated stars. 
           For $grzy$-bands, the median magnitude difference is ${\sim}0.002$ mag, 
           and can be ignored.}

       \item \plan{Using $i$-band as example, although worse seeing leads to 
           lower $\mathrm{S}/\mathrm{N}$, it does not strongly impact the 
           accuracy of the PSF magnitude. 
           The bad seeing \tract{} suggests almost identical accuracy comparing
           to the good seeing one.}

       \item \plan{However, the story for highly blended stars is quite different. 
           \hscpipe{} systematically under-estimate the total flux of highly blended
           stars by $0.05$--$0.10$ mag at fixed input magnitude.}

       \item \plan{We also test the \unforced{} PSF photometry in all five 
           bands. 
           The results suggest very similar accuracy.
           We highlight the differences between \forced{} and 
           \unforced{} PSF photometry in Fig \ref{fig:psf_diff}. 
           The overall difference is quite small, but it seems that \hscpipe{} 
           \unforced{} PSF photometry recovers slightly more flux at very 
           faint end.} 
           \todo{Move the figure to appendix?}

    \end{itemize}

\subsubsection{Accuracy of the PSF colors}

    \begin{itemize}

        \item \plan{Briefly explain why PSF colors are important; we focus on the 
            \forced{} PSF colors here}

        \item The results for accuracy of \forced{} PSF colors are shown in 
            Fig \ref{fig:psf_color} and Fig \ref{fig:psf_cdist}.

        \item \plan{In Fig \ref{fig:psf_color}, we investiage how \forced{}
            PSF colors depend on input magnitude and colors using both $(g-i)$ and 
            $(i-y)$ colors. 
            Results suggest \hscpipe{} \forced{} PSF photometry can provide 
            accurate and unbiased colors for stars in our tests.
            The uncertainties do not show any significant dependence on neither input
            magnitude nor color.}

        \item \plan{For $(g-i)$ color, average uncertainty around 
            $i_{\mathrm{Input}}{\sim}19.0$ mag is ${\sim}0.025$ mag. 
            The uncertainty increases to ${\sim}0.11$ mag at
            $i_{\mathrm{Input}}{\sim}25.0$ mag. 
            Results for $(i-y)$ color is similar.}

        \item \plan{In Fig \ref{fig:psf_cdist}, we use the results from the bad 
            seeing \tract{} to show that the \forced{} PSF photometry 
            can accurately recover the color-color sequences of stars down to 
            $i{\sim}25$ mag.}

        \item \plan{Accuracy of \forced{} PSF photometry apparently does not 
            strongly depend on seeing condition. 
            Impact from blendedness on PSF color is less serious than the magnitude, 
            although highly blended stars still show worse color accuracy and 
            systematically bluer (redder) $(g-i)$ ($i-y$) colors. 
            \unforced{} PSF photometry show similar but slightly worse 
            accuracy than the \forced{} ones. }

    \end{itemize}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Magnitude v.s S/N for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sn}
    \end{center}
    \caption{
        Relations between input magnitudes (\textbf{left}: $i$-band; \textbf{right}: 
        $g$-band) of synthetic galaxies and the $log(S/N)$ measured by \hscpipe{}
        \cmodel{} photometry.  
        Other formats are very similar to Fig \ref{fig:star_sn}
        }
    \label{fig:cmodel_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Accuracy of Forced cModel Magnitudes for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_galaxy_mag}
    \end{center}
    \caption{
        Accuracies of the \texttt{hscpipe} \cmodel{} photometry for synthetic 
        galaxies measured by the difference between input and output \forced{} 
        \cmodel{} magnitudes. 
        Plot [a, b, c, d, e] shows the results for [$g$, $r$, $i$, $z$, $y$]-band. 
        Other formats are identical to Fig \ref{fig:psf_mag}.
        }
    \label{fig:cmodel_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_cmodel_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        \cmodel{} photometry for synthetic galaxies in $g$- (\textbf{left} and 
        $y$-band (\textbf{right}). 
        Other formats are identical to the plots in Fig \ref{fig:psf_diff}.
        }
    \label{fig:cmodel_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Accuracies of the cModel color for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_color}
    \end{center}
    \caption{
        Accuracies of the color measurements for synthetic galaxies via the 
        differences between input and \forced{} \cmodel{} colors. 
        The \textbf{Upper panels} and \textbf{lower panels} are for $g-i$ and $i-y$ 
        colors separately.
        The \textbf{left} column shows the relation between input magnitude and 
        the color difference, and the \textbf{right} one uses the input colors as
        x--axis instead. 
        Other formats are identical to the plot in Fig \ref{fig:psf_color}.        
        }
    \label{fig:cmodel_color}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Color-Color Distributions of the Input and Measurements for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_cdist}
    \end{center}
    \caption{
        Evaluations of the accuracy of color measurements for synthetic galaxies 
        using the color-color distributions. 
        \textbf{Left} uses $(g-r)$ v.s. $(r-i)$ colors, and the \textbf{right} one 
        uses $(i-z)$ and $(z-y)$ colors. 
        The filled contours and shaded histograms reflect the distributions for input 
        colors. 
        The empty contours and solid-line histograms show the distributions recovered 
        by \hscpipe{} \cmodel{} photometry.
        }
    \label{fig:galaxy_cdist}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%     
\subsection{cModel photometry of Galaxies}
    \label{ssec:cmodel}

    \cmodel{} algorithm is the primary photometric method for galaxies in 
    \hscpipe{} that can be seen as an improved version of the SDSS \cmodel{} 
    photometry. 
    For more details, please see Section 4.9.8 of Bosch\etal (in prep.). 
    Basically, \cmodel{} tries to fit each object using a combination of de Vaucouleur
    and exponential profiles that are approximated by multiple Gaussian profiles. 
    Despite the limitation of the \cmodel{} method (e.g. sensitive to background 
    subtraction and deblender failure), it can deliver robust PSF-corrected fluxes and 
    colors for galaxies.

    \begin{itemize}

        \item \plan{${\sim}2$--$3$\% of the injected synthetic galaxies are 
            ambiguously blended with real objects.}

        \item \plan{Excluding the ambiguously blended ones, ${\sim}4$--$8$\% of the 
            remaining synthetic galaxies do not have any match.
            The unmatched galaxies are systematically fainter than the matched ones.
            And the fraction of unmatched synthetic galaxies is also slightly higher 
            in the bad seeing \tract{} than the one with good seeing.
            Similarly, $g$- and $i$-band have higher fraction of unmatched objects.
            }

       \item \plan{From the rest of the matched galaxies, we only keep the primary 
           detections (\texttt{detect.is-primary=True}) and the ones with clean 
           photometry (see Appendix \ref{app:qc} for details). 
           For each \tract{}, there are ${\sim}30000$ galaxies left.}

       \item \plan{A tiny fraction of synthetic galaxies is misclassified as 
           point sources by \hscpipe{}. 
           And the fraction of misclassifications seems to increase with seeing 
           (${\sim}1$\% in good seeing \tract{}; ${\sim}4$\% in bad seeing \tract{}). 
           We will further discuss in Section \ref{ssec:sg}.}

       \item \plan{We also use $b>0.05$ to define samples of highly blended galaxies.
           Since synthetic galaxies are more extended, their distribution of $b$
           skew more toward the higher value. 
           About ${\sim}7$--$9$\% of galaxies are highly blended by this standard.
           The $g$-band has significantly higher fraction of highly blended objects.}

    \end{itemize}

\subsubsection{Input magnitude and the $\mathrm{S}/\mathrm{N}$ of \cmodel{} photometry}

    \begin{itemize}

        \item In Fig \ref{fig:cmodel_sn}, we show the relations between the input
            magnitude and the $\mathrm{S}/\mathrm{N}$ recovered by \cmodel{}
            photometry in $g$- and $i$-band.

        \item \plan{Define $\mathrm{S}/\mathrm{N}$. 
                    Unlike point sources, at fixed total magnitude, the 
                    synthetic galaxies have a range of size and shape that 
                    lead to different average surface brightness. 
                    Comparing to stars, at the same input magnitude, synthetic
                    galaxies have lower average $\mathrm{S}/\mathrm{N}$ and larger 
                    scatter. }

        \item \plan{Seeing condition still impacts the $\mathrm{S}/\mathrm{N}$ of
                    galaxies}

        \item \plan{Blendedness can reduce the $\mathrm{S}/\mathrm{N}$ a little.}

        \item \plan{Briefly discuss the detection limit of point sources. 
            Basically consistent with the reported values in \citet{HSCDR1}.
            Although at $i{\sim}25.2$, the average $\mathrm{S}/\mathrm{N}$ is quite high,
            small fraction of galaxies will have very low $\mathrm{S}/\mathrm{N}$.}

    \end{itemize}

\subsubsection{Accuracy of the \cmodel{} magnitude}

    \begin{itemize}

        \item In Fig \ref{fig:cmodel_mag}, we evaluate the accuracies of the \forced{}
            \cmodel{} magnitudes in all five bands.

        \item \plan{Generally speaking, the \forced{} \cmodel{} provides good estimates 
            of magnitudes for galaxies between 20.0 and 25.0 magnitude in all filters and 
            different seeing conditions. }

        \item \plan{For $i$-band the absolute flux accuracy is around 7\% level at 
            $i{\sim}21.0$ mag, and it increases to ${\sim}10$\% at $i{\sim}25.0$ mag.
            The \forced{} \cmodel{} photometry slightly underestimate the total flux of 
            galaxies at $21.0 < i < 23.0$ mag. 
            The average magnitude difference is at 0.04 mag level. 
            For $i > 23.0$ mag, the \forced{} \cmodel{} photometry is unbiased.}

        \item \plan{The \forced{} \cmodel{} magnitude error from \hscpipe{} only considers 
            the per pixel noise, but not the systematic uncertainties involved in the 
            modeling fitting process.
            Therefore it is no surprise that the typical uncertainty from the \synpipe{} 
            test is larger than the magnitude error.}

        \item \plan{The accuracies of the \forced{} \cmodel{} magnitudes in other bands 
            are similar to the $i$-band.  
            The only exception is for galaxies with $y<24.0$ mag where \forced{} \cmodel{}
            seems to overestimate the flux slightly.}

        \item \plan{For galaxies with $i < 20.0$, the statistics are uncertain due to small 
            sample size}

        \item \plan{Unlike the synthetic stars, the highly blended synthetic galaxies only 
            have slightly biased photometry, where \forced{} \cmodel{} underestimates the 
            fluxes of these galaxies by ${\sim}0.05-0.10$ mag. 
            High blendedness does lead to larger scatter in photometric uncertainties.}

       \item \plan{We also test the \unforced{} \cmodel{} photometry in all five bands, 
           and the results suggest similar accuracies.
           We highlight the differences between \forced{} and 
           \unforced{} \cmodel{} photometry in Fig \ref{fig:cmodel_diff}.}

    \end{itemize}

\subsubsection{Accuracy of the \cmodel{} color}

    \begin{itemize}

        \item \plan{Briefly explain why \cmodel{} colors are important.}

        \item The results for accuracy of \forced{} \cmodel{} colors are shown in 
            Fig \ref{fig:cmodel_color} and Fig \ref{fig:cmodel_cdist}.

        \item \plan{\forced{} \cmodel{} photometry provides accurate and unbiased 
            colors for galaxies at $20.0 < i < 25.0$ mag at different seeing conditions.
            The accuracies of colors also do not show dependence on input colors. 
            }

        \item \plan{For $(g-i)$ color, average uncertainty around 
            $i_{\mathrm{Input}}{\sim}21.0$ mag is ${\sim}0.04$ mag. 
            The uncertainty increases to ${\sim}0.11$ mag at
            $i_{\mathrm{Input}}{\sim}25.0$ mag. 
            Results for $(i-y)$ color is similar.}

        \item \plan{High blendedness can increase the uncertainties of the color 
            and bias the color measurements. 
            The highly blended galaxies on average show slightly bluer $(g-i)$ colors 
            and redder $(i-y)$ colors comparing to the input values.}

        \item \plan{In Fig \ref{fig:cmodel_cdist}, we show that the \forced{} \cmodel{} 
            photometry can reasonably recover the color-color distributions of galaxies 
            down to $i{\sim}25$ mag.
            }

    \end{itemize}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Astrometric Calibration
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_astrometry}
    \end{center}
    \caption{
        Astrometric accuracy for synthetic star (\textbf{left}) and galaxy 
        (\textbf{right}) via the differences in R.A and Dec between the input
        coordinates and the ones measured on the coadd images. 
        Hexagonal bins (filled histogram) and contours (solid-line histogram) are for 
        the synthetic objects from \tracts{} with good and bad seeing conditions. 
        $\Delta\mathrm{R.A}=0$ and $\Delta\mathrm{Dec}=0$ are marked by dashed lines.
        }
    \label{fig:astrometry}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Star-Galaxy Separation
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_star_galaxy}
    \end{center}
    \caption{
        Illustrates the problems with current star--galaxy separation method in 
        \hscpipe{} which is based on the magnitude difference between the PSF and 
        \cmodel{} photometry. 
        \textbf{Left} plot is for the synthetic stars. 
        Hexagonal bins and filled histogram show distributions of correctly classified 
        synthetic stars in the good seeing \tract{}. 
        The scatter plot (dashed-line histogram) and contour (solid-line histogram) 
        show distributions of synthetic stars that are misclassified as extended 
        objects in \tracts{} with bad and good seeing conditions. 
        \textbf{Right} one is for synthetic galaxies. 
        Here the hexagonal bins (filled histogram) and contours (solid-line histogram)
        are for galaxies from good and bad seeing \tracts{}. 
        The scatter plot and dashed-line histogram are for synthetic galaxies that are 
        misclassified as point sources.
        Zero magnitude difference is highlighted using dashed lines on both plots.
        }
    \label{fig:sg}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Relations between Blendedness and Photomertic Errors
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_blendedness_err}
    \end{center}
    \caption{
        Shows the relationships between the $\log(\mathrm{Blendedness})$ parameter
        and the photometric accuracies for synthetic stars (\textbf{upper row}) and 
        galaxies (\textbf{lower row}). 
        \textbf{Left} column is for uncertainties of PSF or \cmodel{} magnitude, 
        and the \textbf{right} one is for uncertainties of $(g-i)$ color. 
        Hexagonal bins and contours show results for good and bad seeing 
        \tracts{}.
        $\log(\mathrm{Blendedness}) = -1.0$ is marked using vertical dashed line. 
        }
    \label{fig:blend}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: Compare with the behaviours of the Kron photometry for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_kron}
    \end{center}
    \caption{
        Basic performance of the Kron photometry for synthetic galaxies. 
        \textbf{Left} shows the accuracy in $i$-band. 
        magnitude. The format is identical to Fig \ref{fig:galaxy_mag}.
        \textbf{Right} shows the accuracy of the $(g-i)$ color using 
        Kron photometry. 
        Other formats are identical to Fig \ref{fig:cmodel_color}.
        }
    \label{fig:galaxy_kron}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Fig: HSC cModel has problem recovering the size and shape of galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_ba}
    \end{center}
    \caption{
        Shows the accuracy of the weighted axis ratio measurements for synthetic 
        galaxies using \cmodel{} photometry. 
        \textbf{Left} shows the relation between the input $(b/a)$ and the 
        uncertainties of $(b/a)$ estimates. 
        \textbf{Right} shows the relation between input magnitude of synthetic galaxy
        and the $(b/a)$ uncertainties. 
        Contours are for the default test with \texttt{usePixelWeights=False}, and the 
        hexagonal bins are used for same tests with \texttt{usePixelWeights=True}.
        }
    \label{fig:galaxy_ba}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

\section{Results on Other Performance}
    \label{sec:others}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Astrometric Calibration} 
    \label{ssec:astrometry}

    \begin{itemize}

        \item \plan{Explain what we are testing here; and why it is useful}

        \item In Fig \ref{fig:astrometry}, we briefly look into the accuracies 
            of the astrometric comparison for both synthetic stars and galaxies in 
            \tracts{} with both good and bad seeing.

        \item \plan{We see small astrometric offsets in both R.A and Dec directions
            that show small differences between the two \tracts{}.
            The average offset is at ${\sim}10$--20 mas, which is consistent with 
            the estimate in \citet{HSCDR1}}.

    \end{itemize}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Star-Galaxy Separation}
    \label{ssec:sg}

    \begin{itemize}

        \item \plan{Explain how star--galaxy separation is done in \hscpipe{}, and 
            why it is difficult at faint magnitudes.}

        \item In Fig \ref{fig:sg}, we highlight the misclassified stars and galaxies 
            on the relations between the input magnitude and 
            $i_{\mathrm{PSF}}-i_{\mathrm{cModel}}$.

        \item \plan{The fraction of misclassified stars show dependence on both 
            input magnitude and seeing.}

        \item \plan{Most of the misclassified galaxies have $i > 24.0$. 
            The fraction of misclassification and distribution of 
            $i_{\mathrm{PSF}}-i_{\mathrm{cModel}}$ for synthetic galaxies also depend on 
            seeings}.

    \end{itemize}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Blendedness and Photometric Errors}
    \label{ssec:blendedness}

    \plan{Briefly explain the definition of blendedness}

    \plan{Relationship between magnitude and blendedness}

    \plan{How blendedness affects the photometric uncertainties}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Comparing to Kron photometry}

    \plan{Uses $i$-band as example to briefly discuss Kron photometry}

    \plan{Although magnitude for galaxies is Ok, the color becomes worse}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Shape and Structural Parameters of Galaxies}
    \label{ssec:shape}

    \plan{Explain why the current \cmodel{}{} algorithm has problem with measuring 
          the size and shape of galaxies}

    \plan{Briefly showcase the problem, and how small change of \hscpipe{} 
          configuration can improve the situation}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Summary and Conclusion}
    \label{sec:summary}

    %% Short summary
    The on-going HSC survey just had its first public data release, and is still 
    producing increasingly larger amount of high-quality imaging data. 
    To help achieve the designed goals of the survey and facilitate scientific 
    outputs, we develop \synpipe{}, a flexible framework that is based on the 
    \hscpipe{} to exam the data reduction process and check the accuracy of all the
    pipeline outputs.
    \synpipe{} injects synthetic stars and galaxies to the desired locations on the
    single-\visit{} images with the help of the original astrometric and 
    photometric calibrations.
    These data are then processed by \hscpipe{} to generate the coadd images and 
    all the photometric measurements. 
    In this way, \synpipe{} can simulate the realistic data reduction process
    to an extent that many subtle systematical effects can be taken into 
    consideration, which is helpful for the outstanding scientific goals of the HSC 
    survey. 
    In this work, we choose two \tracts{} in the HSC/Wide layer with 
    representative seeing conditions (good/bad) to test the general behaviors of HSC 
    photometry for both stars and galaxies.

    From these tests, we find:

    \begin{enumerate}

        \item The \forced{} PSF photometry provides excellent magnitude and color 
            measurements for isolated stars at $18.0 < i < 25.0$. 
            The typical uncertainties of HSC forced PSF photometry for stars range 
            from around 0.01 mag at $i{\sim}18.0$ mag to 0.08 mag at $i{\sim}25.0$ 
            mag (1\%-7\% accuracy in the $i$-band).            
            The \forced{} PSF photometry can accurately recover the color--color 
            sequences for stars.

        \item The \forced{} \cmodel{} photometry is reliable for synthetic galaxies 
            with realistic distributions of structural parameters and colors at 
            $20.0 < i < 25.0$ mag. 
            The average accuracy of \forced{} \cmodel{} magnitude ranges from 
            ${\sim}7$\% at $i{\sim}20.0$ mag to ${\sim}10$\% at $i{\sim}25.0$. 
            The \forced{} \cmodel{} colors for galaxies are accurate and unbiased. 
            The accuracies of colors do not show dependence on input magnitude and 
            color.

        \item The \forced{} PSF and \cmodel{} photometry are robust against different 
            seeing conditions, despite that worse seeing leads to lower 
            $\mathrm{S}/\mathrm{N}$ at fixed input magnitude.

        \item High blendedness ($b>0.1$) impacts the accuracy of both PSF and \cmodel{}
            photometry, especially for stars.
            For highly blended stars, the \forced{} PSF photometry on average 
            overestimate the magnitudes of stars by 0.1–-0.2 mag. 
            For galaxies, high-blendedness typically adds an additional 0.05 mag 
            uncertainty in both magnitude and color estimates.

    \end{enumerate}

    Meanwhile, we also notice a few issues in the current \hscpipe{}:

    \begin{enumerate}

        \item The performance of current star--galaxy separation algorithm in \hscpipe{}
            is not perfect, especially at faint magnitude end and under poor seeing 
            condition.  
            The fraction of stars that are misclassified as extended objects is still 
            quite high ($>20$\% at $i> 22.5$ mag).

        \item The \cmodel{} setting right now (\texttt{usePixelWeights=True}) results 
            in very biased estimates of shape and size for galaxies, especially for the 
            relative faint ones. 
            The axis ratio of galaxy using \cmodel{} is highly overestiamted while 
            the effective radius is underestimated. 
            We show that, using \texttt{usePixelWeights=True} instead and significantly 
            improve this situation. 
            However, it may impact the accuracy of deblender, which we will test in the 
            future.

    \end{enumerate}

    The HSC collaboration is working on improving these issues in the future version 
    of \hscpipe{} and the following data release.  
    Meanwhile, we suggest the users of HSC data to keep these caveats in mind. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%     
    %% Other applications
    Although we mainly focus on the general behaviors of photometric measurements here, 
    \synpipe{} is quite flexible and has plenty of other applications for any HSC 
    dataset.  
    For instance, in a following paper, Murata\etal (in prep.) uses \synpipe{} to test 
    the impacts of extremely blended objects on the WL measurements. 
    Also, \citep{Niikura2017} applies \synpipe{} to the high-cadence HSC observations 
    of M31 to help the search of interesting microlensing events. 
    At this moment, it is also being used to help select high-$z$ Lyman-break galaxies 
    and Lyman-$\alpha$ emitters (Ono\etal in prep.; Kono\etal in prep); identify 
    low surface brightness dwarf galaxies (Greco\etal in prep); and test photometric 
    and WL measurements around nearby clusters (Chiu\etal in prep).

    %% Caveats
    In the end, it is worth brief discussing the limitations of the current \synpipe{}:

    \begin{enumerate}

        \item \citet{HSCDR1} points out that \hscpipe{} tends to over-subtract the
            background around bright objects. 
            Unfortunately, \synpipe{} now takes the original background subtraction on 
            the single-\visit{} image for granted, hence lacks the capability to 
            test or help improve this problem.

        \item Right now, \synpipe{} simply adopts the PSF measured by \hscpipe{} and 
            use it for model of point source and PSF convolution for galaxy. 
            Hence it can not be used to test how the uncertainty of the PSF modeling 
            affects the photometry and shape measurements. 
            This could be important for regions with exquisite seeing that makes the 
            PSF modeling difficult, and for accurate WL measurements.

        \item \synpipe{} now works in ``unit'' of \visit{} for single exposure 
            test, and \tract{} for test of \texttt{coadd} images. 
            In case of tests that focus on specific regions that are much smaller than
            the size of \visit{} or \tract{} (e.g. quality of photometry 
            around rich galaxy clusters), the usage of \synpipe{} often leads to 
            low efficiency.

    \end{enumerate}

    %% Future developments
    Right now, both \hscpipe{} and \synpipe{} still undergo active development, 
    and \synpipe{} will update accordingly to provide photometric benchmark for 
    each data release of HSC survey, and help \hscpipe{} improve. 
    At the same time, we are also working on improving the efficiency of \synpipe{}.
    Since subtle effects involved in the image stacking process are not the concerns 
    of many topics, we will update \synpipe{} so that it can directly inject 
    synthetic objects to the coadd images and speed up these tests.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

\begin{ack}
    \label{sec:ack}

    % HSC part
    The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of 
    Japan and Taiwan, and Princeton University.  
    The HSC instrumentation and software were developed by the National Astronomical 
    Observatory of Japan (NAOJ), the Kavli Institute for the Physics and Mathematics of
    the Universe (Kavli IPMU), the University of Tokyo, the High Energy Accelerator 
    Research Organization (KEK), the Academia Sinica Institute for Astronomy and 
    Astrophysics in Taiwan (ASIAA), and Princeton University.
    Funding was contributed by the FIRST program from Japanese Cabinet Office, the 
    Ministry of Education, Culture, Sports, Science and Technology (MEXT), the Japan 
    Society for the Promotion of Science (JSPS), Japan Science and Technology Agency 
    (JST), the Toray Science  Foundation, NAOJ, Kavli IPMU, KEK, ASIAA, and Princeton 
    University.

    % LSST software part
    This paper makes use of software developed for the Large Synoptic Survey Telescope. 
    We thank the LSST Project for making their code available as free software at 
    http://dm.lsstcorp.org.

    % Pan-STARRS1 part
    The Pan-STARRS1 Surveys (PS1) have been made possible through contributions of the 
    Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, 
    the Max-Planck Society and its participating institutes, the Max Planck Institute 
    for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, 
    Garching, The Johns Hopkins University, Durham University, the University of 
    Edinburgh, Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, 
    the Las Cumbres Observatory Global Telescope Network Incorporated, the National 
    Central University of Taiwan, the Space Telescope Science Institute, the National 
    Aeronautics and Space Administration under Grant No. 
    NNX08AR22G issued through the Planetary Science Division of the NASA Science 
    Mission Directorate, the National Science Foundation under Grant No. AST-1238877, 
    the University of Maryland, and Eotvos Lorand University (ELTE) and the Los Alamos 
    National Laboratory.

    % Software part
    This research made use of
    {\texttt{Astropy}}, 
        a community-developed core Python package for Astronomy (\citealt{Astropy}; 
        \url{http://www.astropy.org/}); 
    {\texttt{astroML}}, 
        a machine learning library for astrophysics (\citealt{astroml};
        \url{http://www.astroml.org/});
    {\texttt{SciPy}},
        an open source scientific tools for Python (\citealt{SciPy}; 
        \url{http://www.scipy.org/});
    {\texttt{NumPy}}, 
        a fundamental package for scientific computing with Python (\citealt{NumPy}; 
        \url{http://www.numpy.org/});
    {\texttt{Matplotlib}}, 
        a 2-D plotting library for Python (\citealt{Matplotlib}; 
        \url{http://matplotlib.org/});
    {\texttt{scikit-learn}},
        a machine-learning library in Python (\citealt{scikit-learn}; 
        \url{http://scikit_learn.org/}).

\end{ack}
%% ------------------------------------------------------------------------------------ %%

%%%%%%%%%%: Bibliographic Section %%%%%%%%%%

%\bibliography{synpipe}

%% ------------------------------------------------------------------------------------ %% 
\bibliographystyle{apj}

%% ------------------------------------------------------------------------------------ %% 
\begin{thebibliography}{}
    \label{sec:ref}
    \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

    %% A %%
    \bibitem[Abazajian et al.(2004)]{Abazajian2004} Abazajian, K., Adelman-McCarthy, 
             J.~K., Ag{\"u}eros, M.~A., et al.\ 2004, \aj, 128, 502

    \bibitem[Aihara et al.(2017)]{HSCDR1} Aihara, H., Armstrong, R., Bickerton, S., 
             et al.\ 2017, arXiv:1702.08449

    \bibitem[Antilogus et al.(2014)]{2014JInst...9C3048A} Antilogus, P., Astier, P., 
             Doherty, P., Guyonnet, A., \& Regnault, N.\ 2014, Journal of 
             Instrumentation, 9, C03048

    \bibitem[Astropy Collaboration et al.(2013)]{Astropy} Astropy Collaboration, 
             Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33

    \bibitem[Axelrod et al.(2010)]{Axelrod2010} Axelrod, T., Kantor, J., Lupton, 
             R.~H., \& Pierfederici, F.\ 2010, \procspie, 7740, 774015

    %% B %% 
    \bibitem[Bartelmann \& Schneider(2001)]{Bartelmann2001} Bartelmann, M., \& 
             Schneider, P.\ 2001, \physrep, 340, 291

    \bibitem[Ben{\'{\i}}tez(2000)]{Benitez2000} Ben{\'{\i}}tez, N.\ 2000, \apj, 
             536, 571

    \bibitem[Bertin(2011)]{Bertin2011} Bertin, E.\ 2011, Astronomical Data Analysis 
             Software and Systems XX, 442, 435

    \bibitem[Bertin(2013)]{Bertin2013} Bertin, E.\ 2013, Astrophysics Source 
             Code Library, ascl:1301.001

    \bibitem[Bolzonella et al.(2000)]{Bolzonella2000} Bolzonella, M., Miralles, J.-M., 
            \& Pell{\'o}, R.\ 2000, \aap, 363, 476

    \bibitem[Bovy Jo et al.(2011)]{Bovy2011} Bovy Jo, Hogg, D.~W., \& Roweis, 
             S.~T.\ 2011, Annals of Applied Statistics, 5,

    %% C %%
    \bibitem[Chang et al.(2015)]{Chang2015} Chang, C., Busha, M.~T., Wechsler, R.~H., 
             et al.\ 2015, \apj, 801, 73

    %% D %%
    \bibitem[Dressler et al.(2012)]{Dressler2012} Dressler, A., Spergel, D., Mountain,
             M., et al.\ 2012, arXiv:1210.7809

    %% G %% 
    \bibitem[Greisen \& Calabretta(2002a)]{WCS1} Greisen, E.~W., \& Calabretta, M.~R.\ 
             2002, \aap, 395, 1061

    \bibitem[Calabretta \& Greisen(2002b)]{WCS2} Calabretta, M.~R., \& Greisen, E.~W.\ 
             2002, \aap, 395, 1077

    %% H %% 
    \bibitem[Hunter (2007)]{Matplotlib} Hunter J. D., 2007, Computing In Science \& 
             Engineering, 9, 90

    %% I %% 
    \bibitem[Ilbert et al.(2009)]{Ilbert2009} Ilbert, O., Capak, P., Salvato, 
             M., et al.\ 2009, \apj, 690, 1236

    \bibitem[Ivezic et al.(2008)]{Ivezic2008} Ivezic, Z., Axelrod, T., Brandt, W.~N., 
             et al.\ 2008, Serbian Astronomical Journal, 176, 1

    %% J %%     
    \bibitem[Jones et al.(2001)]{SciPy} Jones E., Oliphant T., Peterson P., et al., 
             2001, SciPy: Open source scientific tools for Python, 
             \url{http://www.scipy.org/}

    \bibitem[Juri{\'c} et al.(2015)]{Juric2015} Juri{\'c}, M., Kantor, J., Lim, K., 
             et al.\ 2015, arXiv:1512.07914

    %% K %% 
    \bibitem[Kaiser \& Squires(1993)]{Kaiser1993} Kaiser, N., \& Squires, G.\ 
             1993, \apj, 404, 441

    %% L %%
    \bibitem[Lackner \& Gunn(2012)]{Lackner2012} Lackner, C.~N., \& Gunn, J.~E.\ 
             2012, \mnras, 421, 2277

    \bibitem[Laureijs et al.(2012)]{Laureijs2012} Laureijs, R., Gondoin, P., Duvet, L., 
             et al.\ 2012, \procspie, 8442, 84420T

    \bibitem[Leauthaud et al.(2007)]{Leauthaud2007} Leauthaud, A., Massey, R., 
             Kneib, J.-P., et al.\ 2007, \apjs, 172, 219

    \bibitem[Lupton et al.(2001)]{Lupton2001} Lupton, R., Gunn, J.~E., Ivezi{\'c}, 
             Z., Knapp, G.~R., \& Kent, S.\ 2001, Astronomical Data Analysis Software 
             and Systems X, 238, 269

    %% M %%
    \bibitem[Magnier et al.(2013)]{Magnier2013} Magnier, E.~A., Schlafly, E., 
             Finkbeiner, D., et al.\ 2013, \apjs, 205, 20

    \bibitem[Mandelbaum et al.(2014)]{Mandelbaum2014} Mandelbaum, R., Rowe, B., 
            Bosch, J., et al.\ 2014, \apjs, 212, 5

    \bibitem[Mandelbaum et al.(2015)]{Mandelbaum2015} Mandelbaum, R., Rowe, B., 
             Armstrong, R., et al.\ 2015, \mnras, 450, 2963

    \bibitem[Miyazaki et al.(2012)]{Miyazaki2012} Miyazaki, S., Komiyama, Y., Nakaya, 
             H., et al.\ 2012, \procspie, 8446, 84460Z

    %% N %%
    \bibitem[Niikura et al.(2017)]{Niikura2017} Niikura, H., Takada, M., Yasuda, N., 
             et al.\ 2017, arXiv:1701.02151

    %% P %% 
    \bibitem[Padmanabhan et al.(2008)]{Padmanabhan2008} Padmanabhan, N., Schlegel, 
             D.~J., Finkbeiner, D.~P., et al.\ 2008, \apj, 674, 1217-1233

    \bibitem[Pedregosa et al.(2011)]{scikit-learn} Pedregosa F., et al., 2011, 
             Journal of Machine Learning Research, 12, 2825

    %% R %%                       
    \bibitem[Rowe et al.(2015)]{Rowe2015} Rowe, B.~T.~P., Jarvis, M., Mandelbaum, R., 
             et al.\ 2015, Astronomy and Computing, 10, 121

    %% S %%         
    \bibitem[Schlafly et al.(2012)]{Schlafly2012} Schlafly, E.~F., Finkbeiner, D.~P.,
             Juri{\'c}, M., et al.\ 2012, \apj, 756, 158

    \bibitem[Scoville et al.(2007)]{Scoville2007} Scoville, N., Abraham, R.~G., 
             Aussel, H., et al.\ 2007, \apjs, 172, 38

    \bibitem[S{\'e}rsic(1963)]{Sersic1963} S{\'e}rsic, J.~L.\ 1963, Boletin de la 
             Asociacion Argentina de Astronomia La Plata Argentina, 6, 41

    \bibitem[Suchyta et al.(2016)]{Suchyta2016} Suchyta, E., Huff, E.~M., Aleksi{\'c}, 
             J., et al.\ 2016, \mnras, 457, 786

    \bibitem[Steidel et al.(1996)]{Steidel1996} Steidel, C.~C., Giavalisco, M., 
             Pettini, M., Dickinson, M., \& Adelberger, K.~L.\ 1996, \apjl, 462, L17

    %% T %%
    \bibitem[The Dark Energy Survey Collaboration(2005)]{DES2005} The Dark Energy 
             Survey Collaboration 2005, arXiv:astro-ph/0510346

    \bibitem[Tonry et al.(2012)]{Tonry2012} Tonry, J.~L., Stubbs, C.~W., Lykke, K.~R., 
             et al.\ 2012, \apj, 750, 99

    %% V %%                  
    \bibitem[VanderPlas et al.(2014)]{astroml} VanderPlas, J., Fouesneau, M., \& 
             Taylor, J.\ 2014, Astrophysics Source Code Library, ascl:1407.018

    %% W %% 
    \bibitem[Walt et al.(2011)]{NumPy} Walt S. v. d., Colbert S. C., Varoquaux G., 
             2011, Computing in Science and Engg., 13, 22

\end{thebibliography}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\appendix
%% ------------------------------------------------------------------------------------ %% 
\section{Quality Control of Synthetic Objects}
    \label{app:qc}

    To select synthetic stars and galaxies from the results to test the photometry, 
    we apply some basic quality control cuts here.

    The HSC survey defines the full--depth \& full--color regions (\texttt{FDFC}) to 
    make sure the data used for science reach the expected number of exposures in each 
    band 
    ($\mathtt{(\#gri\ \geq\ 4)\ and\ (\#yz\ \geq\ 6)\ and\ (Limiting\ imag> 25.6)}$). 
    And, the two \tracts{} we used in this work are almost entirely covered in 
    the \texttt{FDFC} region so we do not apply this cut.  
    We apply the following cuts to the matched synthetic objects"

    \begin{itemize}

        \item[ ] \texttt{ idetect\_is\_primary==True }
        \item[ ] \texttt{ ideblend\_skipped==False }
        \item[ ] \texttt{ iflags\_badcentroid==False }
        \item[ ] \texttt{ icentroid\_sdss\_flags==False }
        \item[ ] \texttt{ iflags\_pixel\_edge==False }
        \item[ ] \texttt{ iflags\_pixel\_interpolated\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_saturated\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_cr\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_bad==False }
        \item[ ] \texttt{ iflags\_pixel\_suspect\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_clipped\_any==False }
        \item[ ] \texttt{ iflags\_pixel\_bright\_object\_center ==False }

    \end{itemize}

    \todo{Details for stars and galaxies}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{$b$: The Blendedness Parameter}
    \label{app:defineb}

    To evaluate at what degree an object is blended with others, \hscpipe{} introduces
    the blendedness parameter: $b$ (\texttt{[grizy]blendedness\_abs\_flux}). 
    Please refer to Bosch\etal (in prep.) and Murata\etal (in prep.) for more details, 
    here we briefly explain the definition.

    For object ${\rm A}$: 
    \begin{eqnarray*}
        b({\rm A}) \equiv
        1-\frac{\int_{\mathbb{R}^2} {\rm d}x~ {\rm d}y\ \mathcal{N}_{\mathrm{A}}({\bf x}\ |\  \mu_{\mathrm{A}}, {\bf\Sigma}_{\mathrm{A}}) 
        F_{\mathrm{A}}({\bf x})}{\int_{\mathbb{R}^2} {\rm d}x~ {\rm d}y\ \mathcal{N}_{\mathrm{A}}({\bf x}\ |\  \mu_{\mathrm{A}},
        {\bf\Sigma}_{\rm{A}}) F_{\rm{total}}({\bf x})},      
        \label{eq:defineb}
    \end{eqnarray*}

    \noindent 
    where $\mathcal{N}_{\rm{A}}({ \bf x }\ |\  \mu_{\rm{A}}, {\bf\Sigma}_{\rm{A}})$ 
    is a 2-D Gaussian function at pixel position $\bf x$, $\mu_{\mathrm{A}}$ is the 
    estimated centroid of object ${\rm A}$, and covariance ${\bf \Sigma}_{\mathrm{A}}$ 
    is estimated based on the Gaussian-weighted adaptive second moments of object 
    ${\rm A}$ (no PSF correction). 
    $F_{\mathrm{total}}({\bf x })$ and $F_{\rm{A}}({\bf x })$ are pixels values of 
    ${\rm A}$ before and after \hscpipe{} deblending process.

    By definition, the parameter is bound from 0 to 1. 
    When the deblending process is carried out correctly, the $b({\rm A})$ parameter 
    reflects the fraction of fluxes that comes from other objects in the 
    region of ${\rm A}$.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%%%%%%%%%%%: Possible Tables %%%%%%%%%%%%%
%\input{table1.tex}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\label{lastpage}
\end{document}
%% ------------------------------------------------------------------------------------ %%