### 2017-04-19 draft/ms_susan.tex
New edits from Susan

% Version 0; Only rough outline by SH
% Version 1; First draft by SH
% Version 2; Main text is finished; with edits from Susan Foster

%% ------------------------------------------------------------------------------------ %%
\documentclass[useamsfonts]{pasj01}
\twocolumn

%% ------------------------------------------------------------------------------------ %%
% Packages
\usepackage{url}
\usepackage{natbib}
\usepackage{latexsym}
\usepackage{morefloats}
\usepackage{xspace}
\usepackage[usenames, dvipsnames]{xcolor}

% Package Settings
%\usepackage{hyperref}
%\hypersetup{colorlinks=true,
%            citecolor=cyan,
%            linkcolor=cyan,
%            filecolor=magenta,
%            urlcolor=cyan}
%\urlstyle{same}
%% ------------------------------------------------------------------------------------ %%

% Figure extension
\DeclareGraphicsExtensions{.pdf,.png,.eps}

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %%
% Song Huang's definition
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\amin{$^\prime$}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over \sim \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over \sim \;$}}
\def\micron{{$\mu$m}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$}
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}

\def\logms{{$\log (M_{\star}/M_{\odot})$}~}
\def\mstar{{$M_{\star}$}~}
\def\logmh{{$\log (M_{\mathrm{halo}}/M_{\odot})$}~}
\def\m2l{{$M_{\star}/L_{\star}$}~}

\def\redm{\texttt{redMaPPer}}
\def\hscpipe{\texttt{hscPipe}}
\def\synpipe{\texttt{SynPipe}}
\def\cmodel{\texttt{cModel}}
\def\forced{\texttt{forced}}
\def\coadd{\texttt{coadd}}
\def\unforced{\texttt{unforced}}
\def\tract{\texttt{Tract}}
\def\visit{\texttt{Visit}}
\def\tracts{\texttt{Tracts}}
\def\visits{\texttt{Visits}}
\def\galsim{\texttt{G}{\scriptsize \texttt{AL}}\texttt{S}{\scriptsize \texttt{IM}}}
\def\hst{{\textit{HST}}}
\def\s2n{{$\mathrm{S}/\mathrm{N}$}}

%% ------------------------------------------------------------------------------------ %%

% Commenting:
%% ------------------------------------------------------------------------------------ %%
\newcommand{\plan}[1]{\textcolor{blue} {\textbf{#1}}}
\newcommand{\todo}[1]{\textcolor{red} {\textbf{#1}}}
\newcommand{\song}[1]{\textcolor{cyan} {\textbf{#1}}}
\newcommand{\addref}{{\textcolor{red}{\textbf{REF}}}}
\newcommand{\term}[1]{\textbf{\texttt{#1}}}
\newcommand{\susan}[1]{\textcolor{magenta} {\textbf{#1}}}

\newcommand{\plus}{\raisebox{.4\height}{\scalebox{.6}{+}}}
\newcommand{\minus}{\raisebox{.4\height}{\scalebox{.8}{-}}}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %%

\begin{document}

\SetRunningHead{Huang et al.}{HSC-SSP SynPipe}

\Received{$\langle$April 2017$\rangle$}
\Accepted{$\langle$2017$\rangle$}
\Published{$\langle$2017$\rangle$}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %%
\title{HSC Synthetic Pipeline I: Basic Design and Characterization of HSC
       Photometric Performance}

\author{Song Huang \altaffilmark{1,2}
        Alexie Leauthaud \altaffilmark{1,2},
        Ryoma Murata \altaffilmark{1, 4},
        Jim Bosch \altaffilmark{3},
        Paul Price \altaffilmark{3},
        Robert Lupton \altaffilmark{3},
        Rachel Mandelbaum \altaffilmark{5},
        Claire Lackner \altaffilmark{1},
        Steven Bickerton \altaffilmark{1},
        Satoshi Miyazaki \altaffilmark{6}
        HSC Collaboration \altaffilmark{7}}
%% ------------------------------------------------------------------------------------ %%

\altaffiltext{1}{Kavli Institute for the Physics and Mathematics of the
    Universe, The University of Tokyo Institutes for Advanced Study,
    the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan}
\altaffiltext{2}{Department of Astronomy and Astrophysics, University of California,
    Santa Cruz, 1156 High Street, Santa Cruz, CA 95064 USA}
\altaffiltext{3}{Department of Astrophysical Sciences, Princeton University,
    4 Ivy Lane, Princeton, NJ 08544}
\altaffiltext{4}{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
\altaffiltext{5}{McWilliams Center for Cosmology, Department of Physics,
    Carnegie Mellon University, Pittsburgh, PA 15213, USA}
\altaffiltext{6}{SOKENDAI (The Graduate University for Advanced Studies), Mitaka,
    Tokyo, 181--8588, Japan}
\altaffiltext{7}{Other affiliations}

\email{song.huang@ipmu.jp}
%% ------------------------------------------------------------------------------------ %%

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %%
\KeyWords{Surveys,
          Methods: observational,
          Techniques: photometric}

\maketitle

%% ------------------------------------------------------------------------------------ %%
\begin{abstract}

    The Subaru Strategic Program (SSP) is an ambitious wide-area, multi-band survey 
    using the Hyper Suprime-Cam (HSC) on the Subaru telescope.
    As SSP reaches to detection limit of $i>26.0$ mag in the Wide field, it is 
    challenging to achieve accurate, unbiased, and consistent photometry across all 
    five bands.
    To improve data processing and achieve accurate photometric measurements, we have 
    developed \synpipe{}, a \texttt{Python}-based, flexible framework that interfaces 
    with the HSC data reduction pipeline (\hscpipe{}) to process data from the first 
    public HSC--SSP data release. 
    In the first paper of this series, we explain the design and implementation of 
    \synpipe{}, and use \coadd{} images in two Wide layer \tracts{} with representative 
    seeing conditions to examine the photometric performance of \hscpipe{}.
    We find that uncertainties of HSC \forced{} PSF photometry for stars typically
    range from ${\sim}0.01$ mag at $i{\sim}18.0$ mag to ${\sim}0.06$ at $i{\sim}25.5$
    mag (1\%-7\% accuracy in the $i$-band). 
    For synthetic galaxies based on a single-\ser{} model, uncertainties of the
    forced \cmodel{} photometry typically range from 0.14 mag at $i{\sim}20.0$ mag 
    to 0.20 mag at $i{\sim} 25.2$ mag.
    Over the color and magnitude ranges used in these \synpipe{} tests, we find that 
    both \forced{} PSF and \cmodel{} photometry provide unbiased color estimates 
    that are robust in different seeing conditions.
    We also notice a few issues in the current \hscpipe{} that are worth keeping
    in mind when using HSC data: 
    1) The degree to which an object is blended with others impacts the photometric 
    accuracy, especially for point sources.
    Highly blended objects tend to have larger photometric uncertainty, systematically
    underestimated flux, and slightly biased color.
    2) A relatively high fraction of faint stars (${\sim}16$\% at $i>22.5$ mag) can be
    misclassified as extended objects.
    3) The shape and size estimates for galaxies using \cmodel{} suffer from strong
    biases.

\end{abstract}
%% ------------------------------------------------------------------------------------ %%

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %%
\section{Introduction}
    \label{sec:intro}

    %% Background
    Wide-field, multi-band imaging surveys have stepped onto the central stage of
    modern astrophysics and cosmology over the past decade.
    They also have a bright future on many current and planned ambitious projects, 
    such as the Large Synoptic Survey Telescope
    (LSST)\footnote{\url{https://www.lsst.org/}}, the Wide-Field Infrared Survey
    Telescope
    (WFIRST; \citealt{Dressler2012})\footnote{\url{https://wfirst.gsfc.nasa.gov/}},
    and the \textit{Euclid} project
    (\citealt{Laureijs2012})\footnote{\url{http://sci.esa.int/euclid/}}.
    Among many ongoing efforts, the Subaru Strategic Program
    (SSP; \citealt{HSCDR1})\footnote{\url{http://hsc.mtk.nao.ac.jp/ssp/}}, which
    uses the Hyper Suprime-Cam (HSC; \citealt{Miyazaki2012}) on the prime focus of
    the Subaru telescope, is the most efficient in terms of 
    etendue\footnote{The collecting area multiplied by the field of view.}.
    Together, these surveys will greatly help us test cosmological models, trace the
    evolution of galaxies, study the stellar structure of our Milky Way, and discover
    a large number of interesting transient objects.

    %% Challenges
    Before we can tackle any outstanding scientific questions, we must first learn
    how to handle the huge amount of data\footnote{Right now, SSP has accumulated 
    ${\sim}300$ TB of data products.} generated by these projects as well as achieve
    accurate photometric measurements of different types of objects.
    Data handling becomes increasingly challenging in the age of modern imaging
    surveys, in part due to variations and effects related to the cameras' 
    charge-coupled devices (CCDs).
    The cameras are often made up of multiple CCDs, each with slightly different 
    characteristics, and have large fields-of-view (FoV) over not perfectly flat 
    focus planes.
    During observation, the seeing and background conditions always show spatial
    and temporal variations across the FoV.
    The full-depletion, thick CCDs selected for the cameras enable long exposure 
    times and excellent sensitivity in the red-end, but they often suffer from the 
    so-called ``brighter-fatter'' effect (\citealt{Antilogus2014}), which means that 
    brighter stars appear larger than they are.
    These variations and effects make accurate astrometric and photometric calibration,
    background subtraction, and point spread function (PSF) modeling intrinsically
    difficult (e.g., \citealt{Schlafly2012}).
    The multi-epoch, multi-band nature of the surveys also complicate data reduction 
    because the scientist must pay special attention to image stacking, multi-band 
    object detection, and measurement processes.
    Furthermore, as ongoing surveys reach to increasingly deeper detection limit 
    (e.g., SSP Wide reaches to $5\ \sigma$ detection limit of point source of 
    ${\sim}26.4$ mag), the number of objects in a unit area, especially faint objects,
    becomes much higher. 
    In addition, it is much more common to observe objects that overlap each other at
    different levels (a phenomenon often referred to as ``blended'').
    These modern imaging survey challenges call for photometric measurement methods 
    that are more powerful and precise than those used in previous, shallower surveys.

%% ------------------------------------------------------------------------------------ %%
    %% Reason
    The challenges are not merely technical details; their resolution is crucial
    to achieving the surveys' key scientific goals.
    For instance, weak gravitational lensing (WL; \citealt{Kaiser1993, Bartelmann2001})
    is a powerful tool for measuring the large-scale distribution of dark matter and
    following its evolution.
    However, WL measurements critically depend on the ability to model PSF and measure 
    the shape of background galaxies with high precision 
    (e.g., \citealt{Mandelbaum2015}).
    The photometric redshift measurement (e.g., \citealt{Benitez2000, Bolzonella2000,
    Ilbert2009}) is fundamentally important for studying the evolution of galaxies,
    and the ``drop-out'' method is very useful in selecting high-redshift galaxies
    (e.g., \citealt{Steidel1996}).
    However, both of those methods are based on accurate multi-band photometric 
    measurements.
    There are many other examples, but these are sufficient to demonstrate that we 
    need to develop new frameworks to oversee the data reduction process of modern 
    imaging surveys.
    Several tools with similar goals have been developed (e.g., \citealt{Chang2015,
    Suchyta2016}) for the Dark Energy Survey (DES; \citealt{DES2005}). 
    As for the HSC--SSP (hereafter referred to as the HSC survey), the higher quality 
    images and more complex \hscpipe{} also require more carefully designed tool to 
    test them.

    %% Organization
    Here, we present \synpipe{}, a \texttt{Python}-based software that is a flexible 
    framework with many applications that can help future developments of the HSC data 
    reduction pipeline (\hscpipe{}).
    Below, we briefly introduce the HSC survey and the current status of data 
    reduction (Section \ref{sec:ssp}), explain \synpipe{} design and implementation 
    (Section \ref{sec:synpipe}), and demonstrate \synpipe{} usage via straightforward 
    tests (Section \ref{sec:test}) that focus on the general photometric quality for 
    both synthetic stars and galaxies (Section \ref{sec:result}).
    We then summarize the work and discuss future developments in 
    Section \ref{sec:summary}.

    The \synpipe{} code, along with basic documents and a few examples, is available
    on GitHub at: \texttt{https://github.com/dr-guangtou/synpipe}.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Flowchart
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_flowchart_new}
    \end{center}
    \caption{
        Illustration of \synpipe{} basic procedures.
        Gray boxes indicate the required inputs at different stages.
        The blue box identifies the \texttt{addFakes.py} step, when \synpipe{} injects 
        simulated images of synthetic objects onto a single-frame image. 
        We show an example CCD image with and without synthetic objects on it
        (highlighted by green circles).
        Red boxes depict the subsequent image coadding and multiband measurement steps
        using \texttt{stack.py} and \texttt{multiBand.py} from \hscpipe{}.
        We show a small piece of \coadd{} image with synthetic objects on it.
        The figure on the right side of the middle row shows the spatial relation
        between \tract{} (outlined by the black dashed lines; each small region
        is a \texttt{Patch}) and all the \visits{} that contribute to it. 
        (Each \visit{} has a  unique color; each small region is the size of a CCD.)
        }
    \label{fig:flowchart}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\section{Hyper-Suprime Cam Subaru Strategic Program (HSC Survey)}
    \label{sec:ssp}

    %% Basic
    Taking advantage of the new prime focus camera on the 8.2--m Subaru telescope,
    the ambitious HSC survey consists of three layers: Wide, Deep, and UltraDeep.
    The Wide layer will map a total of ${\sim}1400$ deg$^2$ of sky in five broad bands
    ($grizy$; Kawanomoto\etal in prep.).
    The Deep (four separated fields; ${\sim}27$ deg$^2$) and UltraDeep (two separated
    fields; ${\sim}3.5$ deg$^2$) layers will use a few additional narrow-band filters
    and employ a slightly different surveying strategy.
    Strauss\etal (in prep.) describes the HSC survey in more detail, and identifies the 
    HSC collaborators.

    The HSC camera (\citealt{Miyazaki2012}) is made up of 124 full-depletion thick 
    CCDs: 112 for science and another 12 for guiding and focusing.
    The camera has a circular FoV with a 1.5 deg diameter.
    Each CCD contains 2048$\times$4096 pixels distributed in four amplifiers.
    The pixel size is equal to 0.168\asec{} on the sky.

    In February 2017, the HSC collaboration released the first 1.7 years of data to 
    the public (DR1; 
    \citealt{HSCDR1})\footnote{\texttt{https://hsc-release.mtk.nao.ac.jp }}.
    For our work, we focused on data from the Wide layer as they relate to key 
    scientific goals of the HSC survey, which include WL cosmology, galaxy evolution 
    since $z{\sim}1$, and studies of galaxy clusters.
    The current Wide layer includes ${\sim}108$ deg$^2$ of sky isolated in
   six different regions (XMM--LSS, GAMA09H, WIDE12H, GAMA15H, HECTOMAP, and VVDS)
    \footnote{A small AEGIS field is also observed for calibration purposes.}.
    Except for the HECTOMAP field, the regions are all close to the equator. 
    In the Wide layer, the $g$ and $r$ bands have 10 mins of exposure broken into four
    dithers\footnote{Large, ${\sim}1/3$ of the FoV dither pattern is used to ensure
    uniform coverage.}; for the remaining three bands, 20 mins of exposure time and six
    dithers are used\footnote{Several 30 s exposures are taken to improve the 
    dynamic range.}.
    The survey prioritizes the observations so that the $i$-band can have excellent
    seeing conditions to help make accurate WL shape measurements.
    The current Wide layer has impressive median seeing with FWHM${\sim}0.6$\asec{},
    and has a $i{\sim} 26.4$ mag $5\sigma$ point source detection limit. 
    \citealt{HSCDR1} provides details on the first data release and data status.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\section{The HSC Data Reduction Pipeline}
    \label{ssec:hscpipe}

    The HSC survey data are currently reduced by \hscpipe{} \texttt{v4.0.5}, a
    version of the LSST reduction pipeline (\citealt{Ivezic2008, Axelrod2010,
    Juric2015}), which is described in detail in Bosch\etal (in prep.).
    Since \synpipe{} is intrinsically connected to the complex data reduction
    processes in \hscpipe{}, we briefly introduce the main steps below, along with
    a few key HSC/LSST terms.

    \begin{enumerate}

        \item \textbf{Single-\texttt{Visit~}Processing}:
            Each individual exposure is called a \term{Visit} and is assigned 
            an even integer.
            After initial data screening (in background, seeing, and transparency
            Furusawa\etal in prep.), \hscpipe{} detrends these data (subtraction
            of overscan, bias, and dark; flatfielding), generates variance and mask
            images, subtracts background, and provides initial astrometric and
            photometric calibrations.
            \textbf{\synpipe{} will use the spatially varying PSF models, photometric 
            zeropoints, and the World Coordinate System (WCS; \citealt{WCS1, WCS2}) 
            corrected for optical distortion.}.

        \item \textbf{Multi-\visit{} Processing}:
            After single-\visit{} processing, \hscpipe{} warps and mosaics the 
            reduced CCD images into much deeper \coadd{} images while improving
            the astrometric and photometric calibrations via processes similar to 
            the \"{u}ber-calibration in the Sloan Digital Sky Survey (SDSS) 
            \citep{Padmanabhan2008}.
            \textbf{\synpipe{} will use these improved calibrations.}
            \hscpipe{} organizes these \coadd{} images into equiareal rectangular 
            regions, or \tracts{}, which are predefined as iso-latitude tessellations.
            One \tract{} covers approximately $1.7\times 1.7$ deg$^2$ of sky, and 
            adjacent \tracts{} overlap each other by ${\sim}1$\amin{}.
            Each \tract{} is further divided into $9\times9$ \texttt{Patches}.
            A \texttt{Patch} is a $4200\times4200$ pixels rectangular region, and
            adjacent \texttt{Patches} overlap each other by 100 pixels.
            \textbf{The \tract{} and \texttt{Patch} are basic units to deal
            with HSC \coadd{} images and \synpipe{} tests.}

        \item \textbf{Multi-band Measurements}:
            To achieve consistent photometry across all filters, \hscpipe{} first
            detects and deblends objects on \coadd{} images in each band
            independently (the \unforced{} measurements).
            The collection of above-threshold pixels for each object is referred to as
            a \texttt{footprint}.
            \hscpipe{} merges the \texttt{footprints} and flux peaks in the different
            bands and then selects a reference band for each object based on its $S/N$ 
            in each band (usually it is the $i$-band).
            With the centroids, shape, and other non-amplitude parameters fixed at the
            reference values, \hscpipe{} performs \forced{} photometry, which is
            optimized to generate accurate colors.

        \item \textbf{HSC Photometry}:
            The multi-band catalogs generated by \hscpipe{} contain a variety of 
            photometric information (see \citealt{HSCDR1} for details).
            \textbf{Here, we focus on the \forced{} \texttt{psf} and \cmodel{}
            photometry.}
            \texttt{psf} photometry should provide the most appropriate magnitudes and
            colors for point sources.
            Improved over the SDSS version (\citealt{Lupton2001, Abazajian2004}),
            the HSC \cmodel{} algorithm (described in Bosch\etal, in prep.) fits 
            each object's flux distribution using a combination of the de Vaucouleurs
            profile and an exponential component that takes the PSF convolution 
            into account.
            \cmodel{} is designed to yield accurate colors for galaxies.
            In addition, the star--galaxy classification task is performed based on 
            the difference between the \texttt{psf} and \cmodel{}{} magnitudes.

    \end{enumerate}

    DR1 data quality has been vetted in \citet{HSCDR1} via cross-checking
    with the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 1
    imaging survey (\citealt{Schlafly2012, Tonry2012, Magnier2013}) to examine the
    behaviors of the stellar sequence and the shape of the PSF models.
    Generally speaking, the PSF photometry is accurate to 1--2\% level, and the
    astrometry is accurate to ${\sim}10$ and 40 max internally and externally.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\section{SynPipe Overview}
    \label{sec:synpipe}

%% ------------------------------------------------------------------------------------ %%
\subsection{ \synpipe{} Basic Design}
    \label{ssec:design}

    \hscpipe{} is a complex data reduction pipeline that still faces many challenges
    and is under active development. 
    Unlike most photometric pipelines on the market, it involves high-level reduction
    processes that produce not only deep \coadd{} images but also a series of
    science-ready catalogs.
    To faithfully test \hscpipe{} and the related data products, we have designed an 
    approach that ensures \synpipe{} satisfies a few basic standards:

    \begin{itemize}

        \item \textbf{Realistic Images}:
            Instead of using the fully generative approach to simulate full HSC-like
            images from the ground up (e.g., \citealt{Chang2015}), \synpipe{} injects
            synthetic objects into the observed image so that all the realistic
            features of the real data (e.g., blended objects; proximity to bright
            objects, bleeding trails, other optical artifacts can be included in the 
            test.
            The impact of these features can be important to any topic (e.g., WL 
            analysis; stellar mass function of galaxies) that cares about the selection 
            function and completeness of samples.
            In \synpipe{}, point source is simulated using the HSC PSF model,
            and a realistic galaxy model is created by the modular galaxy image
            simulation toolkit, \galsim{} (\citealt{Rowe2015})
            \footnote{\texttt{https://github.com/GalSim-developers/GalSim}}.

        \item \textbf{Authentic Data Processing}:
            To follow the data reduction process as realistically as possible, we choose to
            start from the single-\visit{} images instead of directly putting synthetic 
            objects on the final \coadd{} images (e.g., \citealt{Suchyta2016}).
            Through this approach, subtle effects like seeing differences among
            \visits{} and small errors in astrometric calibration can be taken
            into account.
            Every synthetic object will experience all the steps we would use for any 
            real object: detection, deblending, stacking, and measurement.
            This can be important for challenging tasks like the WL measurements
            or the detection of high-$z$ galaxies.
            The only downside of this choice is that it will certainly slow down the
            test for \coadd{} images because in this approach we need to inject 
            synthetic objects into all the \visits{} that contribute to a \tract{}.

        \item \textbf{Flexible Capabilities}:
            We must make \synpipe{} flexible enough to be useful for HSC users with 
            various scientific goals.
            We provide default catalogs to generate samples of synthetic stars
            and galaxies with realistic magnitude and color distributions along with
            tools to help the user work on HSC DR1 data; however, the user is allowed 
            to be creative in terms of providing input catalogs and designing the tests.
            \synpipe{} is already being used in a wide range of
            scientific applications (see Section \ref{sec:summary} for details).

    \end{itemize}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{\synpipe{} Implementation }
    \label{ssec:flowchart}

    In this section, we describe the \synpipe{} test implementation depicted in 
    Fig \ref{fig:flowchart}.

%% ------------------------------------------------------------------------------------ %%
\subsubsection{Preparation}
    \label{sssec:prep}

    To start a \synpipe{} test, the user needs to provide information about the data 
    and an input catalog, as discussed below:

    \begin{itemize}

        \item \textbf{Information about the data}
            includes the location of the HSC images and a list of \visits{} for 
            the test.
            To test the \coadd{} images, \synpipe{} can help the user identify
            all the \visits{} that contribute to one \tract{} in any given
            band.
            \synpipe{} also provides tools to create an optional bad-pixel mask 
            (e.g., bright object, bleeding trail) for a specific \tract{} so that 
            the user can avoid putting synthetic objects on problematic regions.

        \item \textbf{Input catalog} in \texttt{FITS} (Flexible Image Transport System)
            format contains the necessary information, including the coordinate, 
            magnitude, and other model parameters.  
            For any given sample of synthetic objects, \synpipe{} can help the user 
            randomly distribute the objects in a selected rectangular region with the
            desired density.
            If necessary, the user can reduce overlaps of synthetic objects by setting 
            a minimum distance between any two of them. 
            \synpipe{} can also help put synthetic objects on an evenly separated
            coordinate grid, which is very useful for certain tests
            (e.g., Murata\etal in prep.).

    \end{itemize}

    For the synthetic object, \synpipe{} provides the following choices:

    \begin{itemize}

        \item \textbf{Point source}:
            \synpipe{} simply uses the PSF model from \hscpipe{} as the model for
            a point source such as a star or a quasi-stellar object (QSO).
            \hscpipe{} uses a special version of \texttt{PSFEx}
            (\citealt{Bertin2011, Bertin2013}) to characterize PSF as a function
            of the point source's position so that a PSF model can be reconstructed 
            for any location on the image.
            The user only needs to provide the integrated magnitude to simulate
            point sources.

        \item \textbf{Single- or Double-\ser{} model for galaxy}:
            \texttt{SynPipe} uses \galsim{} \texttt{v1.4} (\citealt{Rowe2015}) to
            simulate a galaxy.
            \galsim{} is an image simulation tool that has been well tested during
            the GRavitational lEnsing Accuracy Testing 3 (GREAT3) challenge 
            (\citealt{Mandelbaum2014}).
            Currently, \synpipe{} allows a synthetic galaxy to be modeled by single
            or double components that are described by the \citet{Sersic1963}
            profile\footnote{More flexible model choices will be added later.}.
            The \ser{} profile is flexible enough to describe the overall flux
            distributions of galaxies near and far, both early-type or late-type.
            A double-\ser{} model can simulate a galaxy with even more realistic
            structural details (e.g., bulge$+$disk).
            For each \ser{} component, the user needs to provide the magnitude,
            effective radius ($R_{\mathrm{e}}$; in unit of arcsec), \ser{} index
            ($n_{\mathrm{Ser}}$), axis ratio ($\mathrm{b}/\mathrm{a}$), and position
            angle (PA).
            External shear can also be applied to a synthetic galaxy ($g_1$ and 
            $g_2$; see \citealt{Rowe2015} for more details).

        \item \textbf{Models or real images of a galaxy from the GREAT3 challenge}:
            In addition to providing detailed model information for each galaxy, 
            \synpipe{} also allows users to choose from parametric models or real
            high-resolution galaxy images used in the GREAT3 WL challenge.
            The images are from the \hst{}/ACS F814W ($I$-band) images of the galaxies
            in the COSMOS field (e.g., \citealt{Scoville2007}).
            These Advanced Camera for Surveys (ACS) exposures are 50\% completed down 
            to $I=26.0$ mag in aperture with 0.5\asec{} diameter.
            The depth and high resolution (0.09\asec FWHM) make the exposures adequate 
            to test galaxies at Wide depth in the HSC survey.
            In addition to the real cutout image, \synpipe{} models the COSMOS galaxies 
            with $I_{\mathrm{Auto}}\leq25.2$ mag via a method developed by
            \citet{Lackner2012}, and selects between the single-\ser{} and the \ser{} 
            bulge$+$Exponential disk values for the best parametric model.
            Information for these models is stored in the 
            \texttt{GalSim.COSMOSCatalog()} catalogs.
            To access them, the user can provide the ID number of the COSMOS galaxies 
            in the input catalog, or simply let \synpipe{} randomly select the galaxies.

    \end{itemize}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsubsection{Injection of Synthetic Objects into Single-\visit{} Image}
    \label{sssec:addFakes}

    With the input catalog and test data information, \synpipe{} first injects synthetic 
    objects into single-exposure images (step \texttt{addFakes.py}; see the middle panel 
    of Fig \ref{fig:flowchart}).
    \synpipe{} goes through all the individual CCD images that belong to each \visit{} 
    and decides which synthetic objects from the input catalog need to be injected.
    \synpipe{} uses the \textit{initial astrometric calibration} of each single
    \visit{} to convert the input coordinates into locations on the CCD image.
    Using the available models, \synpipe{} reconstructs the PSF at the position of
    each injected object.
    To convert the input magnitudes of synthetic objects into fluxes on the image,
    \synpipe{} also uses the photometric zero point from the single-\visit{}
    calibration.

    For point sources, \synpipe{} generates a rectangular cutout image of the PSF model
    with appropriate size and correct total flux.
    For a galaxy described by single- or double-\ser{} models, \synpipe{} passes the 
    input \ser{} index and effective radius to the \galsim{}\texttt{.Sersic} function to
    create a \ser{} component.
    After stretching and rotating the galaxy to the expected axis ratio and position 
    angle via the \texttt{shear} and \texttt{rotate} methods, \synpipe{} assigns the 
    correct flux to this component.
    For a double-\ser{} model, \synpipe{} uses the \galsim{}\texttt{.Add} method to 
    combine the two \ser{} components.
    For parametric models from the GREAT3 catalog, \synpipe{} uses the
    \texttt{COSMOSCatalog.makeGalaxy} method to generate a corresponding model; in the 
    case of a real high-resolution image, \synpipe{} uses the 
    \galsim{}\texttt{.RealGalaxy} method.
    If necessary, additional shear ($g_1$ and $g_2$) can be applied to the model at 
    this point.
    Then, \synpipe{} passes the reconstructed PSF image into a
    \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy model
    using \galsim{}\texttt{.Convolve}.
    After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to create the 
    image of the simulated galaxy.
    A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires a 
    large image size to cover all of its flux, so \synpipe{} allows the user to 
    truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
    \synpipe{} will also give warning information when a component with $
    n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve 
    accurate PSF convolution in such a model; and the result is, therefore,  
    not a realistic galaxy model.

    After \synpipe{} creates the image of the simulated galaxy, it then shifts the 
    image according to is location on the CCD on a pixel-by-pixel basis; 
    \synpipe{} also crops the images when necessary.
    When that task is complete, \synpipe{} adds appropriate noise to the image based 
    on the pixels it covers and the calibration information of the detector, and creates 
    a variance image that reflects the influence of the synthetic object.
    Since each HSC CCD consists of 4 amplifiers, each with different characteristics,
    \synpipe{} carefully creates a corresponding gain map to provide accurate noise 
    level.
    In the final step, \synpipe{} adds the noise-added image and the variance map to the
    original CCD data while also adding a new \texttt{FAKE} mask bit to the mask plane.
    The middle panel of Fig \ref{fig:flowchart} depicts the result of the
    \texttt{addFakes.py} step by showing an example CCD image before and after the
    synthetic objects are injected.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsubsection{Stacking and Multi-band Measurements}
    \label{sssec:multiband}

    The newly generated single-\visit{} images have the same format and data
    structure as the real HSC data; therefore, \synpipe{} can directly feed them
    to the other \hscpipe{} steps.
    The \texttt{stack.py} step takes the improved astrometric and photometric
    calibrations for each \visit{} from the original reduction and creates
    \coadd{}  images with synthetic objects on them.
    Then the \texttt{multiBand.py} step processes these \coadd{}  images and
    provides standard multi-band measurements in FITS catalogs that are grouped by
    \texttt{Patch} ID.

    Using the infrastructures provided by \hscpipe{}, the user can easily perform the 
    above processes in parallel, and the overall efficiency of \synpipe{} is similar to 
    the real data reduction process.
    The processes also generate the same amount of images and catalog data as the 
    original data reduction process.

    As the last step, \synpipe{} matches the input catalog to the results within a
    radius that is specified by the user, and merges all the \unforced{} and
    \forced{} measurements from all the \texttt{Patches}.
    Each unmatched object from the input catalog is also passed to the results catalog 
    with a unique label.
    In the case of multiple matches, the user can choose between returning only the 
    closest match or returning all objects within the matching radius (the nearest one 
    is labeled as such).

    \song{Susan's edit ends here; and Song has went through the previous sections.}
    \susan{Susan made some minor additional edits to the preceding text on 4/17/2017.}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Star Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_star_sample}
    \end{center}
    \caption{
        Magnitude and color distributions of the synthetic stars.
        \textbf{Upper panel} shows the five-band magnitude distributions of the COSMOS
        stars (filled histograms) and the resampled stars (dashed lines).
        \textbf{Lower panel} shows their color-color distributions
        (left: $g-r$ v.s. $r-i$; right: $i-z$ v.s. $z-y$).
        We show the number density distribution of the COSMOS stars in hexagonal bins,
        and use contours for the resampled stars.
        }
    \label{fig:star_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Galaxy Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sample}
    \end{center}
    \caption{
        Synthetic galaxy distributions of five-band magnitudes (\textbf{upper panel}) and other properties
        (\texttt{lower panel}; from left to right are the \ser{} index, axis ratio,
        effective radius in arcsec, the $g-r$ and $r-z$ colors).
        Objects used for the \textbf{Tract} with good seeing (filled histograms) and
        the one with bad seeing (dashed line) have almost identical distributions.
        }
    \label{fig:galaxy_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\susan{Susan started editing here on 4/17/2017.}
\section{Generation of Synthetic Dataset}
    \label{sec:test}

    Here we demonstrate the application of \synpipe{} via testing the basic performance
    of HSC photometry for both stars and galaxies.

%% ------------------------------------------------------------------------------------ %%
\subsection{Test Data}

    Given the \synpipe{} design and the HSC DR1 data volume \footnote{${\sim}70$
    TB of images}, it is not realistic to test the whole dataset. 
    For this work, we select two \tracts{} in HSC Wide regions
    that fit the following criteria:

    \begin{enumerate}

        \item Not at the edge of any Wide region, so the full \tract{} is
            useful for test.

        \item Not distorted by an extremely bright ($i<12$ mag), saturated star.

        \item Can represent both good and bad seeing conditions in the current $i$-band
            data of the HSC survey.
              \end{enumerate}
              In addition to satisfying these criteria, the selected \tracts {}  show the photometric uncertainties under both good and bad seeing conditions, since for any given star or galaxy, different seeing conditions affect its
            central S/N on the \coadd{} images and, therefore, also impact the
            detection and deblending processes in \hscpipe{}.

The two selected HSC Wide regions are as follows:  
    \tract{}$=9699$ in the VVDS field (median FWHM$=0.449$\asec{} in $i$-band;
   in this paper referred to as \texttt{goodSeeing}) and
    \tract{}$=8764$ in the XMM-LSS field (median FWHM$=0.700$\asec{};
    \texttt{badSeeing}).
    \tract{}$=9699$ also has better seeing in both $r$- and $z$-bands;
    the seeing conditions in $g$- and $y$-bands are very similar for both  \tract{}$=9699$ and \tract{}$=8764$.
    For both \tracts{}, the \coadd{} image in each band includes data from
    20--40 \visits{}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{Input Models for Stars and Galaxies}
    \label{ssec:inputs}

    In this work, we use objects from the high-resolution \hst{}/ACS COSMOS field observations to build input catalogs of realistic synthetic stars and galaxies with appropriate completeness for the Wide survey and with reasonable five-band colors. We first select stars and galaxies as described below;  we then create the input catalogs via randomly distributing the selected synthetic stars and galaxies into the regions covered by the good- and bad-seeing \tracts{}. For stars, we use a number density of 1,000 per CCD; for galaxies, we use a number density of  500 per CCD.

For star selection, we use the sample from \citet{Leauthaud2007}. 
     We use not only the \texttt{SExtractor} stellar classification index (\texttt{CLASS\_STAR}) but also the peak surface brightness above background (\texttt{MU\_MAX}) and
    half-light radius (\texttt{RHL}) to select stars with clean photometry. 
    The classification is reliable down to $I_{\mathrm{F814W}}{\sim}25.2$ mag.
    We match this sample of stars with the HSC UltraDeep--COSMOS data that currently
    reach to about ${\sim}27.2$ mag in $i$-band.
    After filtering the matched sample through basic quality cuts (see \ref{app:qc}),
    we gather five-band HSC photometry for 14,472 stars.
    To get a larger sample for the test, we resample the five-band magnitude
    distributions using \texttt{astroML} (\citealt{astroml}) alternative
    implementation of the \texttt{extreme-deconvolution} algorithm developed by
    \citet{Bovy2011}
    \footnote{\texttt{https://github.com/jobovy/extreme-deconvolution/}} to
    generate 100,000 synthetic stars.
    In Fig \ref{fig:star_sample}, we show that the magnitude and color distributions of
    the real COSMOS stars are nicely recovered by the resampled stars in all
   five-band magnitude and two color-color space distributions. \susan{Later you use "2-D color-color" Should it also be 2-D here?}

For synthetic galaxies, we use the single-\ser{} galaxy models 
    in the COSMOS field that have $I_{F814W} \leq 25.2$ mag.
    These models are created via a method developed by \citet{Lackner2012} on the
    \textit{HST}/ACS images and are included in \galsim{}.
    Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    Although the single-\ser{} model is not always the best choice, its flexibility
    enables it to reasonably describe the flux distribution of most galaxies.
    Also, its simplicity makes it easy for us to diagnose potential problems of the
    \cmodel{} photometry.
    We exclude a tiny fraction of ill-behaved models that have a very high \ser{}
    index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface brightness
    ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing the objects with problematic photometry, we obtain a sample of
    58,210 synthetic galaxies with realistic five-band HSC photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test  \hscpipe{}'s general photometric behaviors, but the sample lacks relatively bright galaxies ($i<20.5$ mag).

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{Running the \synpipe{} Tests}
    \label{ssec:running}

    We use HSC DR1 data stored at Kavli Institute for the Physics and Mathematics of the Universe (Kavli IPMU) to conduct the \synpipe{} tests on the cluster (\texttt{Master}).
    The \texttt{addFakes.py} step takes ${\sim}1.5$ hours per \tract{} in each
    band; the same process takes longer for the galaxy tests (${\sim}3.0$ hours per
    \tract{} in each band) because the \galsim{} simulation is more time consuming.
    The \texttt{stack.py} step and \texttt{multiBand.py} step together take ${\sim}3.5$ hours
    per \tract{}.

    We match the results with the input catalogs using a 2--pixel matching radius
    to generate output catalogs for \forced{} and \unforced{} photometry
    in each band.
    We also match the input catalogs with the original HSC catalogs to identify the
    ``ambiguously blended'' objects that are right on top of the flux peak of a real
    object. \susan{Is ambiguously blended well-understood by your reader? Or is the definition that they are objects that are right on top of the flux peak of a real object? If so, change sentence to the following: We also match the input catalogs with the original HSC catalogs to identify the  ``ambiguously blended'' objects, which are defined as those objects that are right on top of the flux peak of a real
    object.}
    The ambiguously blended objects are useful in the WL-related statistical tests (see Murata et al.
    in prep. for more details); but since they will not help test photometry, they will be excluded later.

    The detailed log of the entire tests can be found at 
    \url{goo.gl/VINOVP}.
    The user should be able to reproduce the results presented in this work using this information. It can also serve as a brief manual for running a \synpipe{} test.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\section{Photometric Performance Results}
    \label{sec:result}

    Here we discuss \hscpipe{}'s general performance, assessed mainly through  \forced{} PSF photometry (for stars) and (\cmodel{}) photometry (for galaxies).
 Although  \hscpipe{} does provide other options (e.g., aperture and Kron photometry), 
these two are the only options that consider the effects of PSF in different bands and are consistent across all bands in terms of position and shape. Therefore, they are the most appropriate and relevant to star and galaxy magnitudes and colors.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $log(S/N)$ measured by \hscpipe{} PSF
        photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}.
        The highly blended stars are highlighted using scatter plots. 
        The gray dashed line marks $S/N = 5$. \susan{"grey" is an acceptable spelling; however, gray is more commonly used in the United States.}
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracy of Forced PSF Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_psf_mag}
    \end{center}
    \caption{
        Accuracy of the \texttt{hscpipe} PSF photometry for synthetic stars measured
        by the difference between input and output \forced{} PSF magnitudes.
        Plots [a, b, c, d, e] show the results for [$g$, $r$, $i$, $z$, $y$]-band, respectively.
        The left panel in each plot shows the relation between input magnitude and
        magnitude difference.
        Hexagonal binned density plots and contours are for the synthetic galaxies from
        good and bad seeing \tracts{}.
        Highly blended objects are highlighted using scattered plots.
        The long-dashed lines mark zero magnitude difference, while the pairs of
        dashed lines outline the running-median of PSF magnitude errors
        (including the uncertainties in aperture correction).
        The right panel in each plot shows the distributions of the magnitude differences for objects
        in good (filled) and bad (solid line) seeing \tracts{}.
        The dashed lines identify the distribution of highly blended objects.
        }
    \label{fig:psf_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        PSF photometry for synthetic stars in $g$- (\textbf{left} and $y$-band
        (\textbf{right}).
        The lines and contours legend is identical to the plots in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracies of the PSF color
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_color}
    \end{center}
    \caption{
        Accuracy of the color measurements for synthetic stars via the differences
        between input and \forced{} PSF colors.
        The \textbf{upper panels} and \textbf{lower panels} are for $g-i$ and $i-y$
        colors separately.
        The \textbf{left} column shows the relation between input magnitude and
        the color difference, and the \textbf{right} column uses the input colors as
        x--axis instead.
        The lines and contours legend is identical to the plot in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_color}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Color-Color Distributions of the Input and Measurements for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_cdist}
    \end{center}
    \caption{
        Evaluations of color measurement accuracy for synthetic stars using
        the color-color distributions.
       The plot on the \textbf{left} (a) uses $(g-r)$ v.s. $(r-i)$ colors, and the plot on the \textbf{right} (b) 
        uses $(i-z)$ and $(z-y)$ colors.
        The filled contours and shaded histograms reflect the distributions for input
        colors.
        The empty contours and solid-line histograms show the distributions recovered
        by \hscpipe{}.
        }
    \label{fig:psf_cdist}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{PSF photometry of stars}
    \label{ssec:psf}

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses the Lanczos interpolation to shift
    the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainty.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see Bosch\etal (in prep.). \susan{Suggest you not include a section number for a document in preparation; section numbers can change before a text is published. }

    In our tests, we randomly inject ${\sim}100,000$ stars into each \tract{} in all
    five bands. Approximately \susan{Spell out approximately and delete approximate sign before 3-4.}
 ${\sim}3$--$4$\% of the stars are located within 2 pixels of the centroids
    of real objects; we remove those stars from the sample to avoid confusion in the comparisons.
    After this step, ${\sim}1$--$3$\% of the remaining synthetic stars still have no match in the \synpipe{} outputs.
    They could either be below the detection threshold or ambiguously blended with real objects.
    The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
    mag fainter than the matched ones.
    The fraction of unmatched stars is slightly higher in the bad seeing \tract{}; this result is consistent with our expectation that magnitude depends on seeing conditions,
    although the difference is not very significant.
    For the matched stars, we select the primary detections
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.

    Next, we isolate the stars that are misidentified by \hscpipe{} as extended
    objects using the \texttt{classification.extendedness} parameter in each band.
    The fraction of misclassification is ${\sim}8$--$15$\%, and clearly depends on
    seeing conditions.
    For the same reason, $g$-band has the highest fraction of misclassification,
    while $i$-band is recommended for selecting point sources.
    The misclassified stars are on average ${\sim}1$ mag fainter than the others.
    We exclude them from the photometric comparison below, and further discuss
    the star--galaxy separation issue in Section \ref{ssec:sg}.

    In the following comparisons, we show the results from the
    \textbf{good-seeing} and \textbf{bad-seeing} \tracts{} separately because it
    is important to understand how the seeing condition affects photometric accuracy as well as to test whether \hscpipe{} can deliver unbiased photometry under different  seeing conditions. \susan{Later you use "slightly worse" and "better" seeing; if you want to use those terms, perhaps define here where you make the distinction? For example, can you quantify slightly worse and better?}

    Besides seeing, ``blendedness'' ($b$; parameter that describes on what level
    the flux distribution of one object is affected by others; please see
    Appendix \ref{app:defineb} for precise definition, and see Murata\etal in prep.
    for details) is another factor that influences the photometric accuracy.
    Therefore, we separate the samples into relatively isolated and highly blended
    groups based on the $b$ values. 
    Assuming zero pixel noise, $b$ of object $A$ reflects the fraction of flux
    within the  $A$ \texttt{Footprint} that comes from other objects.
    Ideally, $b=0$, which means perfect isolation; $b{\sim}1$ means completely
    confused with another object.
    Here we use $b>0.05$ to define highly \susan{Later you say that $b>0.05$ is "highly blended"; it needs to be consistent. If inserting "highly" is incorrect here, please define b for highly blended stars since you use highly blended throughout the text.} blended stars (${\sim}4$--$9$\% are blended). 
    The fraction of highly blended stars slightly increases in the bad seeing \tract{}.
    We highlight these stars separately as well, and will discuss more about the
    impact of blendedness in Section \ref{ssec:blendedness}.

\subsubsection{PSF photometry input magnitude and \s2n{} }

    In discussing the relation between the input magnitude and the
    signal-to-noise ratio of the output PSF photometry, we define \s2n{} simply as
    $\mathrm{Flux}_{\mathrm{PSF}}/\mathrm{Flux\_Err}_{\mathrm{PSF}}$ in each band
    without considering the aperture correction.
    This helps us visualize
    1) the detection limit for point sources,
    2) the impact of seeing condition on PSF photometric,
    3) and the scatter of photometric accuracies at a fixed input magnitude given all
       the realistic features of HSC images (e.g., background subtraction, nearby
       objects, optical artifacts).
    We show the results for $g$- and $i$-bands in Fig \ref{fig:star_sn}, and find close
    relations in both bands.

    As expected, we clearly see that the seeing condition impacts the \s2n{} of point
    sources at fixed input magnitude.
    In  $i$-band, the bad-seeing (0.7\asec{}) \tract{}  shows systematically
    lower \s2n{} than the good-seeing (0.5\asec{}) \tract{}. 
     \s2n{} is similar for the two $g$-band \tracts{} because they share similar seeing conditions.
    Our results show that blendedness can also affect the accuracy of PSF photometry.
    In $i$-band, the highly blended synthetic stars scatter toward lower \s2n{} at
    fixed magnitude.
    Interestingly, the effect is not obvious in $g$-band, suggesting the deblending
    process and photometric quality of blended objects also depend on seeing condition.

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$- and
    $i$-bands at average seeing conditions, which is consistent with the values found
    in \citet{HSCDR1}. \susan{Is there a value for "average seeing conditions"?}
    At the same time, it is worth reminding HSC data users that the detection
    limit will show spatial variations due to the seeing conditions.

\subsubsection{Accuracy of PSF magnitude}

    Now, we look into PSF magnitude accuracy in each of the five bands
    independently.
    Fig \ref{fig:psf_mag} shows the relation between the input magnitude and the
    magnitude difference from the \hscpipe{} \forced{} PSF photometry.

    The overall performance of PSF photometry is excellent.
    In $i$-band, the median magnitude difference for \tract{} with 0.45\asec{} seeing is
     ${\sim}0.015$ mag (${\sim}1.5$\% accuracy) at
    $i_{\mathrm{Input}}{\sim}19.0$ mag.
    At $i_{\mathrm{Input}}{\sim}24.0$ mag, the average accuracy of PSF
    magnitude holds at ${\sim}3$\% level (${\sim}0.036$ mag average difference).
    At $i_{\mathrm{Input}}{\sim}25.2$ mag, the average accuracy decreases to
    ${\sim}7$\% level with an average ${\sim}0.072$ mag difference.
    The bad-seeing \tract{} shows similar accuracy of PSF photometry at
    $i_{\mathrm{Input}}<23.5$ mag.
    At the faint end, the accuracy becomes slightly worse: ${\sim}5$\% accuracy around
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% accuracy around
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the accuracy of PSF photometry via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds a 1-2\% level accuracy,
    which is consistent with our results.
    At fainter magnitude, external comparisons become difficult due to the lack of deep
    imaging data at HSC level and also due to the impact of the color-term when comparing photometry
    using different filters.
    The results from \synpipe{} hence provide more useful evaluations down to the
    detection limit\footnote{Strictly speaking, the uncertainties here should be
    considered as the lower limit because \synpipe{} currently does not consider the uncertainties
    in PSF modeling.}.

    We also see that the accuracy of PSF photometry does not strongly depend on filters.
    The accuracy of the forced PSF magnitude in $r$- and $z$-bands is similar to 
    $i$-band observations (1.5-4.0\% at $19.0 < r_{\mathrm{Input}} < 24.0$ mag, ${\sim}8$\% down to
    $r_{\mathrm{Input}}{\sim}25.2$ mag;
    1.0-5.0\% at $19.0 < z_{\mathrm{Input}} < 24.0$ mag, ${\sim}10$\% down to
    $z_{\mathrm{Input}}{\sim}25.2$ mag).
    The accuracies in $g$- and $y$-bands become slightly worse (1.5-7.0\% at $19.0 < g_{\mathrm{Input}} < 24.0$ mag, ${\sim}13$\% down to
    $g_{\mathrm{Input}}{\sim}25.2$ mag;
    1.5-5.0\% at $19.0 < y_{\mathrm{Input}} < 23.0$ mag, ${\sim}12$\% down to
    $y_{\mathrm{Input}}{\sim}24.0$ mag;); however, the differences can be attributed to the
    differences in seeing conditions.

    For relatively isolated stars, \hscpipe{} provides unbiased PSF photometry down to a
    very faint magnitude in $i$- and $z$-bands.
    For $g$-, $r$-, and $y$-bands, we find small offsets in the distributions of magnitude
    differences, which indicates \hscpipe{} tends to underestimate the fluxes of stars
    in these bands by ${\sim}0.01-0.02$ mag.
    Although the exact cause of this offset is unclear (it could be related to
    differences in seeing), it will not affect the accuracies of magnitudes of \susan{should "of" be "or"?} colors.

    At the same time, the story for highly blended stars is very different.
    \hscpipe{} on average systematically underestimates their total fluxes by
    $0.05$--$0.10$ mag at a fixed input magnitude.
    We see the same effect in all bands, and the impact of blendedness becomes
    increasingly significant at fainter magnitude.
    It is important to keep this in mind when using PSF photometry for point
    sources in HSC data.

    We also test the \unforced{} PSF photometry in all five bands, and find very similar
    accuracies.
    While measuring \forced{} PSF photometry, \hscpipe{} fixes the centroid
    of the PSF model across all five bands.
    Therefore, the  difference between \forced{} and \unforced{} PSF magnitudes
    indicates the photometric uncertainties induced by the differences in the accuracies of
    astrometric calibrations and PSF modeling among different bands.
    We highlight the differences between \forced{} and \unforced{} PSF photometry of
    $g$- and $y$-bands in Fig \ref{fig:psf_diff}.
    The overall difference is very small ($<0.01$ mag), but it also suggests that
    \hscpipe{} \unforced{} PSF photometry systematically recovers slightly more fluxes
    at the very faint end.

\subsubsection{Accuracy of the PSF colors}

    PSF photometry is the most appropriate way to measure colors for point
    sources; therefore, the accuracy of the PSF color estimates is important to many scientific goals of the HSC survey
    (e.g., study of the Milky Way structure, selection of unique stellar objects or
    high-redshift quasars, and accurate star--galaxy separation).

    In Fig \ref{fig:psf_color}, we evaluate the accuracies of the \forced{} $(g-i)$
    and $(i-y)$ PSF colors via comparing the error in color measurements with both
    input magnitudes and input colors.
    These results suggest that \hscpipe{} provides accurate and unbiased PSF color
    estimates for synthetic stars with realistic color distributions down to a very faint
    magnitude.
    For $(g-i)$ color, the average uncertainty around $i_{\mathrm{Input}}{\sim}19.0$ mag
    is ${\sim}0.025$ mag, and increases to ${\sim}0.11$ mag at
    $i_{\mathrm{Input}}{\sim}24.0$ mag.
    The results are very similar for the $(i-y)$ color.
    We also find that the accuracies of PSF colors do not depend on the input colors of
    the synthetic stars or the seeing conditions.
    For instance, the accuracy of $(g-i)$ color for a bad-seeing \tract{} is only 
    slightly worse in the very faint end (${\sim}0.14$ mag at
    $i_{\mathrm{Input}}{\sim}24.0$ mag) compared to the results for a good-seeing \tract{}.

    We further illustrate the accuracies of PSF colors via comparing the distributions
    of colors for the input synthetic stars and the stars recovered by \hscpipe{}
    (see Fig \ref{fig:psf_cdist}).
    Using four different colors and two \susan{2-D?} color--color planes, we show that the \forced{}
    PSF colors can accurately reproduce the distributions of all four colors and the
    2-D \susan{or two?} color-color distributions.

    As for the highly blended stars, on Fig \ref{fig:psf_color}, we see that on average
    they still have larger uncertainties in colors compared to the relatively isolated
    ones.
    For $(g-i)$ color, the average color uncertainty is ${\sim}0.1$ mag larger for the
    highly blended stars at a fixed input magnitude.
    However, the biases in color measurements for these stars are less severe than
    the magnitude measurements.
    Although the blended stars still show slightly bluer color for $(g-i)$ and slightly redder color for $(i-y)$ 
    compared to the isolated stars and the input values, the differences in the
    distributions of color uncertainties of blended and isolated stars are less striking
    than the distribution differences for magnitude.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic galaxies and the $log(S/N)$ measured by \hscpipe{}
        \cmodel{} photometry.
        The lines and contours legend is similar to Fig \ref{fig:star_sn}
        }
    \label{fig:cmodel_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracy of Forced cModel Magnitudes for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_galaxy_mag}
    \end{center}
    \caption{
        Accuracies of the \texttt{hscpipe} \cmodel{} photometry for synthetic
        galaxies measured by the difference between input and output \forced{}
        \cmodel{} magnitudes.
        Plots [a, b, c, d, e] show the results for [$g$, $r$, $i$, $z$, $y$]-bands, respectively.
       The lines and contours legend is identical to Fig \ref{fig:psf_mag}.
        }
    \label{fig:cmodel_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_cmodel_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        \cmodel{} photometry for synthetic galaxies in $g$- (\textbf{left} and
        $y$-band (\textbf{right}).
        The lines and contours legend is identical to the plots in Fig \ref{fig:psf_diff}.
        }
    \label{fig:cmodel_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracies of the cModel color for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_color}
    \end{center}
    \caption{
        Accuracies of the color measurements for synthetic galaxies via the
        differences between input and \forced{} \cmodel{} colors.
        The \textbf{upper panels} and \textbf{lower panels} are for $g-i$ and $i-y$
        colors separately.
        The \textbf{left} column shows the relation between input magnitude (x-axis) and
        the color difference, and the \textbf{right} column uses the input colors as
       the  x--axis.
       The lines and contours legend is identical to the plot in Fig \ref{fig:psf_color}.
        }
    \label{fig:cmodel_color}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Color-Color Distributions of the Input and Measurements for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_cdist}
    \end{center}
    \caption{
        Evaluations of color measurement accuracy for synthetic galaxies
        using the color-color distributions.
        The \textbf{left} panel (a) uses $(g-r)$ vs. $(r-i)$ colors, and the \textbf{right} panel (b)
        uses $(i-z)$ and $(z-y)$ colors.
        The filled contours and shaded histograms reflect the distributions for input
        colors.
        The empty contours and solid-line histograms show the distributions recovered
        by \hscpipe{} \cmodel{} photometry.
        }
    \label{fig:cmodel_cdist}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{cModel photometry of Galaxies}
    \label{ssec:cmodel}

  For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm 
     that is an improved version of the SDSS \cmodel{}
    photometry.
    For more details, please see Bosch\etal (in prep.).\susan{Suggest you not include a section number for a document in preparation; section numbers can change before a text is published. }
    In short, \cmodel{} tries to fit each object using a combination of de Vaucouleurs
    and exponential profiles that are approximated by multiple Gaussian profiles.
    Despite the limitation of the \cmodel{} method (e.g., sensitive to background
    subtraction and deblender failure), it can deliver robust PSF-corrected fluxes and
    colors for galaxies.

    In our tests, we randomly inject ${\sim}40,000$ synthetic galaxies into each 
    \tract{}; and, using an approach similar to that for star selection, we then select galaxy samples for photometric comparisons. 
    First, ${\sim}2$--$3$\% of the synthetic galaxies are ambiguously blended with 
    real objects, and are hence removed from the sample. 
    Among the rest of the sample, ${\sim}2$--$7$\% do not have any match in the 
    \synpipe{} results in different \tracts{} and filters. 
    The fraction of unmatched objects is slightly higher for the $g$- and $y$-bands. 
    For the same filter, the fraction is always higher for the bad-seeing \tract{}.
    The magnitudes of the unmatched galaxies are systematically fainter than the 
    matched ones. 
    In $i$-band, the median input magnitudes for matched and unmatched galaxies are 
    $24.0$ mag and $24.6$ mag, respectively.

    After selecting the matched synthetic galaxies with primary detection and 
    clean photometry (see Appendix \ref{app:qc} for details), we have ${\sim}30,000$
    galaxies in each \tract{} for comparison.  
   \hscpipe{} misclassifies less than $1$\% of these objects as point sources 
    in the good-seeing \tract{}.  
    However,   \hscpipe{} misclassifies more objects ($4$\%) in the bad-seeing \tract{}.
    The median input magnitude for these misclassified sources is ${\sim}24.9$ mag 
    in $i$-band, significantly fainter than the median of other sources.  
    We further discuss this in Section \ref{ssec:sg}.

    We also use $b>0.1$ to define samples of highly blended galaxies.
    Due to the extended nature of the galaxies, the $b$ distributions for synthetic galaxies are 
    very different from the  $b$ distributions for synthetic stars, and they also skew toward higher values (especially for the 
    bad-seeing \tract{}). 
    About ${\sim}7$--$9$\% \susan{Delete the approximately symbol when you use "About"; it's redundant.} of synthetic galaxies are highly blended by this standard.
    The fraction is only slightly higher in the bad-seeing \tract{}, and the 
    $g$-band has the highest fraction of highly blended galaxies. 
    We highlight the highly blended ones in comparisons, and there are other relevant 
    discussions in Section \ref{ssec:blendedness}.

    Although these galaxies span in very wide large ranges of magnitude and 
    structural parameters, the majority of them are faint 
    ($i_{\mathrm{Input}} > 24.0$ mag) and small ($R_{\mathrm{e}} < 0.5$\asec). 
    It is very challenging to obtain good photometric quality of these galaxies, which is crucial
    to achieving several key scientific goals of the HSC survey. 
    The selected samples do not provide enough bright or large galaxies to test. 
    Therefore, we will supplement results for these galaxies using better samples 
    in the future.

%% ------------------------------------------------------------------------------------ %%
\subsubsection{Input magnitude and the \s2n{} of \cmodel{} photometry}

    In Fig \ref{fig:cmodel_sn}, we first show the relation between the input 
    magnitudes and the \s2n{} of the synthetic galaxies estimated by the ratio 
    between \cmodel{} flux and its uncertainties from \hscpipe{}. 
    Since these galaxies are extended objects, the \s2n{} here should be treated 
    as the average \s2n{}, which is different from the \s2n{} of stars using PSF 
    photometry.

    As we did for stars, we use $i$- and $g$-bands as examples for galaxies. 
    The most obvious difference between the  \s2n{} distribution for stars and the  \s2n{} distribution for galaxies is that
    at fixed input magnitude, the galaxy \s2n{} distribution shows much larger
    scatter.  
    At the same total flux, a galaxy can have a different size and shape, which 
    naturally leads to differences in the galaxy's peak and average \s2n{}.  
    This also makes it difficult to have a clear definition of ``detection limit'' 
    for galaxies. 
    For $i$-band, the typical \s2n{} for synthetic galaxies at the magnitude limit 
    of our sample ($25.2$ mag) is ${\sim}20$, but at that limit the low-\s2n{} tail has  
    reached the \s2n{}$=5$ boundary.  
    This suggests that, although HSC Wide can detect galaxies as faint as 
    $i{\sim}26.0$ mag, the sample at that magnitude will be significantly incomplete. 
    Furthermore, the magnitude limit can also introduce selection bias because galaxies with certain structural properties are harder to detect than others. 
    HSC survey data users who want to select flux-limited samples of galaxies,
    or intend to study the population of faint (e.g., high-$z$) galaxies should
    keep this in mind.

    Similar to the synthetic stars, the worse the seeing condition, the more 
    we noticed a decrease in \s2n{} at the same magnitude (see the figure \susan{cite actual figure number.} for
    $i$-band). 
    Meanwhile, the highly blended synthetic galaxies do not show significantly 
    different \s2n{} distributions here.

%% ------------------------------------------------------------------------------------ %%
\subsubsection{Accuracy of the \cmodel{} magnitude}

    In Fig \ref{fig:cmodel_mag}, we illustrate the photometric uncertainties of the 
    \cmodel{} magnitudes in each band using the same format as 
    Fig \ref{fig:psf_mag}.

    The overall performance of \cmodel{} photometry is very reasonable all the way 
    down to $i_{\mathrm{Input}}=25.2$ mag. 
    Compared to the qualities of PSF photometry for stars, the magnitude differences
    of \cmodel{} show much larger scatter for galaxies. 
    Given the diversity of properties of these galaxies and the complexities in the 
    model-fitting process, this difference is expected.  
    At the same time, we can see that the \cmodel{} algorithm in \hscpipe{} can 
    provide unbiased and consistent photometry for galaxies in different bands and 
    seeing conditions.

    In $i$-band, the median flux uncertainty of the \forced{} \cmodel{} is ${\sim}7$\% 
    in the good-seeing \tract{} at $i_{\mathrm{Input}}{\sim}20.0$ mag, 
    and ${\sim}9$\% at $i_{\mathrm{Input}}{\sim}24.0$ mag.  
    Down to $i_{\mathrm{Input}}{\sim}25.0$ mag, the median uncertainty maintains 
    at ${\sim}11$\% level. 
    For the bad-seeing \tract{} in $i$-band, the performance is very similar. 
    From $i_{\mathrm{Input}}{\sim}20.0$ to $25.0$ mag, the median magnitude difference
    between output and input values changes from \plus{}$0.041$ mag to 
    \minus{}$0.005$ mag, which suggests that \forced{} \cmodel{} photometry is 
    unbiased down to the very faint end. 
    It is worth noting that the photometric uncertainties estimated by \synpipe{}
    are significantly larger than the fiducial error from \hscpipe{}, which does not 
    take into account the systematic uncertainties involved in the modeling 
    fitting process. 
    Although the synthetic galaxies in our sample are described by the simple single-\ser{} 
    model, the\cmodel{} algorithm can only approximate them to a certain accuracy, 
    especially when the noise level, deblending uncertainty, and priors of model 
    parameters are considered. 
    Given that real galaxies are more complex in structural details, these 
    uncertainties should be treated as lower limits.

    The performance of \forced{} \cmodel{} photometry is very consistent across all
    filters. 
    For $g$-, $r$-, and $z$-bands, the photometric uncertainties are very similar to 
    the ones for $i$-band between $20.0$ and $25.0$ mag. 
    The \forced{} \cmodel{} magnitudes are also unbiased for these three filters 
    across the entire input magnitude range. 
    The uncertainties for $y$-band are slightly larger, ranging from 9\% to 17\% in the 
    same magnitude bins. 
    In addition, the \forced{} \cmodel{} tends to overestimate the total flux in $y$-band
    at $y_{\mathrm{Input}}>24.0$ mag. 
    Mean magnitude differences change from \minus{}0.06 at 24.0 mag to \minus{}0.22
    mag at 25.0 mag.  
    Both a worse seeing condition and a higher background noise level can make it more 
    difficult to detect a faint object on a $y$-band image.
    At $y_{\mathrm{Input}}{\sim}24.0$, the average \s2n{} for \cmodel{} in $y$-band
    is already close to the detection threshold.

    As mentioned, the statistics for brighter galaxies are very uncertain due to the
    small sample size. 
    However, we do notice that, at $i_{\mathrm{Input}}<20.0$, the \forced{} 
    \cmodel{} photometry shows larger uncertainty and starts to systematically 
    underestimate the total flux.  
    This problem is confirmed via comparing with more carefully derived total fluxes 
    for bright galaxies (e.g., \citealt{HSCDR1}, Huang\etal in prep.). 
    It relates to the known issues of \hscpipe{} (e.g., \hscpipe{}over-deblends around bright objects; inappropriately weights priors in \cmodel{}); but it may also suggest
    that, at the imaging depth of HSC, assumptions in \cmodel{} are no longer the 
    best choice in modeling these galaxies (e.g., for massive elliptical galaxies; 
    see Huang\etal in prep.~for more details).

    For the highly blended synthetic galaxies, we also see systematically 
    underestimated total flux (average offset ${\sim}0.1$ mag) and larger scatter 
    of photometric uncertainty in all five bands.  
    Meanwhile, the contrast between relatively isolated and highly blended galaxies 
    in photometric performance is not as staggering as the one for synthetic
    stars (Fig \ref{fig:psf_mag}).

    We also test the \unforced{} \cmodel{} photometry in all five bands, and the 
    results suggest similar accuracies.
    We highlight the differences between \forced{} and \unforced{} \cmodel{} 
    photometry in $g$- and $y$-bands in Fig \ref{fig:cmodel_diff}.
    The only noticeable difference is for galaxies brighter than $24.0$ mag in 
    $g$-band; their \forced{} \cmodel{} magnitudes are systematically brighter than 
    the \unforced{} ones.

%% ------------------------------------------------------------------------------------ %%   
\subsubsection{Accuracy of the \cmodel{} color}

    The consistent and accurate five-band \cmodel{} colors from \hscpipe{} are crucial 
    to the photometric redshift and spectral energy distribution (SED) fitting results. 
    They are also key to the selection of high-$z$ Lyman-break galaxies (LBG) and 
    and many other interesting objects.

    In Fig \ref{fig:cmodel_color}, we demonstrate the accuracies of \forced{} $(g-i)$
    and $(i-y)$ \cmodel{} colors in the same format used for synthetic stars 
    (Fig \ref{fig:psf_color}).
    These plots show that \hscpipe{} can provide reliable \cmodel{} colors for 
    synthetic galaxies down to $i_{\mathrm{Input}}{\sim}25.0$ mag. 
    For $(g-i)$ colors, the average uncertainty around $i_{\mathrm{Input}}{\sim}21.0$ mag 
    is ${\sim}0.04$ mag.
    The uncertainty increases to ${\sim}0.11$ mag at $i_{\mathrm{Input}}{\sim}25.0$ mag.
    Results for $(i-y)$ colors are similar.

    The performance of \forced{} \cmodel{} color does not show dependence on seeing
    condition, and it is unbiased across the ranges for input total magnitude and color.
    The only noticeable feature is that the $(i-y)$ color at 
    $i_{\mathrm{Input}}>24.0$ is slightly overestimated. 
    As mentioned above, this is not surprising given the shallower imaging depth of  the
    $y$-band.    
    At the same time, high blendedness of synthetic galaxies can still increase the 
   color uncertainties and bias the color measurements.
    The highly blended galaxies on average show slightly bluer  $(g-i)$ colors
     and redder  $(i-y)$ colors  than the input values.

    Similar to the star comparison in Fig \ref{fig:psf_cdist}, we also compare the color--color distributions 
    of the input and output \forced{} \cmodel{} values in Fig \ref{fig:cmodel_cdist}. 
    The general color distributions are well recovered despite the fact that the galaxy accuracy is 
    not as good as the accuracy for stars, especially on the $(i-z)$ vs. $(z-y)$ plane. 
    That result is expected given the narrow dynamical ranges of these colors among red filters.
    The \forced{} \cmodel{} tends to slightly underestimate the $(g-r)$ and $(i-z)$ 
    colors at the very ``blue'' end, while overestimating the $(i-y)$ color at the very 
    red end.

   We should point out that we assume constant color for the entire galaxy.  
    In reality, different colors for individual components and a smooth color gradient 
    can complicate the situation; hence, the uncertainties of \cmodel{} colors 
    here should be considered as lower limits.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Astrometric Calibration
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_astrometry}
    \end{center}
    \caption{
        Astrometric accuracy for a synthetic star (\textbf{left}) and a galaxy
        (\textbf{right}) via the differences in R.A and Dec between the input
        coordinates and the ones measured on the \coadd{} images.
        Hexagonal bins (filled histogram) and contours (solid-line histogram) are for
        the synthetic objects from \tracts{} with good and bad seeing conditions, respectively.
        $\Delta\mathrm{R.A}=0$ and $\Delta\mathrm{Dec}=0$ are marked by dashed lines.
        }
    \label{fig:astrometry}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Star-Galaxy Separation
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_star_galaxy}
    \end{center}
    \caption{
        Problems with the current  \hscpipe{}star--galaxy separation method, which is based on the magnitude difference between the PSF and
        \cmodel{} photometry.
       The plot on the \textbf{left} (a) illustrates distributions for the synthetic stars.
        Hexagonal bins and filled histogram show distributions of correctly classified
        synthetic stars in the good-seeing \tract{}.
        The scatter plot (dashed-line histogram) and contour (solid-line histogram)
        show distributions of synthetic stars that are misclassified as extended
        objects in \tracts{} with bad and good seeing conditions, respectively.
        The plot on the \textbf{right} (b) illustrates distributions for synthetic galaxies.
        Hexagonal bins (filled histogram) and contours (solid-line histogram)
        are for galaxies from good- and bad-seeing \tracts{}, respectively.
        The scatter plot and dashed-line histogram are for synthetic galaxies that are
        misclassified as point sources.
        Zero magnitude difference is highlighted using dashed lines on both plots.
        }
    \label{fig:sg}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Relations between Blendedness and Photomertic Errors
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_blendedness_err}
    \end{center}
    \caption{
        The relation between the $\log(\mathrm{Blendedness})$ parameter
        and the photometric accuracies for synthetic stars (\textbf{upper row}, a and b) and
        galaxies (\textbf{lower row}, c and d) .
       The \textbf{left} column (a and c) is for uncertainties of PSF or \cmodel{} magnitude,
        and the \textbf{right} column (b and d) is for uncertainties of $(g-i)$ color.
        Hexagonal bins and contours show results for good- and bad-seeing
        \tracts{}, respectively.
        $\log(\mathrm{Blendedness}) = -1.0$ is marked using a vertical dashed line.
        }
    \label{fig:blend}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Compare with the behaviours of the Kron photometry for galaxies
%\begin{figure*}
%    \begin{center}
%        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_kron}
%    \end{center}
%    \caption{
%        Basic performance of the Kron photometry for synthetic galaxies.
%        \textbf{Left} shows the accuracy in $i$-band.
%        magnitude. The format is identical to Fig \ref{fig:galaxy_mag}.
%        \textbf{Right} shows the accuracy of the $(g-i)$ color using
%        Kron photometry.
%        Other formats are identical to Fig \ref{fig:cmodel_color}.
%        }
%    \label{fig:galaxy_kron}
%\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: HSC cModel has problem recovering the size and shape of galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_ba}
    \end{center}
    \caption{
        Accuracy of the weighted axis ratio measurements for synthetic
        galaxies using \cmodel{} photometry.
      The  \textbf{left} (a) plot shows the relation between the input $(b/a)$ and the
        uncertainties of $(b/a)$ estimates.
        The \textbf{right} (b) plot shows the relation between input magnitude of a synthetic galaxy
        and the $(b/a)$ uncertainties.
        Contours are for the default test with \texttt{usePixelWeights=False}, and the
        hexagonal bins are used for same tests with \texttt{usePixelWeights=True}.
        }
    \label{fig:galaxy_ba}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

\section{Results on Other Aspects of \hscpipe{}}
    \label{sec:others}

    In this section, we briefly discuss a few other  \hscpipe{} aspects that relate 
    to astrometric calibration, star-galaxy separation, blendedness and photometric errors, and galaxy shape measurements to demonstrate the range of topics that can be tested with \synpipe{}.

%% ------------------------------------------------------------------------------------ %%
\subsection{Astrometric Calibration}
    \label{ssec:astrometry}

    For a \synpipe{} test, we inject synthetic objects to single-\visit{} images using 
    the less accurate initial astrometric calibration. 
    During the image stacking step, the joint calibration process  improves the 
    astrometric solutions. 
    Therefore, the differences between the input and output coordinates from \synpipe{}
    can help us understand the systematic effect of the image stacking step.

    In Fig \ref{fig:astrometry}, we plot the distributions of astrometric offsets for 
    both synthetic stars and galaxies in good- and bad-seeing \tracts{}. 
    Small systematic offsets at ${\sim}10$--20 mas can be noticed in either direction. 
    The two \tracts{} apparently show different systematic offsets. 
    In the same \tracts{}, synthetic stars and galaxies show very similar offsets, 
    although the scatter for galaxies is slightly larger due to their extended nature.

%% ------------------------------------------------------------------------------------ %%
\subsection{Star-Galaxy Separation}
    \label{ssec:sg}

    Star--galaxy separation is an important aspect of a cosmology survey,
    and it becomes increasingly difficult at very faint magnitude ($i<24.0$ mag). 
    In \hscpipe{}, the star--galaxy classification primarily depends on the extendedness 
    value, which is measured by the magnitude difference between PSF and \cmodel{} 
    photometry in $i$-band. 
    For more details about this algorithm, please see Bosch\etal 
    (in prep.). \susan{Suggest you not include a section number for a document in preparation; section numbers can change before a text is published. }

    In Fig \ref{fig:sg}, we show the current status of star--galaxy classification  
    via comparing the input magnitude with the difference between PSF and \cmodel{} 
    magnitudes for both synthetic stars and galaxies, and we highlight the distributions 
    for misclassified objects. 
    As described above, more than $20$\% of synthetic stars at 
    $i_{\mathrm{Input}}>23.0$ mag are misclassified as extended objects, while only a 
    tiny fraction of faint synthetic galaxies ($i_{\mathrm{Input}}>24.0$) are 
    misclassified. 
    This confirms that the current star--galaxy classification strategy in \hscpipe{}
    results in a galaxy sample with \textbf{high completeness} and star sample 
    with \textbf{high purity}.

    It is also clear that seeing condition strongly modifies the distributions 
    of $i_{\mathrm{PSF}}-i_{\mathrm{cModel}}$, where worse seeing conditions clearly make it 
    more difficult to separate star and galaxy.  
    This implies that the completeness or purity of the galaxy or star sample at the faint end may depend on seeing; therefore, seeing condition deserves the attention of HSC users who care about these properties.

    In the next data release, \hscpipe{} will include a better star--galaxy classifier, which will be based on a machine learning method that takes shape and color information into account. We will test its improvement using \synpipe{} in the future.

%% ------------------------------------------------------------------------------------ %%
\subsection{Blendedness and Photometric Errors}
    \label{ssec:blendedness}

    In the above sections, we have pointed out the impact of high blendedness on 
    photometry through a small sample of highly blended objects.  
    Here, in Fig \ref{fig:blend}, we further explore the effect of blendedness via 
    directly plotting it against the uncertainties of magnitudes and colors.

    For both PSF and \cmodel{} magnitude distributions, we can see clear trends that total fluxes are systematically underestimated for highly blended
    ($b>0.05$ for stars or $0.1$ for galaxies) objects. 
    This effect is particularly strong for stars. 
    As for colors, higher blendedness still results in higher uncertainties for both 
    stars and galaxies, but the trends are different for each. 
    Highly blended stars have clearly overestimated $(g-i)$ colors, and the 
    highly blended galaxies seem to have  $(g-i)$ colors that are bluer than the true values.

    These blendedness-related uncertainties are not included in the photometric errors released by \hscpipe{}.  
    We therefore remind the HSC users to treat the highly blended object with 
    greater caution.

%% ------------------------------------------------------------------------------------ %%
%\subsection{Comparing to Kron photometry}

%    \plan{Uses $i$-band as example to briefly discuss Kron photometry}

%    \plan{Although magnitude for galaxies is Ok, the color becomes worse}

%% ------------------------------------------------------------------------------------ %%
\subsection{Shape and Structural Parameters of Galaxies}
    \label{ssec:shape}

    In SDSS,  structural parameters like size and shape from \cmodel{} are often 
    used to study the evolution galaxies. 
    Although the \cmodel{} in \hscpipe{} follows a very similar algorithm, its primary 
    goal now is to provide accurate and consistent magnitude and color in all five 
    HSC filters for a vast majority of faint, small galaxies.  
    To improve the stability of \cmodel{} for these low-\s2n{} and barely resolved 
    galaxies, \hscpipe{} uses Bayesian priors on a radius and axis ratio that strongly 
    favors these galaxies. 
    However, the current implementation of these priors has a flaw that leads to 
    severely underestimated size and overestimated axis ratio for most galaxies.

    In Fig \ref{fig:galaxy_ba}, we demonstrate this issue using the difference between 
    the \cmodel{} axis ratio and the input values as an example.
    Here, $({\mathrm{b}/\mathrm{a}})_{\mathrm{CModel}}$ is a combination of the 
    axis ratios estimated by the exponential and de Vaucouleurs models weighted by their flux ratio.

    During the investigation, we noticed that one of the problems relates to the \cmodel{} \texttt{usePixelWeight} configuration, which decides whether to 
    use per-pixel variances as weights in the nonlinear fit. 
    By default, \hscpipe{} chooses \texttt{usePixelWeight}$=$\texttt{False}. 
    As one can see, this leads to a significantly overestimated axis ratio for almost 
    all galaxies with $i_{\mathrm{Input}}>21.5$ mag\footnote{It also leads to an
    underestimated half-light radius for these galaxies.}.
    At the same time, we also repeat the \synpipe{} run for galaxies with 
    \texttt{usePixelWeight}$=$\texttt{True}, and this change clearly mitigates
    the problem. When we use \texttt{usePixelWeight}$=$\texttt{True},
     $({\mathrm{b}/\mathrm{a}})_{\mathrm{CModel}}$ can provide an unbiased axis ratio estimation 
    for galaxies with $i_{\mathrm{Input}}< 24.0$ mag.

    For the current HSC data release, we warn users against using the \cmodel{} 
    size and shape to study a galaxy with $i_{\mathrm{Input}}>21.5$. 
    At the same time, \synpipe{} will try to help \hscpipe{} improve the details 
    of the \cmodel{} so that it can provide more useful structural information without hurting its photometric performance.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\section{Summary and Conclusion}
    \label{sec:summary}

    %% Short summary
    In February 2017, the HSC survey made its first public data release, and the survey continues to
    produce increasingly larger amounts of high-quality imaging data.
    To help achieve the designed goals of the survey and facilitate scientific
    outputs, we developed \synpipe{}, a flexible framework that is based on 
    \hscpipe{}, to examine the data reduction process and check the accuracy of all the HSC pipeline outputs.
    \synpipe{} injects synthetic stars and galaxies into the desired locations on the
    single-\visit{} images with the help of the original astrometric and
    photometric calibrations.
    These data are then processed by \hscpipe{} to generate the \coadd{} images and
    all the photometric measurements.
    In this way, \synpipe{} can simulate the realistic data reduction process
    to an extent that many subtle systematical effects can be taken into
    consideration, which is helpful for achieving the outstanding scientific goals of the HSC
    survey.
    In this work, we choose two \tracts{} in the HSC Wide layer with
    representative seeing conditions (good and bad) to test the general behaviors of HSC
    photometry for both stars and galaxies.

    From these tests, we find the following:

    \begin{enumerate}

        \item The \forced{} PSF photometry provides excellent magnitude and color
            measurements for isolated stars at $18.0 < i < 25.0$.
            The typical uncertainties of HSC forced PSF photometry for stars range
            from around 0.01 mag at $i{\sim}18.0$ mag to 0.08 mag at $i{\sim}25.0$
            mag (1\%-7\% accuracy in the $i$-band).
            The \forced{} PSF photometry can accurately recover the color--color
            sequences for stars.

        \item The \forced{} \cmodel{} photometry is reliable for synthetic galaxies
            with realistic distributions of structural parameters and colors at
            $20.0 < i < 25.0$ mag.
            The average accuracy of \forced{} \cmodel{} magnitude ranges from
            ${\sim}7$\% at $i{\sim}20.0$ mag to ${\sim}10$\% at $i{\sim}25.0$.
            The \forced{} \cmodel{} colors for galaxies are accurate and unbiased.
            The accuracies of colors do not show dependence on input magnitude and
            color.

        \item The \forced{} PSF and \cmodel{} photometry are robust against different
            seeing conditions, despite the fact that worse seeing leads to a lower
            $\mathrm{S}/\mathrm{N}$ at fixed input magnitude.

        \item High blendedness ($b>0.1$) \susan{This is the value you've given for highly blended galaxies; should the text in the parentheses read ($b>0.1$ for galaxies and $b>0.05$ for stars)}impacts the accuracy of both PSF and \cmodel{}
            photometry, especially for stars.
            For highly blended stars, the \forced{} PSF photometry on average
            overestimates the magnitudes of stars by 0.1–-0.2 mag.
            For galaxies, high-blendedness typically adds an additional 0.05 mag
            uncertainty in both magnitude and color estimates.

    \end{enumerate}

    We also notice a few issues in the current \hscpipe{}:

    \begin{enumerate}

        \item The performance of the current star--galaxy separation algorithm in \hscpipe{}
            is not perfect, especially at the faint magnitude end and under poor seeing
            conditions.
            The fraction of stars that are misclassified as extended objects is still
            quite high ($>20$\% at $i> 22.5$ mag).

        \item The nominal \cmodel{} setting  (\texttt{usePixelWeights=True}) results
            in very biased estimates of shape and size for galaxies, especially for the
            relatively faint ones.
            The axis ratio of a galaxy using \cmodel{} is highly overestimated, while
            the effective radius is underestimated.
            We show that we can significantly improve the result if we use \texttt{usePixelWeights=True} instead.
            However, that choice may impact the accuracy of the deblender, which we will test in the future.

    \end{enumerate}

    The HSC collaboration is working on improving these issues in the future version
    of \hscpipe{} and the subsequent data release.
    Meanwhile, we suggest that HSC data users keep these caveats in mind.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
    %% Other applications
    Although we mainly focus on the general behaviors of photometric measurements in this study,
    \synpipe{} is quite flexible and has many other applications for any HSC
    dataset.
    For instance,  Murata\etal (in prep.) uses \synpipe{} to test
    the impacts of extremely blended \susan{Quantify? Is "extremely blended" more blended than "highly blended"?} objects on the WL measurements.
    Also, \citep{Niikura2017} applies \synpipe{} to the high-cadence M31 HSC observations
    to help users search for interesting microlensing events.
  \synpipe{} is also currently being used to estimate the completeness of the selections 
    for high-$z$ LBG and Lyman- $\alpha$ emitters 
    (Ono\etal in prep.; Konno\etal in prep.), check the dependence of the detection and 
    completeness of high-$z$ LBG pairs on the pair separation (Harikane\etal in prep.),
    identify low surface--brightness dwarf galaxies (Greco\etal in prep.),
    and test photometric and WL measurements around nearby clusters (Chiu\etal in prep.).

    %% Caveats
    The limitations of the current \synpipe{} include the following:

    \begin{enumerate}

        \item \citet{HSCDR1} points out that \hscpipe{} tends to over-subtract the
            background around bright objects.
            Unfortunately, \synpipe{} now takes the original background subtraction on
            the single-\visit{} image for granted; hence, it lacks the capability to
            test or help improve this problem.

        \item  \synpipe{} simply adopts the PSF measured by \hscpipe{} and
            uses it as a model of point source and PSF convolution for a galaxy.
            Hence, it cannot be used to test how the uncertainty of the PSF modeling
            affects the photometry and shape measurements.
            This could be important for regions with exquisite seeing \susan{Quantify "exquisite"? Is it better than "good" or "better" seeing? Suggest using consistent adjectives.}  that makes the
            PSF modeling difficult, and for accurate WL measurements.

        \item \synpipe{} works in a ``unit'' of \visit{} for a single-exposure
            test, and uses a \tract{} to test \coadd{} images.
            In the case of tests that focus on specific regions that are much smaller than
            the size of a \visit{} or a \tract{} (e.g., quality of photometry
            around rich galaxy clusters), using \synpipe{} often leads to
            low efficiency.

    \end{enumerate}

    %% Future developments
    Both \hscpipe{} and \synpipe{} are currently undergoing active development. The plan is to update \synpipe{} to provide a photometric benchmark for
    each HSC survey data release, which will help \hscpipe{} improve.
    At the same time, we are also working on improving the efficiency of \synpipe{}.
    Since many survey data users are not concerned with the subtle effects involved in the image stacking process, we will update \synpipe{} so that it can directly inject synthetic objects to the \coadd{} images and speed up these tests.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\begin{ack}
    \label{sec:ack}

    % HSC part
    The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of
    Japan and Taiwan, and Princeton University.
    The HSC instrumentation and software were developed by National Astronomical
    Observatory of Japan (NAOJ), Kavli Institute for the Physics and Mathematics of
    the Universe (Kavli IPMU), University of Tokyo, High Energy Accelerator
    Research Organization (KEK) in Japan,  Academia Sinica Institute for Astronomy and
    Astrophysics  (ASIAA) in Taiwan, and Princeton University in the United States.
    Funding was contributed by the FIRST program from Japanese Cabinet Office; Ministry of Education, Culture, Sports, Science and Technology (MEXT); Japan
    Society for the Promotion of Science (JSPS); Japan Science and Technology Agency
    (JST); Toray Science  Foundation; NAOJ; Kavli IPMU; KEK; ASIAA; and Princeton University.

    % LSST software part
    This paper makes use of software developed for the Large Synoptic Survey Telescope.
    We thank the LSST Project for making their code available as free software at
    http://dm.lsstcorp.org.

    % Pan-STARRS1 part
    Pan-STARRS1 Surveys (PS1) have been made possible through contributions of 
    Institute for Astronomy, University of Hawaii, Pan-STARRS Project Office,
    Max-Planck Society and its participating institutes (Max Planck Institute
    for Astronomy, Heidelberg, and Max Planck Institute for Extraterrestrial Physics,
    Garching), Johns Hopkins University, Durham University, University of
    Edinburgh, Queen's University Belfast, Harvard-Smithsonian Center for Astrophysics,
    Las Cumbres Observatory Global Telescope Network Incorporated,  National
    Central University of Taiwan, Space Telescope Science Institute,  National
    Aeronautics and Space Administration under Grant No.
    NNX08AR22G issued through Planetary Science Division of NASA Science
    Mission Directorate, National Science Foundation under Grant No. AST-1238877,
    University of Maryland, Eötvös Loránd University (ELTE), and Los Alamos
    National Laboratory.

    % Software part
    This research made use of
    {\texttt{Astropy}},
        a community-developed core Python package for Astronomy (\citealt{Astropy};
        \url{http://www.astropy.org/});
    {\texttt{astroML}},
        a machine learning library for astrophysics (\citealt{astroml};
        \url{http://www.astroml.org/});
    {\texttt{SciPy}},
        an open source scientific tool for Python (\citealt{SciPy};
        \url{http://www.scipy.org/});
    {\texttt{NumPy}},
        a fundamental package for scientific computing with Python (\citealt{NumPy};
        \url{http://www.numpy.org/});
    {\texttt{Matplotlib}},
        a 2-D plotting library for Python (\citealt{Matplotlib};
        \url{http://matplotlib.org/}); and
    {\texttt{scikit-learn}},
        a machine learning library in Python (\citealt{scikit-learn};
        \url{http://scikit_learn.org/} \susan{I think correct URL is http://scikit-learn.org/ (hyphen, not underscore.}).

\end{ack}
%% ------------------------------------------------------------------------------------ %%

%%%%%%%%%%: Bibliographic Section %%%%%%%%%%

%\bibliography{synpipe}

%% ------------------------------------------------------------------------------------ %%
\bibliographystyle{apj}

%% ------------------------------------------------------------------------------------ %%
\susan{Susan did not review bibliographic citations.}
\begin{thebibliography}{}
    \label{sec:ref}
    \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

    %% A %%
    \bibitem[Abazajian et al.(2004)]{Abazajian2004} Abazajian, K., Adelman-McCarthy,
             J.~K., Ag{\"u}eros, M.~A., et al.\ 2004, \aj, 128, 502

    \bibitem[Aihara et al.(2017)]{HSCDR1} Aihara, H., Armstrong, R., Bickerton, S.,
             et al.\ 2017, arXiv:1702.08449

    \bibitem[Antilogus et al.(2014)]{2014JInst...9C3048A} Antilogus, P., Astier, P.,
             Doherty, P., Guyonnet, A., \& Regnault, N.\ 2014, Journal of
             Instrumentation, 9, C03048

    \bibitem[Astropy Collaboration et al.(2013)]{Astropy} Astropy Collaboration,
             Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33

    \bibitem[Axelrod et al.(2010)]{Axelrod2010} Axelrod, T., Kantor, J., Lupton,
             R.~H., \& Pierfederici, F.\ 2010, \procspie, 7740, 774015

    %% B %%
    \bibitem[Bartelmann \& Schneider(2001)]{Bartelmann2001} Bartelmann, M., \&
             Schneider, P.\ 2001, \physrep, 340, 291

    \bibitem[Ben{\'{\i}}tez(2000)]{Benitez2000} Ben{\'{\i}}tez, N.\ 2000, \apj,
             536, 571

    \bibitem[Bertin(2011)]{Bertin2011} Bertin, E.\ 2011, Astronomical Data Analysis
             Software and Systems XX, 442, 435

    \bibitem[Bertin(2013)]{Bertin2013} Bertin, E.\ 2013, Astrophysics Source
             Code Library, ascl:1301.001

    \bibitem[Bolzonella et al.(2000)]{Bolzonella2000} Bolzonella, M., Miralles, J.-M.,
            \& Pell{\'o}, R.\ 2000, \aap, 363, 476

    \bibitem[Bovy Jo et al.(2011)]{Bovy2011} Bovy Jo, Hogg, D.~W., \& Roweis,
             S.~T.\ 2011, Annals of Applied Statistics, 5,

    %% C %%
    \bibitem[Chang et al.(2015)]{Chang2015} Chang, C., Busha, M.~T., Wechsler, R.~H.,
             et al.\ 2015, \apj, 801, 73

    %% D %%
    \bibitem[Dressler et al.(2012)]{Dressler2012} Dressler, A., Spergel, D., Mountain,
             M., et al.\ 2012, arXiv:1210.7809

    %% G %%
    \bibitem[Greisen \& Calabretta(2002a)]{WCS1} Greisen, E.~W., \& Calabretta, M.~R.\
             2002, \aap, 395, 1061

    \bibitem[Calabretta \& Greisen(2002b)]{WCS2} Calabretta, M.~R., \& Greisen, E.~W.\
             2002, \aap, 395, 1077

    %% H %%
    \bibitem[Hunter (2007)]{Matplotlib} Hunter J. D., 2007, Computing In Science \&
             Engineering, 9, 90

    %% I %%
    \bibitem[Ilbert et al.(2009)]{Ilbert2009} Ilbert, O., Capak, P., Salvato,
             M., et al.\ 2009, \apj, 690, 1236

    \bibitem[Ivezic et al.(2008)]{Ivezic2008} Ivezic, Z., Axelrod, T., Brandt, W.~N.,
             et al.\ 2008, Serbian Astronomical Journal, 176, 1

    %% J %%
    \bibitem[Jones et al.(2001)]{SciPy} Jones E., Oliphant T., Peterson P., et al.,
             2001, SciPy: Open source scientific tools for Python,
             \url{http://www.scipy.org/}

    \bibitem[Juri{\'c} et al.(2015)]{Juric2015} Juri{\'c}, M., Kantor, J., Lim, K.,
             et al.\ 2015, arXiv:1512.07914

    %% K %%
    \bibitem[Kaiser \& Squires(1993)]{Kaiser1993} Kaiser, N., \& Squires, G.\
             1993, \apj, 404, 441

    %% L %%
    \bibitem[Lackner \& Gunn(2012)]{Lackner2012} Lackner, C.~N., \& Gunn, J.~E.\
             2012, \mnras, 421, 2277

    \bibitem[Laureijs et al.(2012)]{Laureijs2012} Laureijs, R., Gondoin, P., Duvet, L.,
             et al.\ 2012, \procspie, 8442, 84420T

    \bibitem[Leauthaud et al.(2007)]{Leauthaud2007} Leauthaud, A., Massey, R.,
             Kneib, J.-P., et al.\ 2007, \apjs, 172, 219

    \bibitem[Lupton et al.(2001)]{Lupton2001} Lupton, R., Gunn, J.~E., Ivezi{\'c},
             Z., Knapp, G.~R., \& Kent, S.\ 2001, Astronomical Data Analysis Software
             and Systems X, 238, 269

    %% M %%
    \bibitem[Magnier et al.(2013)]{Magnier2013} Magnier, E.~A., Schlafly, E.,
             Finkbeiner, D., et al.\ 2013, \apjs, 205, 20

    \bibitem[Mandelbaum et al.(2014)]{Mandelbaum2014} Mandelbaum, R., Rowe, B.,
            Bosch, J., et al.\ 2014, \apjs, 212, 5

    \bibitem[Mandelbaum et al.(2015)]{Mandelbaum2015} Mandelbaum, R., Rowe, B.,
             Armstrong, R., et al.\ 2015, \mnras, 450, 2963

    \bibitem[Miyazaki et al.(2012)]{Miyazaki2012} Miyazaki, S., Komiyama, Y., Nakaya,
             H., et al.\ 2012, \procspie, 8446, 84460Z

    %% N %%
    \bibitem[Niikura et al.(2017)]{Niikura2017} Niikura, H., Takada, M., Yasuda, N.,
             et al.\ 2017, arXiv:1701.02151

    %% P %%
    \bibitem[Padmanabhan et al.(2008)]{Padmanabhan2008} Padmanabhan, N., Schlegel,
             D.~J., Finkbeiner, D.~P., et al.\ 2008, \apj, 674, 1217-1233

    \bibitem[Pedregosa et al.(2011)]{scikit-learn} Pedregosa F., et al., 2011,
             Journal of Machine Learning Research, 12, 2825

    %% R %%
    \bibitem[Rowe et al.(2015)]{Rowe2015} Rowe, B.~T.~P., Jarvis, M., Mandelbaum, R.,
             et al.\ 2015, Astronomy and Computing, 10, 121

    %% S %%
    \bibitem[Schlafly et al.(2012)]{Schlafly2012} Schlafly, E.~F., Finkbeiner, D.~P.,
             Juri{\'c}, M., et al.\ 2012, \apj, 756, 158

    \bibitem[Scoville et al.(2007)]{Scoville2007} Scoville, N., Abraham, R.~G.,
             Aussel, H., et al.\ 2007, \apjs, 172, 38

    \bibitem[S{\'e}rsic(1963)]{Sersic1963} S{\'e}rsic, J.~L.\ 1963, Boletin de la
             Asociacion Argentina de Astronomia La Plata Argentina, 6, 41

    \bibitem[Suchyta et al.(2016)]{Suchyta2016} Suchyta, E., Huff, E.~M., Aleksi{\'c},
             J., et al.\ 2016, \mnras, 457, 786

    \bibitem[Steidel et al.(1996)]{Steidel1996} Steidel, C.~C., Giavalisco, M.,
             Pettini, M., Dickinson, M., \& Adelberger, K.~L.\ 1996, \apjl, 462, L17

    %% T %%
    \bibitem[The Dark Energy Survey Collaboration(2005)]{DES2005} The Dark Energy
             Survey Collaboration 2005, arXiv:astro-ph/0510346

    \bibitem[Tonry et al.(2012)]{Tonry2012} Tonry, J.~L., Stubbs, C.~W., Lykke, K.~R.,
             et al.\ 2012, \apj, 750, 99

    %% V %%
    \bibitem[VanderPlas et al.(2014)]{astroml} VanderPlas, J., Fouesneau, M., \&
             Taylor, J.\ 2014, Astrophysics Source Code Library, ascl:1407.018

    %% W %%
    \bibitem[Walt et al.(2011)]{NumPy} Walt S. v. d., Colbert S. C., Varoquaux G.,
             2011, Computing in Science and Engg., 13, 22

\end{thebibliography}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\appendix
%% ------------------------------------------------------------------------------------ %%
\section{Quality Control of Synthetic Objects}
    \label{app:qc}

    To select synthetic stars and galaxies from the HSC survey data in order to test the photometry,
    we apply some basic quality control cuts.

    The HSC survey defines the full--depth \& full--color regions (\texttt{FDFC}) to
  ensure the data used for science reach the expected number of exposures in each
    band
    ($\mathtt{(\#gri\ \geq\ 4)\ and\ (\#yz\ \geq\ 6)\ and\ (Limiting\ imag> 25.6)}$).
    Since the two \tracts{} we used in this work are almost entirely covered in
    the \texttt{FDFC} region, we did not apply this cut.
    We did apply the following cuts to the matched synthetic objects.

    \begin{itemize}

        \item[ ] \texttt{ idetect\_is\_primary==True }
        \item[ ] \texttt{ ideblend\_skipped==False }
        \item[ ] \texttt{ iflags\_badcentroid==False }
        \item[ ] \texttt{ icentroid\_sdss\_flags==False }
        \item[ ] \texttt{ iflags\_pixel\_edge==False }
        \item[ ] \texttt{ iflags\_pixel\_interpolated\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_saturated\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_cr\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_bad==False }
        \item[ ] \texttt{ iflags\_pixel\_suspect\_center==False }
        \item[ ] \texttt{ iflags\_pixel\_clipped\_any==False }
        \item[ ] \texttt{ iflags\_pixel\_bright\_object\_center ==False }

    \end{itemize}

    \todo{Details for stars and galaxies}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\section{$b$: The Blendedness Parameter}
    \label{app:defineb}

    To evaluate the degree to which an object is blended with others, \hscpipe{} introduces
    the blendedness parameter: $b$ (\texttt{[grizy]blendedness\_abs\_flux}).
    We briefly define  $b$ below; please refer to Bosch\etal (in prep.) and Murata\etal (in prep.) for more details.

    For object ${\rm A}$:
    \begin{eqnarray*}
        b({\rm A}) \equiv
        1-\frac{\int_{\mathbb{R}^2} {\rm d}x~ {\rm d}y\ \mathcal{N}_{\mathrm{A}}({\bf x}\ |\  \mu_{\mathrm{A}}, {\bf\Sigma}_{\mathrm{A}})
        F_{\mathrm{A}}({\bf x})}{\int_{\mathbb{R}^2} {\rm d}x~ {\rm d}y\ \mathcal{N}_{\mathrm{A}}({\bf x}\ |\  \mu_{\mathrm{A}},
        {\bf\Sigma}_{\rm{A}}) F_{\rm{total}}({\bf x})},
        \label{eq:defineb}
    \end{eqnarray*}

    \noindent
    where $\mathcal{N}_{\rm{A}}({ \bf x }\ |\  \mu_{\rm{A}}, {\bf\Sigma}_{\rm{A}})$
    is a 2-D Gaussian function at pixel position $\bf x$, $\mu_{\mathrm{A}}$ is the
    estimated centroid of object ${\rm A}$, and covariance ${\bf \Sigma}_{\mathrm{A}}$
    is estimated based on the Gaussian-weighted adaptive second moments of object
    ${\rm A}$ (no PSF correction).
    $F_{\mathrm{total}}({\bf x })$ and $F_{\rm{A}}({\bf x })$ are pixel values of
    ${\rm A}$ before and after \hscpipe{} deblending process, respectively.

    By definition, the parameter is bound from 0 to 1.
    When the deblending process is carried out correctly, the $b({\rm A})$ parameter
    reflects the fraction of fluxes that comes from other objects in the
    region of ${\rm A}$.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%%%%%%%%%%%: Possible Tables %%%%%%%%%%%%%
%\input{table1.tex}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\label{lastpage}
\end{document}
%% ------------------------------------------------------------------------------------ %%