### 2017-08-16 draft/response_to_referee.txt
Response to the referee report; 1

We thank the referee for the careful reading and many valuable comments. Here we attached the updated draft
where important changes are highlighted in cyan.

We believe that all the comments and questions from the referee have been addressed, and below are the
detailed response to some of the referee's concerns.

0. About the general comment that this draft resembles a "technical supplement of the hscPipe paper", we
think:
    * SynPipe paper presents important evaluations of the curent HSC data.  These results are useful to
      the general users of HSC instrument.  Given that the hscPipe paper already contains lots of technical 
      details and is quite long, we believe that a separate paper can better draw people's attention to these
      results. 
    * As we now presented in the new Discussion section, as an independent software, SynPipe is being used for
      many scientific goals, both within and outside the SSP community.  For some of them (e.g. Murata et al.
      in prep. about the impact of blended objects in weak lensing analysis), SynPipe is the key technical
      component.  We think it is useful to have a separate paper as reference to the readers of these works
      and the future SynPipe users.  In addition, SynPipe has recently been adopted by the LSST community, and 
      will be used to test the performance of the LSST pipeline.  We think it is useful to summarize the
      developement of SynPipe on the HSC side as an independent paper. 
    * We would like to draw the referee's attention to papers with the similar purpose, e.g. Balrog for the
      DES survey (http://adsabs.harvard.edu/abs/2016MNRAS.457..786S). These papers are proved to be useful and
      well-cited for similar reasons as we stated above.

# 1.

# 8. p. 7-8: "High level reduction pipeline": as stated more clearly in the hscPipe technical paper (Bosch et
al. 2017). Low-level reduction involves basic image detrending and calibration; high-level processes involves
the ones that generate coadded images and science-ready catalogs.  Now use the same explanation here to be
consistent with the technical paper and provide more details.

# 9. "purity and completeness"

# Minor comments:

1. p.7: should be 'mas' not 'max', corrected.

2. p.10, about 'rectangular cutout': the images are stored in a 2-D array, hence has a rectangular nature to
it.  And, it is much easier to convolve the model and shift the 2-D rectangular array before we add them to
the HSC coadd images.  When we generate the model image array using GalSim, we try to make sure that all fluxes 
of the model is contained in the image. For galaxies, we cutout the model at 10 x Re of the Sersic model. This 
is fine for most galaxies, but still miss small fraction of light for galaxy with very high Sersic index.
First, we try to avoide model with very high Sersic index in the input catalog, as they are often problematic
models for real galaxies; Second, given the magnitude distribution of the model galaxies (most have i>23.0
mag), the results we are showing here are not sensitive to the flux distributions outside 10 x Re.  
The referee is correct that estimation of R_e could be sensitive to the very outskirt when Sersic index is
high, but we'd like to point out that the CModel method used in hscPipe does not use any pixel with
S/N < 5, and only provides n=4 model fits (dev) to the "footprint" of the galaxy.  We do not expect accurate
R_e estimates for high Sersic index galaxies in the beginning.

We add a footnote to briefly explain this choice.

3. p.11, about "noise": it is Poisson noise, and we state that clearly in the text.

4. p.13: We think the location and context of this figure makes it very clear that it is for the input
parameters.  Now we state that clearly in the caption.

5. Fig. 4 and 5: has been corrected.

6.