### 2017-09-13 draft/response_to_referee.txt
All minor comments

We thank the referee for the careful reading and many valuable comments. Here we attached the updated draft
where important changes are highlighted in cyan.

We believe that all the comments and questions from the referee have been addressed, and below are the
detailed response to some of the referee's concerns.

-----
0. About the general comment that this draft resembles a "technical supplement of the hscPipe paper", we
think:
    * SynPipe paper presents important evaluations of the curent HSC data.  These results are useful to
      the general users of HSC instrument.  Given that the hscPipe paper already contains lots of technical 
      details and is quite long, we believe that a separate paper can better draw people's attention to these
      results. 
    * As we now presented in the new Discussion section, as an independent software, SynPipe is being used for
      many scientific goals, both within and outside the SSP community.  For some of them (e.g. Murata et al.
      in prep. about the impact of blended objects in weak lensing analysis), SynPipe is the key technical
      component.  We think it is useful to have a separate paper as reference to the readers of these works
      and the future SynPipe users.  In addition, SynPipe has recently been adopted by the LSST community, and 
      will be used to test the performance of the LSST pipeline.  We think it is useful to summarize the
      developement of SynPipe on the HSC side as an independent paper. 
    * We would like to draw the referee's attention to papers with the similar purpose, e.g. Balrog for the
      DES survey (http://adsabs.harvard.edu/abs/2016MNRAS.457..786S). These papers are proved to be useful and
      well-cited for similar reasons as we stated above.

-----

# 1. Discussion:

We agree with the referee that this work can benefit from a more detailed discussion, which we have added
before the Summary section.  A few relevant points were included in the earlier Summary and Conclusion
section, and have been moved to the new Discussion section.

We emphasize the importance of such synthetic object pipeline for modern cosmology survey, explain again the
philosophy of our desgin, and compare it with other efforts for testing the data reduction pipline.  
Right now, SynPipe is specifically tailored for HSC images, but it will be included in the LSSTpipe in the
near future.  We demonstrate the broad scientific topics that can benefit from SynPipe, and describe the
on-going efforts for further improvements.

-----
# 2 Interpreter Language

The referee is correct that, in general, C++ should be more efficient in the heavy duty data reduction process.
However, SynPipe here mainly works as a special user interface to hscPipe, which is also use Python for
high-level algorithm.  The reason for this choise is described in details in the Section 2 of the technical paper 
for hscPipe (Bosch et al. 2017).  Below is the most relevant paragraph:

    """
    Both the LSST Data Management codebase and the HSC Pipeline are written in a combination of C++ and Python. 
    Flexibility and extensibility are core development goals at all levels of the code-base: we have attempted 
    to make virtually every algorithm not just configurable but replaceable (even by external code). 
    This flexibility is important for LSST in part because many algorithmic choices in LSST processing still 
    require significant research and are under active development.....

    High-level algorithms are written in Python, primarily as subclasses of our Task class...

    Low-level algorithms that perform direct pixel-level processing or other computationally intensive operations 
    are written in C++....
    """

As the referee can see, flexibility and extensibility is the main reason that hscPipe choose to use Python as
the high-level language.  GalSim also uses the same logic.  Since SynPipe 1) directly depends on hscPipe and
GalSim; 2) will develop along with the hscPipe (and LSSTpipe in the future); 3) does not do any low-level
computationally intensive operation, we think it is natural to follow this decision and use Python as our
working language.

The efficiency of SynPipe, as the referee mentioned, certaintly relates to the available cores of the CPU. 
What we meant to explain is that, under the same computational condition, SynPipe test costs similar amount
time as the real data reduction.  Most heavy duty is carried out by hscPipe, so using C or C++ for SynPipe
will not increase the efficiency.  The most time-consuming part of SynPipe test is the injection of synthetic
objects to the individual CCD image.  During this process, the most computationally intensive part is carried
out by GalSim using C++, so using C or C++ for SynPipe here will not increase the efficiency much.  At the
end, we would like to emphasize that the design of SynPipe currently focuses on testing data reduction in the
most reaslistic way, in the future, we will gradually work on increase the efficiency of SynPipe.

Following the suggestion by the referee, we now try to explain this better in Section 2.2, 
Section 3.1, and Section 3.2.3.

-----
# 3.  Missing Definitions

The referee is correct that some of these basic definitions should be explained better.  
Most of these concepts are explained and discussed in great details in the hscPipe paper (Bosch et al. 2017)
and the Data Release 1 Paper (Aihara et al. 2017).  We emphasize the basic definitions of these values at the
end of the introduction section and point the authors to the Bosch+2017 paper for more details.

<1> HSC pipeline performs multi-step photometric calibrations against the Pan-STARRS system, hence all the
fluxes and magnitudes measured by hscPipe are already calibrated into the AB magnitude system.  Therefore, the
pixel values on the calibrated single-Visit and coadd images can be directly related to fluxes in physical
unit, and can be converted into magnitudes using a simple calibration zeropoint (27.0).  Users of HSC survey
data should not worry that the definition of magnitude is affected by filter or exposure time.
Through out this work, we assume that all magnitudes of synthetic objects are also in AB system, and they are 
converted into fluxes using the same zeropoint.

<2> The CCD processing provides us calibrated exposures that include both the detrended, background-subtracted
image and an image containing per-pixel estimate of the variance. The following coaddition process also takes
the variance information into account, uses the inverse of the mean variance of the input images to set the
weights for coaddition, and calculates the per-pixel variance for the coadd images. Both the PSF and CModel
photometry take these variance images into account, and provide meaningful estimations of flux error for the
coadd images.  
    - The HSC pipeline uses matched-filter method to estimate the PSF fluxes.  Per-pixel variance information
      is taken into account when estimating the flux error.  Please see Section 4.9.5 of Bosch+2017 for more
      technical details.  
    - The HSC CModel photometry is modified and improved based on the SDSS version.  It fits linear
      combination of two basic parametric models to the 2-D flux distribution of galaxy after taking PSF
      convolution into account.  It has the advantage of providing consistent multiband photometry for
      extended objects and having reliable performance for very faint object. 
      Please see Section 4.9.9 and Appendix 2 of Bosch+2017 for technical details of the CModel algorithm. 
    - For CModel photometry, it is optional to use per-pixel inverse square root of the variance information
      as weights during the fitting (when usePixelWeights=True).  By default, usePixelWeights=False, and a
      constant weight is used for an object.  Again we will refer the reader to Bosch+2017 for more details
      about this.

<3> We understand the referee's concern, and it is true that based on statistics of photons, we could have
more clear definition of signal-to-noise ratio (S/N).  However, in practice, it is not realistic to use "photon
number" for statistics or separately account for the noise from readout, dark, and sky background.  Each CCD
on the camira has four amplifiers that have slightly different effective gain; individual CCD images need to
be detrended, calibrated, and background subtracted before coaddition; single exposures with different seeings
are warped before they are coadded into a final image using appropriate weighting.  All these processes make
it difficult to trace back to the origianl photon statistics.  For instance, on the coadd image, there are
covariances between adjacent pixels due to the warping process.  On the other hand, hscPipe calibrates the
pixels values on both coadd and single-Visit images to physically meaningful flux scale, and provides
correponding variance information that reflect the per-pixel uncertainties after all these complex reduction
steps.  Based on this design, scientific user of HSC data can use the flux value and its uncertainty from
hscPipe to estimate the S/N of the object under certain photometric method.  
    - Throughout this work, S/N is defined as the ration between a flux measurement and its uncertainty (flux
      / flux_err).  Same definition is adopted by both Bosch+2017 (hscPipe paper) and Aihara+2017 (HSC DR1
      paper), along with a series of science papers.  
    - Under this definition, 5 sigma detection corresponds to a S/N=5 flux measurement under certain
      photometric method.  The same concept is adopted by both Bosch+2017 (hscPipe paper) and Aihara+2017 
      (HSC DR1 paper), along with a series of science papers. 
    - Both the PSF and CModel photometry algorithms convolve a model for the true object’s morphology with 
      the effective PSF model at the position of the object, which maximizes S/N (see Bosch+2017)

The changes of text related to this comment are: 
    - At the end of Introduction section, we make it clear that we are using AB system, and explain the
      definiton of S/N and Nxsigma detection.  
    - At the end of Section 2.2, we provide more details related to the flux measurements for PSF and CModel
      photometry.  We also refer the reader to Bosch+2017 for more technical details. 
    - Section 6.4: slightly more details about the usePixelWeights option and reference to hscPipe paper.

-----
# 4. CModel algorithm

We agree with the referee that the details of CModel algorithm should be explained better.  However, limited
by the scope of this work, we can not afford to show all the technical details.  As mentioned, we expanded
the description in Section 2.2, and points the readers to the relevant sections in Bosch+2017.

Related to the referee's detailed comments: 
    - We should clearly point out that good performace of cModel photometry is **not** a natural consequence
      of hscPipe, and hscPipe **does not** use GalSim model for galaxy photometry.  These are two completely
      independent software, and SynPipe interacts with them at different stages of the test.  In our tests, the
      synthetic galaxy is described using single-Sersic profile, and is generated using GalSim.  For hscPipe,
      we test CModel photometry which uses very different model assumptions and algorithms for galaxy photometry. 
      It first fits an exponential model to the image with both the shape and the amplitude free. 
      Then it fits a de Vaucouleurs model in the same way.  As last step, it fits both models 
      simultaneously, keeping their ellipses fixed at the results from the previous two fits, and 
      allowing only the two amplitudes to vary.
      Also, in CModel, both of these models are actually described by multi-Gaussian approximations to allow
      fast convolution (e.g. Hog & Lang 2013, Sheldon 2014).  
      The CModel algorithm is not designed to model detailed flux distributions of galaxies accurately, but to
      provide robust and efficient photometry for large amount of galaxies, especially for the poorly-resolved
      and/or low S/N ones.  
    - Although these tests are much more than just "internal consistency", we should mention that 1)
      single-Sersic is still oversimplified model for real galaxies; 2) we assume that there is no error is
      PSF modelling.  These are the limitations of our current tests, and we mention them in the discussion
      section now.  
    - About the central pixel value of Sersic model, please refer to Rowe et al. (2015; 
      http://www.sciencedirect.com/science/article/pii/S221313371500013X) for details of this code.  
      The Section 5.3 and Section 6 explain the integration over pixels and image rendering method in great
      details.  We should point out that synthetic galaxies generated by GalSim have been used for various
      photometric and weak lensing tests.  The algorithms involved should be considered well tested.

We modify the text in the following way: 
    - Section 3.2.1: We point the readers to the relevant sections in Rowe+2015 for more details. 
    - At the end of Section 4.2, we clearly point out that CModel uses very different model assumptions, hence
      we are not only testing the internal consistency of CModel pipeline. 
    - In the discussion section, we add discussion about the limitation of our single-Sersic model test.

-----
# 5. Fig 4 and 9.

We understand the referee's point that random noises from photon statistics could be useful for comparison.

However, as we pointed out earlier, given the design of the camera, observational strategy, and the complex
processes of data reduction, it is impossible to make statistics based on the "input photon number".  
Each pixel on the coadd image consists of information from CCDs with different characteristics and from
separated exposures under different seeing and sky conditions.  These individual exposures are first detrended,
calibrated, background subtracted, and warped before they are coadded using weights that are related to their
variance information.  Given these reasons, it is not realistic to conduct tests using photo number and random
statistical errors as the referee requested.

Meanwhile, we should point out that, since hscPipe (and LSSTpipe in the future) is designed to provide
high-level science-ready results for users, it carefully calibrates the fluxes into AB magnitude system
(absolute flux level), which is physically more meaningful and convinient.  In addition, the hscPipe tries to
provide per-pixel variance information for both single-Visit and coadd images, which includes statistical
uncertainties from sources mentioned and not mentioned by the referee (e.g. correction of brighter-fatter
effect, crosstalk, cosmic-ray removal, number of single-visit images that go into the coadd etc.).
For coadd image, the variance plane provided by hscPipe should be treated as a summary of all statistical
errors, and all the flux error in hscPipe is estimated based on it.  Notice that, for both PSF and CModel
photometry, the estimation of statistical uncertainty is not as straightfoward as for aperture photometry.
And it is not realistic for us to provide independent estimations for each object.

As mentioned earlier, such definition of S/N has been adopted within the HSC SSP collaboration, and has
been used for multiple technical and scientific purposes.  Therefore, we think the S/N plotted in Fig 4 and 
Fig 9 are the most meaningful and useful ones for HSC data user.

We slightly modify the text to explain the definition of S/N and flux error better: 
    - In Section 5.1.1, we explain the PSF flux error in more details.
    - In Section 5.2.1, we explain that for extended objects modeled by CModel, definition of S/N is not as
      straightforward as point sources.  Here we choose to use the CModel flux and flux error to simply
      demonstrate the typical "significancy" of galaxies at fixed magnitude, and show that hscPipe can
      relibaly detect and measure flux for galaxies down to certain magnitude.  We also mention that
      uncertainty of CModel flux is dominated by systematics, and here the flux error only summarises the
      statistical ones.  Therefore, our tests should provide more realistic estimation of average flux error
      (including both systematic and statistical ones) for galaxy at certain magnitude.

-----
# 6. Pure statistical errors in figures.

We agree with the referee that it is very useful to compare with the statistical uncertainties. 
However, as we explained for comment #5, "pure statistical random errors" based on photon 
statistics are not very practical to get from hscPipe.

Instead, the flux error from hscPipe for both PSF and CModel photometry are relied on the
detailed variance information that summarize most of the statistical errors. These flux or
magnitude error are also provided to the HSC data user as the most frequently used
information on photometric uncertainties, it is useful to compare them with the results of
our SynPipe tests.

On Fig 5, 6, 7, 10, 11, 12, and 15, we have overplot a pair of grey dashed lines that
outline the running median of statistical uncertainties measured by the hscPipe (in
positive and negative form).  For magnitude, we simply use the magnitude error provided by
hscPipe, for color, we add the magnitude uncertainties for both bands in quadrature.

Given the realistic situation, we think these comparisons cover the requirements by the
referee, although are not exactly the same.  We did not specifically mentioned these
comparesions because the results are exactly as we expected based on the algorithms
themselves: for PSF photometry, the statistical uncertainties from hscPipe are similar but
still slightly lower compared to the median uncertainties from SynPipe test; For CModel,
the statistical flux errors from hscPipe are smaller compared to the uncertainties from
SynPipe test as systematic uncertainties (model-dependent errors) dominate the real error
budget of CModel photometry.

We now briefly mention these comparisons in each section to draw readers attention to
these information.

-----
# 7. Precision and accuracy of PSF magnitude

We believe this comment is also covered by the previous one.  For PSF magnitude, the
statistical uncertainties obtained by hscPipe is similar but still slightly lower than the
photometric uncertainties based on the SynPipe test.  As mentioned, due to the correlated
noise from the image warping and coadding process, the statistical uncertainties of PSF
flux measured by hscPipe is likely to be over-optimistic.

Relevant text has been added to Section 5.1.2 and 5.1.3.

-----
# 8. p. 7-8: "High level reduction pipeline": as stated more clearly in the hscPipe technical paper (Bosch et
al. 2017). Low-level reduction involves basic image detrending and calibration; high-level processes involves
the ones that generate coadded images and science-ready catalogs.  Now use the same explanation here to be
consistent with the technical paper and provide more details.

-----
# 9. "purity and completeness"

We agree with the referee that the purity and completeness tests are very important for
survey like this, and SynPipe can be a useful tool for this purpose.

At the same time, careful tests for this purpose are beyond the scope of this paper, where
we simply want to demonstrate the application of SynPipe and provide simple photometric
benchmark tests for stars and galaxies.

Following the suggestions from the referee, we modify the text with more discussions on
this, and here are the main points:
    - Section 4.2: We explain that it is beyond the scope of this work to look into this
      carefully.  Detection limits and completeness are normally defined for point sources.  
      As we show using the PSF S/N, it depends on the seeing condition and show spatial 
      changes which requires more careful designs and higher density of fake stars. 
      Detection completeness of point sources for HSC DR1 is investigated in Aihara+2017
      (see Section 5.7) using a different method. 
    - Misclassification is briefly discussed in Section 6.2. The design of our tests and the 
      number density of synthetic galaxies are not perfect to thoroughly test the star-galaxy 
      separation. So our results just demonstrate the capability and briefly discuss the 
      current situation for star/galaxy separation.  We find that, right now, hscPipe tends 
      to provide a very complete sample for galaxy but with considerable fraction of star 
      contamination at the faint end.  In Aihara+2017, the star/galaxy separation issue 
      is discussed in Section 5.7, where they draw the same conclusion using different 
      method. 
    - The current method adopted by hscPipe focuses on the magnitude different between PSF 
      and CModel photometry and is known to have limitations (see Section 4.9.10 of 
      Bosch+2017) and issues (e.g. the one related to the current CModel algorithm; see 
      Appendix 2 in Bosch+2017).  All these conclusions are reflected in our simple test. 
      On the other hand, hscPipe team is working on replacing the star/galaxy separation 
      algorithm with a much improved one (based on extreme deconvolution method, also 
      see Section of 4.9.10 of Bosch+2017).  We will test its performance after the next 
      data release. 
    - We expand the discussion in Section 6.2 to provide more information about star/galaxy 
      separation tests and emphasize the importance of such tests as suggested by the referee.

-----
# Minor comments:

----
1. p.7: should be 'mas' not 'max', corrected.

----
2. p.10, about 'rectangular cutout': the images are stored in a 2-D array, hence has a rectangular nature to
it.  And, it is much easier to convolve the model and shift the 2-D rectangular array before we add them to
the HSC coadd images.  When we generate the model image array using GalSim, we try to make sure that all fluxes 
of the model is contained in the image. For galaxies, we cutout the model at 10 x Re of the Sersic model. This 
is fine for most galaxies, but still miss small fraction of light for galaxy with very high Sersic index.
First, we try to avoide model with very high Sersic index in the input catalog, as they are often problematic
models for real galaxies; Second, given the magnitude distribution of the model galaxies (most have i>23.0
mag), the results we are showing here are not sensitive to the flux distributions outside 10 x Re.  
The referee is correct that estimation of R_e could be sensitive to the very outskirt when Sersic index is
high, but we'd like to point out that the CModel method used in hscPipe does not use any pixel with
S/N < 5, and only provides n=4 model fits (dev) to the "footprint" of the galaxy.  We do not expect accurate
R_e estimates for high Sersic index galaxies in the beginning.

We add a footnote to briefly explain this choice.

----
3. p.11, about "noise":

It is Poisson noise, and we state that clearly in the text.

----
4. p.13:

We think the location and context of this figure makes it very clear that it is for the input
parameters.  Now we state that clearly in the caption.

----
5. Fig. 4 and 5: has been corrected.

----
6. p.22:

For stars, first, fraction of stars with such extreme colors are very low.  The
discrepency can barely be noticed on the comparisons of 1-D color distributions. 
We think the likely reasons for this are: 
    - There are still scattered light on the y-band images that are not accounted for.
      This known prolblem can affect the accuracy of y-band photometry and is discussed in 
      Section 5.8.14 in Aihara+2017. 
    - Background level is much higher in z and y-band, and background subtraction could
      suffer from higher uncertainties.  Right now, SynPipe can still not test the impact
      of background subtraction on photometry, and it will be investigated in the future. 
    - We modify text in Section 5.1.3 to breifly explain this.

As for galaxies, the same background and y-band scatter light issues also exist.  However,
the dynamical ranges for i-z and z-y colors are narrower.  Also the CModel color
uncertainties are higher than the PSF ones.  Both of these factors make it harder to spot
discrepencies of color distributions.  However, we still see discrenpencies between the
input and recovered z-y colors. 
    - We expand the discussion in Section 5.2.3 a little to address this issue.

----
7. p.22

The referee is correct that, for highly blended stars, the biases on their PSF colors are 
still quite big.  What we meant here is that the biases for PSF magnitude are much worse 
as the total fluxes of blended stars are highly underestimated.  As for PSF colors, 
although there are still long tails that indicate the biases, at least the peak of the 
distributions are around zero now.  We adjust this sentence to make it more clear.

As for general comments for PSF photometry of highly blended stars.  Unfortunately we are 
not perfectly clear what causes this issue.  It is clearly related to the performace of 
debelnder in hscPipe, which is know to still have issue and is being improved now 
(please see Section 4.8 of Bosch+2017).  Given the depth of HSC survey, the situation for 
object deblending becomes much more tricky.  At this point, we have to leave this as a 
caveat.  PSF magnitude is still considered the best photometry for point sources, but we 
will remind the users to avoid using it for highly blended objects.  The text has been 
modified to reflect this point.

----
8. P.29

Actually it should be (z-y) color, and has been corrected.

----
9. Astrometric Calibration Fig 14

HSC image has pixel scale of 0.168 arcsec/pixel.  We use arcsec as unit since when evaluating the astrometric
calibration for HSC DR1, arcsec or mas are used in Aihara+2017.  Information about pixel scale is added to the
caption of the figure and the text.

----
10. the definition of highly blended

The referee is correct, that we should use consistent definition for highly blended
objects, which is b>0.05.  We have make it consistent across the draft.

---- 
11. Table 1, 2, 3, and 4

The statistical uncertainty is actually the standard deviation of (mag_output - mag_input)
within each input magnitude bin.  This is the what we used to describe "precision". 
For accuracy, we use the mean value of magnitude difference in each input magnitude bin.

All the captions of tables have been updated to make it more clear.